### 2024-02-28 slides/data.tex
slides
27/02 presentation

\section{Data}

\begin{transitionframe}

    \rmfamily % Font

    \begin{center}
    {\Huge \textbf{\textcolor{white}{Data}}}
    \end{center}

\end{transitionframe}

\begin{frame}

    \frametitle{Data} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item \textcolor{fgre}{USBLS's LAUS Database} \(\to\) county-level unemployment statistics (1990-2023)
        \item \textcolor{fgre}{US Census Bureau's Population Estimates Program (PEP)} \(\to\) county-level population and demographics (1969-2020)
        \item \textcolor{fgre}{USDOL's Changes in Basic Minimum Wages} (1968-2023)
        \item \textcolor{fgre}{Horwitz et al. (2020)} \(\to\) Prescription Drug Monitoring Programs (1990-2019)
    \end{wideitemize}

    \vspace{9pt} 
    Current working sample:
    \vspace{9pt}

    \begin{wideitemize}
        \item Unemployment, PDMPs, Minimum Wage rates and demographics for counties of 50 States + DC, 1998-2019
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{PDMPs} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    Prescription Drug Monitoring Programs (\textcolor{fblu}{PDMPs}): electronic database that tracks controlled substance prescriptions
    \vspace{9pt}

    \begin{wideitemize}
        \item \textcolor{fblu}{Modern System PDMPs}: the electronic database becomes accessible to any authorized user (eg, physician, pharmacist, or member of law enforcement)
        \item \textcolor{fblu}{Must Query PDMPs}: the law mandates the prescriber to check the database before prescribing a listed opioid
    \end{wideitemize}

\end{frame}

### 2024-02-28 slides/estimation.tex
slides
27/02 presentation

\section{Estimation}

\begin{transitionframe}

    \rmfamily % Font

    \begin{center}
    {\Huge \textbf{\textcolor{white}{Estimation}}}
    \end{center}

\end{transitionframe}

\begin{frame}

    \frametitle{Empirical model} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Individual-level event study for different minimum wage quantiles: \textcolor{fgre}{Arkhangelsky, Kazuharu and, Zohar (in progress)}
    \end{wideitemize}

    \begin{equation*}
        u_{it} = \alpha_i + \lambda_t + \sum_{k\geq h}\tau_{ik} + \beta X_{it} + \nu_{it}
    \end{equation*}
    with 
    \vspace{9pt}
    \begin{wideitemize}
        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{i}\): county, \(\textcolor{fblu}{t}\): time (year-month), \(\textcolor{fblu}{k = t - z_i}\): time difference from PDMP implementation year in county \(\textcolor{fblu}{i}\), \(\textcolor{fblu}{z_i}\)
        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{u_{it}}\): unemployment rate, \(\textcolor{fblu}{X_{it}}\): demographic covariates, \(\textcolor{fblu}{\nu_{it}}\): idiosyncratic shock 
    \end{wideitemize}
    \vspace{9pt}
    State-level changes must be exogenous at the county level for \textcolor{fblu}{identification} and no anticipation up to \(\textcolor{fblu}{h}\) periods

\end{frame}

\begin{frame}

    \frametitle{Results} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{center}
        \includegraphics[scale=0.45]{mop_effects.png}
    \end{center}

\end{frame}

\begin{frame}

    \frametitle{Results} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{center}
        \includegraphics[scale=0.45]{pmq_effects.png}
    \end{center}

\end{frame}

\begin{frame}

    \frametitle{What is to be done?} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Minimum wage measures of bindingness
        \vspace{9pt}
        \begin{wideitemize}
            \item Idea: we can estimate the effect of minimum wage on employment for wage bins
            \item Then assign, if significant, the coefficient to each county
            \item Sector compositions of the counties can also be relevant
        \end{wideitemize}
        \item I'm \textcolor{fblu}{open to suggestions}
        \item Participation rates might capture some of the effect
        \item Distribution of individual effects
    \end{wideitemize}

\end{frame}

### 2024-02-28 slides/introduction.tex
slides
27/02 presentation

\section{Introduction}

\begin{frame}

    \frametitle{The Opioid Crisis} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item 110,000 Americans died in 2022 as a consequence of drug abuse
    \end{wideitemize}

    \begin{center}
        \includegraphics[scale=0.5]{ODD.png}
    \end{center}

    {\footnotesize \textit{Opioid deaths for both the non-college and college-educated as measured per 100,000 people in the respective education class (\textcolor{fgre}{Greenwood, Guner and Kopecky, 2022})}}

\end{frame}

\begin{frame}

    \label{Minimum Wage}
    \frametitle{State minimum wage trends} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize} 
        \item States have been increasing their \textcolor{fblu}{minimum wage rates}, with varying intensity
    \end{wideitemize}

    \begin{center}
        \includegraphics[scale=0.5]{min_wage_plot_simp.png}
    \end{center}

    \hyperlink{min_wage_plot_allstates}{\beamerbutton{All states}}

\end{frame}

\begin{frame}

    \frametitle{Main idea} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Opioid misuse can lead the victims to negative labor market outcomes
        \item Minimum wage rates affect these outcomes 
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] Negatively, through their impact on \textcolor{fblu}{labor demand for low-skill jobs} 
            \item[\textcolor{fblu}{\textbullet}] Positively, if they address the reasons that cause \textcolor{fblu}{addiction} in the first place \(\to\) \textcolor{fblu}{Deaths of despair}
        \end{wideitemize}
        \item The dominance of each effect depends on how \textcolor{fblu}{binding} minimum wage rates are across labor markets
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{In this paper I ask} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Do the effects of \textcolor{fblu}{Opioid Use Disorders} (\textcolor{fblu}{OUDs}) on labor market outcomes interact with minimum wage policies?
        \item If so, how have these policies shaped the ongoing Crisis?
        \item Is the interaction homogeneous across States, or can we expect differences across labor markets?
    \end{wideitemize}

    \vspace{9pt}
    I will use \textcolor{fblu}{changes in state prescription drug regulations} as a source of variation
    \vspace{9pt}

    \begin{wideitemize}
        \item These changes can be seen as \textcolor{fblu}{exogeneous} from the \textcolor{fblu}{county level}
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{Policy interactions} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item More stringent regulations on opioid access can make the victims to 
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Quit}, which might imply some disutility, or
            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Switch to the black market}
        \end{wideitemize}
        \item This decision affects the MPL of the worker
        \item In this scenario, minimum wage increases can
        \vspace{9pt}
        \begin{wideitemize}
            \item[\textcolor{fblu}{\textbullet}] Improve labor outcomes if better living conditions tackle the causes of addiction and relapsing (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding} after the change)
            \item[\textcolor{fblu}{\textbullet}] Foster addiction, if quitting doesn't improve employment prospects because \textcolor{fblu}{minimum wage becomes binding} 
        \end{wideitemize}
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{Contribution} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item \textcolor{fgre}{Cengiz et al. (2019)} \(\to\) minimum wage changes reduced employment in low-wage jobs of tradeable sectors
        \item \textcolor{fgre}{Aliprantis, Fee, and Schweitzer (2023)} \(\to\) large and significant negative impact of prescription rates on labor force participation
        \item \textcolor{fgre}{Alpert, Powell, and Pacula (2018)} \(\to\) positive relation between OxyContin misuse and heroin deaths 
    \end{wideitemize}

    \vspace{9pt}
    \textbf{Contribution}: shedding light on how the policy responses to the \textcolor{fblu}{Opioid Crisis} have been affected by \textcolor{fblu}{minimum wage changes}
    \vspace{9pt}

    \begin{wideitemize}
        \item Interest in how \textcolor{fblu}{labor market regulations} interact with \textcolor{fblu}{Substance Abuse Disorders}
    \end{wideitemize}

\end{frame}

### 2024-02-28 main.R
estimates
(in progress)

@@ -1,0 +1,189 @@
+
+rm(list=ls()) 
+set.seed(123)
+
+library(tidyverse)
+library(fixest)
+
+# Set the path
+file_location <- rstudioapi::getSourceEditorContext()$path
+setwd(dirname(file_location)) #set path to location
+
+# Load the data
+data_location <- paste(getwd(),"data/joined_data.csv",sep = "/")
+df <- read.csv(data_location,header=TRUE,sep=",")
+
+# Prepare the covariates
+names_covar <- names(df[,grep("_ratio$", colnames(df))])
+covariates <- paste(names_covar, collapse = " + ")
+
+# One-side length of effects
+window = 24
+
+# Not yet treated (mop)
+not_yet_treated_mop <- df[df[["time_marker"]] < df$first_treatment_mop,]
+
+# Not yet treated (pmq)
+not_yet_treated_pmq <- df[df[["time_marker"]] < df$first_treatment_pmq,]
+
+# First stage (mop)
+first_stage_mop <- fixest::feols(
+  stats::as.formula(paste0("urate ~ ", covariates, " |", "fips", " + ", "time_marker")),
+  data = not_yet_treated_mop,
+  combine.quick = FALSE,
+  warn = FALSE,
+  notes = FALSE)
+
+# First stage (pmq)
+first_stage_pmq <- fixest::feols(
+  stats::as.formula(paste0("urate ~ ", covariates, " |", "fips", " + ", "time_marker")),
+  data = not_yet_treated_pmq,
+  combine.quick = FALSE,
+  warn = FALSE,
+  notes = FALSE)
+
+# Fitted values and diffs.
+df[[paste0("urate", "_hat_mop")]] <- stats::predict(first_stage_mop, newdata = df) 
+df[[paste0("urate", "_tilde_mop")]] <- df[["urate"]] - df[[paste0("urate", "_hat_mop")]]
+
+df[[paste0("urate", "_hat_pmq")]] <- stats::predict(first_stage_pmq, newdata = df) 
+df[[paste0("urate", "_tilde_pmq")]] <- df[["urate"]] - df[[paste0("urate", "_hat_pmq")]]
+
+# Initialize matrices
+effects_mop <- matrix(0,length(unique(df$state)),2*window+1)
+constants_mop <- matrix(0,length(unique(df$state)),1)
+
+effects_pmq <- matrix(0,length(unique(df$state)),2*window+1)
[…]

### 2024-02-28 data/data_merge.R
data_merge changes

@@ -4,10 +4,18 @@ set.seed(123)
 
 library(tidyverse)
 
+# Set path, load the data
+
 setwd(dirname(rstudioapi::getSourceEditorContext()$path))
-unemp <- read.csv(paste(getwd(),"unemployment.csv",sep = "/"),header=TRUE,sep=",")
-horwitz <- read.csv(paste(getwd(),"horwitz2020.csv",sep = "/"),header=TRUE,sep=",")
-minwg <- read.csv(paste(getwd(),"minwage_clean_ub.csv",sep = "/"),header=TRUE,sep=",")
+
+unemp <- read.csv(file.path(getwd(), "unemployment.csv"), header = TRUE, sep = ",")
+horwitz <- read.csv(file.path(getwd(), "horwitz2020.csv"), header = TRUE, sep = ",")
+minwage <- read.csv(file.path(getwd(), "minwage_clean_fed_ub.csv"), header = TRUE, sep = ",")
+demographics <- read.csv(file.path(getwd(), "county_demographics.csv"), header = TRUE, sep = ",")
+
+# PDMPs
+
+## Divide months and years in PDMPs dataset
 
 horwitz <- horwitz %>% 
   separate_wider_delim(Enactment, delim = "-", names = c("enact_month", "enact_year")) %>% 
@@ -17,13 +25,21 @@ horwitz <- horwitz %>%
   separate_wider_delim(Prescriber.must.query, delim = "-", names = c("pmq_month", "pmq_year"),too_few = "align_start") %>% 
   mutate_if(is.character, list(~na_if(.,""))) 
 
+## Numeric values, select the policies
+
 horwitz$enact_month <- replace(horwitz$enact_month,horwitz$enact_month == "Pre","0")
 horwitz[,2:length(horwitz)] <- sapply(horwitz[,2:length(horwitz)], as.numeric)
 horwitz <- horwitz[c("state", "pmq_year", "pmq_month", "mop_year", "mop_month")]
 
+## Add time markers
+
 horwitz <- horwitz %>% 
   mutate(first_treatment_pmq = (pmq_year-1960)*12 + pmq_month) %>%
   mutate(first_treatment_mop = (mop_year-1960)*12 + mop_month) 
+
+# Unemployment
+
+## Rename, time marker, select and numeric values
 
 unemp <- unemp %>% 
   rename(state = state_name) %>%
@@ -33,18 +49,36 @@ unemp <- unemp %>%
[…]

### 2024-02-28 data/try_main.R

@@ -1,8 +1,0 @@
-
-install.packages("reticulate")
-
-library(reticulate)
-use_python("C:/Users/guill/AppData/Local/Programs/Python/Python311/python.exe")
-py_run_file("C:/Users/guill/Documents/GitHub/masters_thesis/data/minwage.py")
-
-py_run_file("C:/Users/guill/Documents/GitHub/masters_thesis/data/ahh.py")

### 2024-02-28 data/minwage_cleaning.py
speeed

@@ -1,0 +1,48 @@
+
+'''
+returns minwage_clean_states_ub.csv and minwage_clean_fed_ub.csv
+
+'''
+
+import pandas as pd
+
+df = pd.read_csv(r'./minwage_raw.csv',sep=',')
+
+df = df.drop(['1968 (a)','1970 (a)','1972','1976 (a)','1979','1980','1981','1988'],axis=1)
+df = df[df['State or otherjurisdiction'] != 'Guam']
+df = df[df['State or otherjurisdiction'] != 'Puerto Rico']
+df = df[df['State or otherjurisdiction'] != 'U.S. Virgin Islands']
+df = df.replace({'\$':''}, regex=True)
+
+for i in range(len(df.iloc[:,0])):
+    for j in range(len(df.columns)):
+        if df.iloc[i,j] == '...':
+            df.iloc[i,j] = df.iloc[0,j]
+
+df = df.replace({'\[c\]':''}, regex=True).replace({'\(d\)':''}, regex=True).replace({'\(e\)':''}, regex=True)
+df = df.replace({'\(g\)':''}, regex=True).replace({'\(h\)':''}, regex=True)
+
+for i in range(len(df.iloc[:,0])):
+    for j in range(len(df.columns)):
+        if '-' in df.iloc[i,j]:
+            df.iloc[i,j] = df.iloc[i,j].split('-')[1].lstrip().split(' ')[0]
+        elif '/' in df.iloc[i,j]:
+            df.iloc[i,j] = df.iloc[i,j].split('/')[1].lstrip().split(' ')[0]
+
+df.iloc[:,1:] = df.iloc[:,1:].replace({' ':''}, regex=True)
+df.iloc[:,1:-1] = df.iloc[:,1:-1].astype(float)
+df = df.melt(id_vars="State or otherjurisdiction", 
+        var_name="Year", 
+        value_name="Value")
+
+df.to_csv(r'./minwage_clean_states_ub.csv',sep=',',index=False)
+
+new_df = pd.DataFrame()
+for i in df['Year'].unique():
+    year_data = df[df['Year'] == i]
+    fed_value = year_data.loc[year_data['State or otherjurisdiction'] == 'Federal (FLSA)', 'Value'].values[0]
+    year_data['Value'] = year_data['Value'].apply(lambda x: max(x, fed_value))
+    new_df = pd.concat([new_df,year_data])
+new_df = new_df.reset_index(drop=True)
+
[…]

### 2024-02-28 data/county_unemp.py
speed, again

@@ -1,0 +1,32 @@
+
+'''
+returns unemployment.csv
+
+'''
+
+import pandas as pd
+
+df = pd.read_csv(r'./ladata64County.txt',sep="\t")
+df = df.drop('footnote_codes',axis=1)
+df = df.rename(columns={df.columns.values.tolist()[0]: 'series_id', df.columns.values.tolist()[1]: 'year',
+                        df.columns.values.tolist()[2]: 'period', df.columns.values.tolist()[3]: 'value',})
+
+df['state_fip'] = df['series_id'].str.slice(5,7)
+df['county_fip'] = df['series_id'].str.slice(7,10)
+df["fips"] = df["state_fip"] + df["county_fip"]
+df['series'] = df['series_id'].str.slice(18,20)
+df['month'] = df['period'].str.slice(1,3)
+
+df = df.drop(df[(df.series != '03') | (df.period == 'M13')].index)
+df = df.drop('series_id',axis=1)
+df = df.drop('series',axis=1)
+df = df.drop('period',axis=1)
+df = df.drop('state_fip',axis=1)
+df = df.drop('county_fip',axis=1)
+
+dfips = pd.read_csv(r'./county_fips_master.csv',encoding = "ISO-8859-1")
+dfips = dfips[['fips','county_name','state_abbr','state_name']]
+dfips['fips'] = dfips['fips'].astype(str).str.rjust(5,'0')
+
+df_final = df.merge(dfips, left_on='fips', right_on='fips')
+df_final.to_csv(r'./unemployment.csv',sep=',',index=False)

### 2024-02-28 data/county_unemp.py
comments

@@ -17,6 +17,29 @@ df["fips"] = df["state_fip"] + df["county_fip"]
 df['series'] = df['series_id'].str.slice(18,20)
 df['month'] = df['period'].str.slice(1,3)
 
+'''
+measure_code	measure_text
+
+03	unemployment rate
+
+04	unemployment
+
+05	employment
+
+06	labor force
+
+07	employment-population ratio
+
+08	labor force participation rate
+
+09	civilian noninstitutional population
+
+https://users.nber.org/~notom/for_tal/LAUS/readme.txt 
+
+M=Monthly, M13=Annual
+
+'''
+
 df = df.drop(df[(df.series != '03') | (df.period == 'M13')].index)
 df = df.drop('series_id',axis=1)
 df = df.drop('series',axis=1)

### 2024-02-28 data/minwage_cleaning.py

@@ -6,7 +6,9 @@ returns minwage_clean_states_ub.csv and minwage_clean_fed_ub.csv
 
 import pandas as pd
 
-df = pd.read_csv(r'./minwage_raw.csv',sep=',')
+df = pd.read_csv(r'./minwage_raw.csv',sep=',') # see minwage.py
+
+# drop non-relevant obs
 
 df = df.drop(['1968 (a)','1970 (a)','1972','1976 (a)','1979','1980','1981','1988'],axis=1)
 df = df[df['State or otherjurisdiction'] != 'Guam']
@@ -14,13 +16,52 @@ df = df[df['State or otherjurisdiction'] != 'Puerto Rico']
 df = df[df['State or otherjurisdiction'] != 'U.S. Virgin Islands']
 df = df.replace({'\$':''}, regex=True)
 
+# decisions
+
+'''
+> from: https://www.debt.org/jobs/minimum-wage/#:~:text=What%20Happens%20if%20a%20State's,they%20can%20be%20paid%20less.
+>
+> **Do States Have to Follow the Federal Minimum Wage?**
+>
+> States are required to follow federal minimum wage law. States can pass their own laws to make the wage higher, equal to or lower than the federal law, but they can’t make >other changes that overrule the federal law, for instance, who is exempt or how many hours constitutes a work week.
+>
+> Federal minimum wage overrides state if the state wage is lower. It doesn’t if the state wage is higher. Hourly workers who come under Fair Labor Standard Act guidelines, which >most hourly workers do, always get the higher wage. Some cities have higher minimum wage laws than their state wage. In those cases, too, the higher wage prevails.
+>
+> If a state has no minimum wage law, the federal wage prevails.
+
+'''
+
+# apply effective minimum wage
+
 for i in range(len(df.iloc[:,0])):
     for j in range(len(df.columns)):
         if df.iloc[i,j] == '...':
             df.iloc[i,j] = df.iloc[0,j]
 
+'''
+> [c] - Rates applicable to employers of four or more.
+>
+> (d) - Rates applicable to employers of six or more. In West Virginia, applicable to employers of six or more in one location.
+>
+> (e) Rates applicable to employers of two or more.
+
+'''
+
+# select higher minimum wage
+
 df = df.replace({'\[c\]':''}, regex=True).replace({'\(d\)':''}, regex=True).replace({'\(e\)':''}, regex=True)
+
+'''
[…]

### 2024-02-28 data/min_wage_plot.R
plots
cleaner  code

@@ -1,7 +1,0 @@
-
-rm(list=ls()) 
-
-library(tidyverse)
-
-setwd(dirname(rstudioapi::getSourceEditorContext()$path))
-minwg <- read.csv(paste(getwd(),"minwage_clean_ub.csv",sep = "/"),header=TRUE,sep=",")

### 2024-02-28 data/plots.R

@@ -1,0 +1,137 @@
+
+rm(list=ls()) 
+
+library(tidyverse)
+library(extrafont)
+library(gridExtra)
+font_install('fontcm')
+
+# Set the path
+file_location <- rstudioapi::getSourceEditorContext()$path
+setwd(dirname(file_location)) #set path to location
+
+# Load the data
+minwg <- read.csv(paste(getwd(),"minwage_clean_fed_ub.csv",sep = "/"),header=TRUE,sep=",")
+ODD <- read.csv(paste(getwd(),"OD_deaths_educ.csv",sep = "/"),header=TRUE,sep=",")
+
+# Transform the data
+
+minwg <- as_tibble(minwg)
+minwg <- minwg %>% 
+  rename(year = Year) %>%
+  rename(state = State.or.otherjurisdiction) %>%
+  rename(minw = Value)
+minwg <- minwg[minwg$state != 'Federal (FLSA)',]
+minwg <- minwg[minwg$year != 2023,]
+missing_minw <- minwg[minwg$year == 1998,] %>% mutate(year = 1999)
+minwg <- rbind(minwg,missing_minw) 
+
+ODD$perc_coll_educ <- gsub(',', '.', ODD$perc_coll_educ)
+ODD$perc_coll_educ <- as.numeric(ODD$perc_coll_educ)
+
+ODD <- ODD %>%
+  mutate(total_col_pop = total_pop*(perc_coll_educ/100)) %>%
+  mutate(total_non_col_pop = total_pop*(1 - perc_coll_educ/100)) %>%
+  mutate(Any_Non_Coll_pc = Any_Non_Coll/total_non_col_pop*100000) %>%
+  mutate(Heroin_Non_Coll_pc = Heroin_Non_Coll/total_non_col_pop*100000) %>%
+  mutate(Prescription_Non_Coll_pc = Prescription_Non_Coll/total_non_col_pop*100000) %>%
+  mutate(Synthetic_Non_Coll_pc = Synthetic_Non_Coll/total_non_col_pop*100000) %>%
+  mutate(Any_Coll_pc = Any_Coll/total_col_pop*100000) %>%
+  mutate(Heroin_Coll_pc = Heroin_Coll/total_col_pop*100000) %>%
+  mutate(Prescription_Coll_pc = Prescription_Coll/total_col_pop*100000) %>%
+  mutate(Synthetic_Coll_pc = Synthetic_Coll/total_col_pop*100000)
+
[…]

### 2024-02-28 data/data_merge.R
new pipe

@@ -17,12 +17,12 @@ demographics <- read.csv(file.path(getwd(), "county_demographics.csv"), header =
 
 ## Divide months and years in PDMPs dataset
 
-horwitz <- horwitz %>% 
-  separate_wider_delim(Enactment, delim = "-", names = c("enact_month", "enact_year")) %>% 
-  separate_wider_delim(Enactment.funding.contingent, delim = "-", names = c("enact_fc_month", "enact_fc_year"),too_few = "align_start") %>%
-  separate_wider_delim(Enactment.electronic, delim = "-", names = c("enact_e_month", "enact_e_year"),too_few = "align_start") %>% 
-  separate_wider_delim(Modern.system.operational, delim = "-", names = c("mop_month", "mop_year"),too_few = "align_start") %>% 
-  separate_wider_delim(Prescriber.must.query, delim = "-", names = c("pmq_month", "pmq_year"),too_few = "align_start") %>% 
+horwitz <- horwitz |> 
+  separate_wider_delim(Enactment, delim = "-", names = c("enact_month", "enact_year")) |> 
+  separate_wider_delim(Enactment.funding.contingent, delim = "-", names = c("enact_fc_month", "enact_fc_year"),too_few = "align_start") |>
+  separate_wider_delim(Enactment.electronic, delim = "-", names = c("enact_e_month", "enact_e_year"),too_few = "align_start") |> 
+  separate_wider_delim(Modern.system.operational, delim = "-", names = c("mop_month", "mop_year"),too_few = "align_start") |> 
+  separate_wider_delim(Prescriber.must.query, delim = "-", names = c("pmq_month", "pmq_year"),too_few = "align_start") |> 
   mutate_if(is.character, list(~na_if(.,""))) 
 
 ## Numeric values, select the policies
@@ -33,18 +33,18 @@ horwitz <- horwitz[c("state", "pmq_year", "pmq_month", "mop_year", "mop_month")]
 
 ## Add time markers
 
-horwitz <- horwitz %>% 
-  mutate(first_treatment_pmq = (pmq_year-1960)*12 + pmq_month) %>%
+horwitz <- horwitz |> 
+  mutate(first_treatment_pmq = (pmq_year-1960)*12 + pmq_month) |>
   mutate(first_treatment_mop = (mop_year-1960)*12 + mop_month) 
 
 # Unemployment
 
[…]

### 2024-02-28 data/plots.R

@@ -17,28 +17,28 @@ ODD <- read.csv(paste(getwd(),"OD_deaths_educ.csv",sep = "/"),header=TRUE,sep=",
 # Transform the data
 
 minwg <- as_tibble(minwg)
-minwg <- minwg %>% 
-  rename(year = Year) %>%
-  rename(state = State.or.otherjurisdiction) %>%
+minwg <- minwg |> 
+  rename(year = Year) |>
+  rename(state = State.or.otherjurisdiction) |>
   rename(minw = Value)
 minwg <- minwg[minwg$state != 'Federal (FLSA)',]
 minwg <- minwg[minwg$year != 2023,]
-missing_minw <- minwg[minwg$year == 1998,] %>% mutate(year = 1999)
+missing_minw <- minwg[minwg$year == 1998,] |> mutate(year = 1999)
 minwg <- rbind(minwg,missing_minw) 
 
 ODD$perc_coll_educ <- gsub(',', '.', ODD$perc_coll_educ)
 ODD$perc_coll_educ <- as.numeric(ODD$perc_coll_educ)
 
-ODD <- ODD %>%
-  mutate(total_col_pop = total_pop*(perc_coll_educ/100)) %>%
-  mutate(total_non_col_pop = total_pop*(1 - perc_coll_educ/100)) %>%
-  mutate(Any_Non_Coll_pc = Any_Non_Coll/total_non_col_pop*100000) %>%
-  mutate(Heroin_Non_Coll_pc = Heroin_Non_Coll/total_non_col_pop*100000) %>%
-  mutate(Prescription_Non_Coll_pc = Prescription_Non_Coll/total_non_col_pop*100000) %>%
-  mutate(Synthetic_Non_Coll_pc = Synthetic_Non_Coll/total_non_col_pop*100000) %>%
-  mutate(Any_Coll_pc = Any_Coll/total_col_pop*100000) %>%
-  mutate(Heroin_Coll_pc = Heroin_Coll/total_col_pop*100000) %>%
-  mutate(Prescription_Coll_pc = Prescription_Coll/total_col_pop*100000) %>%
+ODD <- ODD |>
+  mutate(total_col_pop = total_pop*(perc_coll_educ/100)) |>
+  mutate(total_non_col_pop = total_pop*(1 - perc_coll_educ/100)) |>
+  mutate(Any_Non_Coll_pc = Any_Non_Coll/total_non_col_pop*100000) |>
[…]

### 2024-02-28 main.R

@@ -72,8 +72,8 @@ for (i in unique(df_mop$state)){
   
   range_mop = (constants_mop[j] - window):(constants_mop[j] + window)
   
-  avg_effects_mop <- df_mop[(df_mop$state == i)&(df_mop$time_marker %in% range_mop),] %>%
-    dplyr::group_by(time_marker) %>%
+  avg_effects_mop <- df_mop[(df_mop$state == i)&(df_mop$time_marker %in% range_mop),] |>
+    dplyr::group_by(time_marker) |>
     summarise(avg = mean(urate_tilde_mop, na.rm = T))
   
   effects_mop[j,] <- matrix(avg_effects_mop$avg,1,length(range_mop))
@@ -87,8 +87,8 @@ for (i in unique(df_pmq$state)){
   
   range_pmq = (constants_pmq[j] - window):(constants_pmq[j] + window)
   
-  avg_effects_pmq <- df_pmq[(df_pmq$state == i)&(df_pmq$time_marker %in% range_pmq),] %>%
-    dplyr::group_by(time_marker) %>%
+  avg_effects_pmq <- df_pmq[(df_pmq$state == i)&(df_pmq$time_marker %in% range_pmq),] |>
+    dplyr::group_by(time_marker) |>
     summarise(avg = mean(urate_tilde_pmq, na.rm = T))
   
   effects_pmq[j,] <- matrix(avg_effects_pmq$avg,1,length(range_pmq))
@@ -104,7 +104,7 @@ for (i in unique(df_mop$state)){
   first_ts[j,4] <- median(df_mop[df_mop$time_marker == first_ts[j,1],]$minw)
 }
 first_ts <- as.data.frame(first_ts)
-first_ts <- first_ts %>% mutate(above = 1*(V3 > V4))
+first_ts <- first_ts |> mutate(above = 1*(V3 > V4))
 
 below_med_states <- state_names_mop[which(first_ts$above == 0)]
 above_med_states <- state_names_mop[which(first_ts$above == 1)]
@@ -130,10 +130,10 @@ gplot_values <- cbind(gplot_values,
                       gplot_values[["V4"]] - 1.96*gplot_values[["V5"]],
                       gplot_values[["V4"]] + 1.96*gplot_values[["V5"]])
 
[…]

### 2024-04-24 slides/data.tex
changes

@@
 
-    \frametitle{Data} % Title
+    \frametitle{Data I} % Title
     \framesubtitle{}  % Subtitle
@@
     \begin{wideitemize}
-        \item \textcolor{fgre}{USBLS's LAUS Database} \(\to\) county-level unemployment statistics (1990-2023)
+        \item \textcolor{fgre}{US Bureau of Labor Statistics (BLS)}  
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{LAUS Database} \(\to\) county-level unemployment statistics (1990-2023)
+            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{OEWS Database} \(\to\) national industry-specific occupations and wages statistics (1997-2023) % Occupational Employment and Wage Statistics
+        \end{wideitemize}
+        \item \textcolor{fgre}{Eckert et al. (2020)} \(\to\) county-level industry-specific employment (1975-2016) 
+        \item \textcolor{fgre}{USDOL's Changes in Basic Minimum Wages} (1968-2023)
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{frame}
+
+    \frametitle{Data II} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item \textcolor{fgre}{Horwitz et al. (2020)} \(\to\) Prescription Drug Monitoring Programs (1990-2019)
         \item \textcolor{fgre}{US Census Bureau's Population Estimates Program (PEP)} \(\to\) county-level population and demographics (1969-2020)
-        \item \textcolor{fgre}{USDOL's Changes in Basic Minimum Wages} (1968-2023)
-        \item \textcolor{fgre}{Horwitz et al. (2020)} \(\to\) Prescription Drug Monitoring Programs (1990-2019)
     \end{wideitemize}
@@
         \item Unemployment, PDMPs, Minimum Wage rates and demographics for counties of 50 States + DC, 1998-2019
+        \item County-level industry-weighted wage distribution for counties of 50 States + DC, 2002-2016
     \end{wideitemize}
@@
 
-    \begin{wideitemize}
+    \begin{wideenumerate}
         \item \textcolor{fblu}{Modern System PDMPs}: the electronic database becomes accessible to any authorized user (eg, physician, pharmacist, or member of law enforcement)
         \item \textcolor{fblu}{Must Query PDMPs}: the law mandates the prescriber to check the database before prescribing a listed opioid
-    \end{wideitemize}
+    \end{wideenumerate}
+
+    %\vspace{9pt}
+    %2 always follows 1, but 1 does not always lead to 2
 

### 2024-04-24 slides/estimation.tex
changes

@@
 
+    \frametitle{Kaitz-\(\textcolor{fblu}{\text{\(\rho\)}}\) index} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item Measure of minimum wage bindingness
+    \end{wideitemize}
+
+    \begin{equation*}
+        \text{\textit{kaitz}}_{ct}(\rho) \equiv \log{w^{\text{\textit{min}}}_{ct}} - \log{w^{\rho}_{ct}}
+    \end{equation*}
+    with 
+    \begin{equation*}
+        w^{\rho}_{ct} = \sum_{i\in c} \dfrac{e_i}{E_c}\,w^{\rho}_{it} \quad\text{and}\quad w^{\rho}_{it} = \sum_{o\in i} \dfrac{e_o}{E_i}\,w^{\rho}_{oit}
+    \end{equation*}
+    and
+    \vspace{9pt}
+    \begin{wideitemize}
+        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{w^{\text{\textit{min}}}_{ct}}\): minimum wage in county \(\textcolor{fblu}{c}\) at time \(\textcolor{fblu}{t}\), \(\textcolor{fblu}{w^{\rho}_{ct}}\): wage percentile \(\textcolor{fblu}{\rho}\)
+        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{w^{\rho}_{cit}}\): wage percentile in occupation \(\textcolor{fblu}{o}\) in industry \(\textcolor{fblu}{i}\) at time \(\textcolor{fblu}{t}\)
+        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{e_o}\): share of employment of occupation \(\textcolor{fblu}{o}\) in industry \(\textcolor{fblu}{i}\), \(\textcolor{fblu}{e_i}\): share of county employment in industry \(\textcolor{fblu}{i}\)
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{frame}
+
+    \frametitle{Kaitz-\(\textcolor{fblu}{\text{\(\rho\)}}\) index} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{center}
+        \includegraphics[scale=0.5]{kaitz_percentiles.png}
+    \end{center}
+
+\end{frame}
+
+\begin{frame}
+
     \frametitle{Empirical model} % Title
@@
     \begin{wideitemize}
-        \item Individual-level event study for different minimum wage quantiles: \textcolor{fgre}{Arkhangelsky, Kazuharu and, Zohar (in progress)}
+        \item Individual-level event study for different minimum wage quantiles: \textcolor{fgre}{Arkhangelsky, Yanagimoto and Zohar (in progress)}
     \end{wideitemize}
@@
     \begin{equation*}
-        u_{it} = \alpha_i + \lambda_t + \sum_{k\geq h}\tau_{ik} + \beta X_{it} + \nu_{it}
+        y_{ct} = \alpha_c + \lambda_t + \sum_{k\geq h}\tau_{ck} + \beta X_{ct} + \nu_{ct}
     \end{equation*}
@@
     \begin{wideitemize}
-        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{i}\): county, \(\textcolor{fblu}{t}\): time (year-month), \(\textcolor{fblu}{k = t - z_i}\): time difference from PDMP implementation year in county \(\textcolor{fblu}{i}\), \(\textcolor{fblu}{z_i}\)
-        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{u_{it}}\): unemployment rate, \(\textcolor{fblu}{X_{it}}\): demographic covariates, \(\textcolor{fblu}{\nu_{it}}\): idiosyncratic shock 
+        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{c}\): county, \(\textcolor{fblu}{t}\): time (year-month), \(\textcolor{fblu}{k = t - z_c}\): time difference from PDMP implementation year in county \(\textcolor{fblu}{c}\), \(\textcolor{fblu}{z_c}\)
+        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{y_{ct}}\): labor market outcome, \(\textcolor{fblu}{X_{ct}}\): demographic covariates, \(\textcolor{fblu}{\nu_{ct}}\): idiosyncratic shock
     \end{wideitemize}
@@
 
+    \label{unemp_rate_result}
     \frametitle{Results} % Title
@@
     \begin{center}
-        \includegraphics[scale=0.45]{mop_effects.png}
+        \includegraphics[scale=0.4]{mop010unemprate.png}
     \end{center}
+
+    \hyperlink{perc_comparison_1}{\beamerbutton{Percentiles comparison}}
+    \hyperlink{dist_1}{\beamerbutton{Distributions}}
 
@@
 
+    \label{unemp_result}
     \frametitle{Results} % Title
@@
     \begin{center}
-        \includegraphics[scale=0.45]{pmq_effects.png}
+        \includegraphics[scale=0.4]{mop010unemp.png}
     \end{center}
 
+    \hyperlink{perc_comparison_2}{\beamerbutton{Percentiles comparison}}
+    \hyperlink{dist_2}{\beamerbutton{Distributions}}
+
 \end{frame}
+
+\begin{frame}
+
+    \label{emp_result}
+    \frametitle{Results} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{center}
+        \includegraphics[scale=0.4]{mop010emp.png}
+    \end{center}
+
+    \hyperlink{perc_comparison_3}{\beamerbutton{Percentiles comparison}}
+    \hyperlink{dist_3}{\beamerbutton{Distributions}}
+
+\end{frame}
+
+\begin{frame}
+
+    \label{labforce_result}
+    \frametitle{Results} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{center}
+        \includegraphics[scale=0.4]{mop010labforce.png}
+    \end{center}
+
+    \hyperlink{perc_comparison_4}{\beamerbutton{Percentiles comparison}}
+    \hyperlink{dist_4}{\beamerbutton{Distributions}}
+
+\end{frame}
+
+\begin{comment}
 
@@
 \end{frame}
+
+\end{comment}

### 2024-04-24 data/data_merge.R

@@ -47,13 +47,13 @@ minwage <- read.csv(
   )
 
 minw_bindgness <- read.csv(
-  file.path(getwd(), "county_percentiles.csv"),
+  file.path(getwd(), "sector_wages/county_percentiles.csv"),
   header = TRUE,
   sep = ","
 )
 
 demographics <- read.csv(
-  file.path(getwd(), "county_demographics.csv"),
+  file.path(getwd(), "demographics/county_demographics.csv"),
   header = TRUE,
   sep = ","
   )

### 2024-04-24 data/plots.R

@@ -89,7 +89,7 @@ state_plot <- function(df1,df2,window,policy,outcome){
       aes(x = plot_range, y = ci_right_above), color = "#ff5ca8", linetype = 2,linewidth = 1
     ) +
     ylab(
-      "Effect on unemployment rate"
+      paste0("Effect on ", outcome)
     ) +
     xlab(
       "Period relative to treatment"
@@ -204,7 +204,7 @@ county_plot <- function(df1,df2,window,policy,percentile,outcome){
       aes(x = plot_range, y = ci_right_above), color = "#ff5ca8", linetype = 2,linewidth = 1
     ) +
     ylab(
-      "Effect on unemployment rate"
+      paste0("Effect on ", outcome)
     ) +
     xlab(
       "Period relative to treatment"
@@ -228,4 +228,346 @@ county_plot <- function(df1,df2,window,policy,percentile,outcome){
   return(cty_plot)
   
 }
+
+
+
+comparison_plot <- function(df1,df2,df3,df4,df5,window,policy,outcome){
+  
+  if (policy == "mop"){
+    policy <- "Modern System PDMPs"
+  } else if (policy == "pmq") {
+    policy <- "Must Query PDMPs"
+  } else {
+    print("Policy must be either mop or pmq")
+  }
+  
+  if (outcome == "unemployment rate"){
+    df1 <- df1[,,1]
+    df2 <- df2[,,1]
+    df3 <- df3[,,1]
+    df4 <- df4[,,1]
+    df5 <- df5[,,1]
+  } else if (outcome == "unemployment"){
+    df1 <- df1[,,2]
+    df2 <- df2[,,2]
+    df3 <- df3[,,2]
+    df4 <- df4[,,2]
+    df5 <- df5[,,2]
+  } else if (outcome == "employment"){
+    df1 <- df1[,,3]
+    df2 <- df2[,,3]
+    df3 <- df3[,,3]
+    df4 <- df4[,,3]
+    df5 <- df5[,,3]
+  } else if (outcome == "labor force"){
+    df1 <- df1[,,4]
+    df2 <- df2[,,4]
+    df3 <- df3[,,4]
+    df4 <- df4[,,4]
+    df5 <- df5[,,4]
+  } else {
+    print("Outcome must be either unemployment rate, unemployment, employment or labor force")
+  }
+  
+  plot_range <- -window:window
+  
+  effect_means_1 <- colMeans(df1,na.rm=T) 
+  
+  effect_sd_1 <- sapply(
+    1:(2*window+1),
[…]

### 2024-04-24 main.R

@@ -9,8 +9,16 @@ rm(list=ls())
 set.seed(123)
 
 library(tidyverse)
+library(reshape2)
+
 library(fixest)
+
 library(ggtext)
+library(ggridges)
+
+# Set the path
+file_location <- rstudioapi::getSourceEditorContext()$path
+setwd(dirname(file_location)) # set path to location
 
 # Auxiliary functions
 source(paste(getwd(), "data/utilities.R", sep = "/"))
@@ -18,10 +26,6 @@ source(paste(getwd(), "data/plots.R", sep = "/"))
 
 # One-side length of effects considered
 window = 24 # two years
-
-# Set the path
-file_location <- rstudioapi::getSourceEditorContext()$path
-setwd(dirname(file_location)) # set path to location
 
 data_location <- paste(
   getwd(),
@@ -51,85 +55,44 @@ not_yet_treated_mop <- df[df[["time_marker"]] < df$first_treatment_mop,] # mop
 not_yet_treated_pmq <- df[df[["time_marker"]] < df$first_treatment_pmq,] # pmq
 
 # First stage regressions
-first_stage_mop_unem_rate <- fixest::feols(
-  stats::as.formula(
-    paste0("unem_rate ~ ", covariates, " |", "fips", " + ", "time_marker")
-    ),
-  data = not_yet_treated_mop,
-  combine.quick = FALSE,
-  warn = FALSE,
-  notes = FALSE
-  )
+samples_fs <- c("not_yet_treated_mop","not_yet_treated_pmq")
+outcomes_fs <- c("unem_rate ~ ","unem ~ ", "emp ~ ","lab_force ~ ")
 
-first_stage_mop_unem <- fixest::feols(
-  stats::as.formula(
-    paste0("unem ~ ", covariates, " |", "fips", " + ", "time_marker")
-  ),
-  data = not_yet_treated_mop,
-  combine.quick = FALSE,
-  warn = FALSE,
-  notes = FALSE
-  )
+# TWFE (linear)
+for (i in samples_fs){
+  sample_i = sub("not_yet_treated_", "", i)
+  for (j in outcomes_fs){
+    outcome_j = sub(" ~ ", "", j)
+    assign(
+      paste0("first_stage_",sample_i,"_",outcome_j),
+      fixest::feols(
+        stats::as.formula(
+          paste0(j, covariates, " |", "fips", " + ", "time_marker")
+        ),
+        data = get(i),
[…]

### 2024-05-08 slides/introduction.tex
presentation upload

@@
 
-    \label{Minimum Wage}
-    \frametitle{State minimum wage trends} % Title
-    \framesubtitle{}  % Subtitle
-    \rmfamily % Font
-
-    \begin{wideitemize} 
-        \item States have been increasing their \textcolor{fblu}{minimum wage rates}, with varying intensity
-    \end{wideitemize}
-
-    \begin{center}
-        \includegraphics[scale=0.5]{min_wage_plot_simp.png}
-    \end{center}
-
-    \hyperlink{min_wage_plot_allstates}{\beamerbutton{All states}}
-
-\end{frame}
-
-\begin{frame}
-
+    \label{main_idea}
     \frametitle{Main idea} % Title
@@
     \begin{wideitemize}
-        \item Opioid misuse can lead the victims to negative labor market outcomes
-        \item Minimum wage rates affect these outcomes 
+        \item Opioid misuse can lead the victims to negative Labor Market Outcomes (\textcolor{fblu}{LMOs})
+        \item The impact of substance abuse on LMOs might be mediated by labor market regulations
+        \item \textcolor{fblu}{Minimum wage rates} can affect this impact 
         \vspace{9pt}
         \begin{wideitemize}
-            \item[\textcolor{fblu}{\textbullet}] Negatively, through their impact on \textcolor{fblu}{labor demand for low-skill jobs} 
+            \item[\textcolor{fblu}{\textbullet}] Negatively, through their effect on \textcolor{fblu}{labor demand for low-skill jobs} 
             \item[\textcolor{fblu}{\textbullet}] Positively, if they address the reasons that cause \textcolor{fblu}{addiction} in the first place \(\to\) \textcolor{fblu}{Deaths of despair}
         \end{wideitemize}
-        \item The dominance of each effect depends on how \textcolor{fblu}{binding} minimum wage rates are across labor markets
+        %\item The dominance of each effect depends on how \textcolor{fblu}{binding} minimum wage rates are across labor markets
+        \item How \textcolor{fblu}{binding} minimum wage rates are across labor markets is the relevant measure
     \end{wideitemize}
+
+    \hyperlink{min_wage_plot}{\beamerbutton{Minimum wage trends}}
+    \hyperlink{min_wage_plot_allstates}{\beamerbutton{All states}}
 
@@
     \begin{wideitemize}
-        \item Do the effects of \textcolor{fblu}{Opioid Use Disorders} (\textcolor{fblu}{OUDs}) on labor market outcomes interact with minimum wage policies?
-        \item If so, how have these policies shaped the ongoing Crisis?
-        \item Is the interaction homogeneous across States, or can we expect differences across labor markets?
+        \item Do the effects of \textcolor{fblu}{Opioid Use Disorders} (\textcolor{fblu}{OUDs}) on labor market outcomes vary depending on minimum wage bindingness?
+        %\item If so, how have these policies shaped the ongoing Crisis?
+        %\item Is the interaction homogeneous across States, or can we expect differences across labor markets?
     \end{wideitemize}
@@
 
-    \frametitle{Policy interactions} % Title
+    \frametitle{Prescription policies} % Title
     \framesubtitle{}  % Subtitle
@@
         \item More stringent regulations on opioid access can make the victims to 
+        \vspace{9pt}
         \begin{wideitemize}
@@
         \item This decision affects the MPL of the worker
-        \item In this scenario, minimum wage increases can
+        \item In this scenario, minimum wage rates might
         \vspace{9pt}
+        %\begin{wideitemize}
+        %    \item[\textcolor{fblu}{\textbullet}] Improve labor outcomes if better living conditions tackle the causes of addiction and relapsing (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding})
+        %    \item[\textcolor{fblu}{\textbullet}] Foster addiction, if quitting doesn't improve labor prospects because \(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{binding} 
+        %\end{wideitemize}
         \begin{wideitemize}
-            \item[\textcolor{fblu}{\textbullet}] Improve labor outcomes if better living conditions tackle the causes of addiction and relapsing (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding} after the change)
-            \item[\textcolor{fblu}{\textbullet}] Foster addiction, if quitting doesn't improve employment prospects because \textcolor{fblu}{minimum wage becomes binding} 
+            \item[\textcolor{fblu}{\textbullet}] Increase the rates of unemployment of the workers that don't quit (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{binding})
+            \item[\textcolor{fblu}{\textbullet}] Not have an impact on employment rates for addicted workers (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding})
         \end{wideitemize}
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{comment}
+\begin{frame}
+
+    \frametitle{Prescription policies} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{equation*}
+        \begin{aligned}
+            w^{min}\,\,\text{not binding} & \xrightarrow[\text{passed}]{\text{Prescription law}}
+            \begin{cases}
+                \text{\textcolor{fblu}{Quit}} \\
+                \text{LMOs improve} \\
+                \downarrow \text{Prescription rates} \\
+                \downarrow \text{Mortality} \\
+            \end{cases} \\
+            & \\
+            w^{min}\,\,\text{binding} & \xrightarrow[\text{passed}]{\text{Prescription law}}
+            \begin{cases}
+                \text{\textcolor{fblu}{Black market}} \\
+                \text{LMOs worsen} \\
+                \downarrow \text{Prescription rates} \\
+                \uparrow \text{Mortality} \\
+            \end{cases} \\ 
+        \end{aligned}
+    \end{equation*}
+
+\end{frame}
+\end{comment}
+
+\begin{frame}
+
+    \frametitle{Prescription policies} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item Minimum wage bindingness can affect the decision of the worker to quit opioids
+        \item This impact is \textcolor{fblu}{ambiguous} and depends on the \textcolor{fblu}{cost of quitting} and the \textcolor{fblu}{impact of addiction on productivity}
+        \item Unless minimum wages force the worker to quit strongly enough, we should see a \textcolor{fblu}{negative impact on employment rates} where the minimum wage is \textcolor{fblu}{more binding}
     \end{wideitemize}
@@
     \vspace{9pt}
-    \textbf{Contribution}: shedding light on how the policy responses to the \textcolor{fblu}{Opioid Crisis} have been affected by \textcolor{fblu}{minimum wage changes}
+    \textbf{Contribution}: shedding light on how the labor effects of the \textcolor{fblu}{Opioid Crisis} have been affected by \textcolor{fblu}{minimum wage rates}
     \vspace{9pt}

### 2024-05-08 data/multiple_causes_death.py

@@ -1,0 +1,10 @@
+import glob
+
+txt_files = glob.glob(r'./data/drugs/*.txt')
+#txt_files.pop(txt_files.index('combined_data_three_naics.csv'))
+
+for file in txt_files:
+    with open(file, 'r') as f:
+        # Do something with the file
+        # For example, print its contents
+        print(f.read())

### 2024-05-08 data/plots.R

@@ -7,7 +7,7 @@ KaitzDensitiesPlot <- function(df){
                       "0.75" = "#f29318",
                       "0.90" = "#B79F00")
   
-  densities_plot <- ggplot(kaitz_at_treatment) +
+  densities_plot <- ggplot(df) +
     geom_density(
       aes(x = kaitz_pct10, ..count../sum(..count..), color = "0.10"), linewidth = 1
       ) +
@@ -15,7 +15,7 @@ KaitzDensitiesPlot <- function(df){
       aes(x = kaitz_pct25, ..count../sum(..count..), color = "0.25"), linewidth = 1
       ) + 
     geom_density(
-      aes(x = kaitz_median, ..count../sum(..count..), color = "0.50"), linewidth = 1
+      aes(x = kaitz_pct50, ..count../sum(..count..), color = "0.50"), linewidth = 1
       ) + 
     geom_density(
       aes(x = kaitz_pct75, ..count../sum(..count..), color = "0.75"), linewidth = 1
@@ -110,12 +110,15 @@ IndividualEffectsPlot <- function(df,policy,outcome,percentile){
         ) +
       labs(
         title = paste0("t = ", gsub("V", "", period))
+        ) + 
+      geom_hline(
+        yintercept = 0, linetype = 5, color = "#000000", linewidth = 0.5
         )
     
     plot <- plot + 
       theme_classic() +
       theme(
-        text = element_text(size = 10, family="serif"), 
+        text = element_text(size = 12, family="serif"), 
         plot.title = element_text(hjust = 0.5)
       )
    
@@ -123,18 +126,28 @@ IndividualEffectsPlot <- function(df,policy,outcome,percentile){
     
   }
   
-  lab_left = textGrob("Individual effects",
-                      rot=90,
-                      gp = gpar(fontsize=15, fontfamily="serif")
-                      )
+  setting_labs <- gpar(fontsize=15, fontfamily="serif")
+  setting_title <- gpar(fontsize=20, fontfamily="serif")
   
-  lab_bottom = textGrob("Kaitz-p values",
-                        gp = gpar(fontsize=15, fontfamily="serif")
-                        )
+  lab_left = textGrob("Individual effects", rot=90, gp = setting_labs)
   
-  lab_top = textGrob(paste0(policy," ", outcomes, " ", percentile),
[…]

### 2024-05-08 data/slides_plots.R

@@ -166,7 +166,7 @@ minw_plot <- ggplot(minwg) +
     label = "DC", x = 2018, y = 15, color = "#B79F00", size = 6, family = "serif"
     ) + 
   geom_text(
-    label = "Alabama, Georgia, Idaho...", x = 2016, y = 6.5, color = "#F8766D", size = 6, family = "serif"
+    label = "Federal rate", x = 2016, y = 6.5, color = "#F8766D", size = 6, family = "serif"
     ) +
   xlab(
     ""
@@ -199,7 +199,7 @@ minw_plot_simple <- ggplot(minw_simple) +
     label = "DC", x = 2018, y = 15, color = "#00BA38", size = 5, family = "serif"
     ) + 
   geom_text(
-    label = "Alabama", x = 2016, y = 6.5, color = "#F8766D", size = 5, family = "serif"
+    label = "Federal", x = 2016, y = 6.5, color = "#F8766D", size = 5, family = "serif"
     ) +
   geom_text(
     label = "Minnesota", x = 2019, y = 11, color = "#619CFF", size = 5, family= "serif"

### 2024-05-08 main.R

@@ -43,8 +43,6 @@ df <- read.csv(
   header=TRUE,
   sep=","
   )
-  
-#df <- df[df$year >= 2003 & df$year <= 2016,]
 
 # Prepare the covariates (as formula)
 names_covar <- names(
@@ -103,9 +101,17 @@ effects_pmq_county = CountyEffects(df = df, policy = policies[2], window = windo
 kaitz_at_treatment_mop = KaitzAtTreatment(df = df, policy = policies[1])
 kaitz_at_treatment_pmq = KaitzAtTreatment(df = df, policy = policies[2])
 
+kaitz_at_treatment_mop <- kaitz_at_treatment_mop |>
+  rename(kaitz_pct50 = kaitz_median)
+
+kaitz_at_treatment_pmq <- kaitz_at_treatment_pmq |>
+  rename(kaitz_pct50 = kaitz_median)
+
 # Plot the distributions of the index at treatment (see data/plots.R)
 densities_plot_mop = KaitzDensitiesPlot(kaitz_at_treatment_mop)
+densities_plot_mop
 densities_plot_pmq = KaitzDensitiesPlot(kaitz_at_treatment_pmq)
+densities_plot_pmq 
 
 # Initialize the list of data frames
 joint_dfs <- list()
@@ -127,9 +133,55 @@ for (outcome in outcomes){
   }
 }
 
-
+png(filename = "C:/Users/guill/Documents/GitHub/masters_thesis/slides/mop10_unemp_rate.png",width = 600, height = 539)
 IndividualEffectsPlot(joint_dfs,policies[1],outcomes[1],percentiles[1])
+dev.off() 
+
+png(filename = "C:/Users/guill/Documents/GitHub/masters_thesis/slides/mop10_emp_rate.png",width = 600, height = 539)
 IndividualEffectsPlot(joint_dfs,policies[1],outcomes[2],percentiles[1])
+dev.off() 
+
+png(filename = "C:/Users/guill/Documents/GitHub/masters_thesis/slides/mop10_lab_force_rate.png",width = 600, height = 539)
 IndividualEffectsPlot(joint_dfs,policies[1],outcomes[3],percentiles[1])
+dev.off() 
+
+png(filename = "C:/Users/guill/Documents/GitHub/masters_thesis/slides/mop10_unemp_rate_comp_t1.png",width = 600, height = 539)
+IndividualEffectsCompPlot(joint_dfs,policies[1],outcomes[1],"V1")
+dev.off() 
+
[…]

### 2024-05-31 slides/estimation.tex
rearranging

@@
     \begin{equation*}
-        w^{\rho}_{ct} = \sum_{i\in c} \dfrac{e_i}{E_c}\,w^{\rho}_{it} \quad\text{and}\quad w^{\rho}_{it} = \sum_{o\in i} \dfrac{e_o}{E_i}\,w^{\rho}_{oit}
+        w^{\rho}_{ct} = \sum_{i\in c} \dfrac{e_i}{E_c}\,w^{\rho}_{it} \quad\text{and}\quad w^{\rho}_{it} = \sum_{o\in i} \dfrac{e_o}{E_{it}}\,w^{\rho}_{oit}
     \end{equation*}
@@
     \begin{center}
-        \includegraphics[scale=0.4]{mop010unemprate.png}
+        \includegraphics[scale=0.4]{mop10_unemp_rate.png}
     \end{center}
@@
     \hyperlink{perc_comparison_1}{\beamerbutton{Percentiles comparison}}
-    \hyperlink{dist_1}{\beamerbutton{Distributions}}
+    \hyperlink{ta_1}{\beamerbutton{Time average}}
 
@@
 
-    \label{unemp_result}
+    \label{emp_rate_result}
     \frametitle{Results} % Title
@@
     \begin{center}
-        \includegraphics[scale=0.4]{mop010unemp.png}
+        \includegraphics[scale=0.4]{mop10_emp_rate.png}
     \end{center}
@@
     \hyperlink{perc_comparison_2}{\beamerbutton{Percentiles comparison}}
-    \hyperlink{dist_2}{\beamerbutton{Distributions}}
+    \hyperlink{ta_2}{\beamerbutton{Time average}}
 
@@
 
-    \label{emp_result}
+    \label{lab_force_rate_result}
     \frametitle{Results} % Title
@@
     \begin{center}
-        \includegraphics[scale=0.4]{mop010emp.png}
+        \includegraphics[scale=0.4]{mop10_lab_force_rate.png}
     \end{center}
@@
     \hyperlink{perc_comparison_3}{\beamerbutton{Percentiles comparison}}
-    \hyperlink{dist_3}{\beamerbutton{Distributions}}
-
-\end{frame}
-
-\begin{frame}
-
-    \label{labforce_result}
-    \frametitle{Results} % Title
-    \framesubtitle{}  % Subtitle
-    \rmfamily % Font
-
-    \begin{center}
-        \includegraphics[scale=0.4]{mop010labforce.png}
-    \end{center}
-
-    \hyperlink{perc_comparison_4}{\beamerbutton{Percentiles comparison}}
-    \hyperlink{dist_4}{\beamerbutton{Distributions}}
+    \hyperlink{ta_3}{\beamerbutton{Time average}}
 

### 2024-05-31 slides/model.tex
rearranging

\section{Model}

\begin{transitionframe}

    \rmfamily % Font

    \begin{center}
    {\Huge \textbf{\textcolor{white}{Model}}}
    \end{center}

\end{transitionframe}

\begin{frame}

    \frametitle{Set Up} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    An addicted worker faces
    \vspace{9pt}
    \begin{wideitemize}
        \item A health/performance \textcolor{fblu}{cost of opioid consumption} \(O_{it}\)
        \item This cost depends on \textcolor{fblu}{legal opioids consumption} \(l_{it}\) and \textcolor{fblu}{black market opioids consumption} \(h_{it}\) (say, heroin)
        \item Which opioid is consumed depends on the \textcolor{fblu}{availability of legal opioids} \(z_{it}\) (\(= 1\) if prescription law is in place)
        \item I assume that illegal opioids are more harmful than legal opioids 
        \[h_{it} = l_{it} + \mathcal{C}\]
    \end{wideitemize}

\end{frame}

\begin{frame}

    \label{decision_making}
    \frametitle{Policy and decision making} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item The \textcolor{fblu}{total cost of opioid consumption} is
        \[
        O_{it} = z_{t}\left(1-q_{it}\right)h_{it} + (1-z_{t})l_{it}
        \]
        where \(q_{it} = 1\) \textcolor{fblu}{if the worker quits opioids}
        \item The worker quits if

        \[
        q_{it} = \mathbbm{1}\{\text{Exp. utility of quitting} \geq \text{Disutility of quitting}\}
        \]

    \end{wideitemize}
    \hyperlink{quitting}{\beamerbutton{Quitting}}

\end{frame}

\begin{frame}

    \frametitle{Labor demand} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item The \textcolor{fblu}{representative firm} maximizes profits
        \[
        \pi_{it} = Y_{it} - w_{it}L_{it} \quad \text{s.t} \quad w_t \geq w^{min}_t
        \]
        \vspace{-15pt}
        \item The \textcolor{fblu}{MPL} is given by
        \[
        \dfrac{\partial Y_{it}}{\partial L_{it}} = f(\theta_{it}) - \kappa\cdot\mathbbm{1}\{O_{it}\geq o\}
        \]
        where \(\theta_{it}\) is the \textcolor{fblu}{worker's ability} and \(\kappa\) captures the \textcolor{fblu}{impact of addiction on productivity}
    \end{wideitemize}

\end{frame}

\begin{frame}

    \frametitle{Impact on labor market outcomes} % Title
    \framesubtitle{}  % Subtitle
    \rmfamily % Font

    \begin{wideitemize}
        \item Assume for the moment that \textcolor{fblu}{the worker cannot quit opioids} and that \(l_{it} < o < h_{it}\)
        \item Then if the \textcolor{fblu}{minimum wage is binding} (\(w_{t} = w^{min}_t\)) and \textcolor{fblu}{the law is passed} (\(z_t = 1\))
        \[
        f(\theta_{it}) - \kappa < w^{min}_t \,\Rightarrow\, \text{unemployment}
        \]
        \vspace{-15pt}
        \item But if the \textcolor{fblu}{minimum wage is not binding} (\(w_{t} >> w^{min}_t\)) 
        \[
        f(\theta_{it}) - \kappa = w_t 
        \]
        \vspace{-15pt}
        \item What if the worker can quit opioids?
    \end{wideitemize}
    \hyperlink{quitting}{\beamerbutton{Quitting}} 
\end{frame}

### 2024-05-31 code/county_demographics.py

@@ -1,0 +1,223 @@
+from pyspark.sql import SparkSession
+from pyspark.sql import functions as F
+from pyspark.sql.types import IntegerType
+import os
+
+'''
+returns county_demographics.csv
+
+'''
+
+# Build session 
+os.environ["JAVA_HOME"] = 'C:\Program Files\Java\jre-1.8'
+spark = SparkSession.builder.appName('demographics').getOrCreate()
+
+# Read the data
+demo_data = spark.read.text(r'./data/source/demographics/us_1990_2020_19ages_adjusted.txt')
+
+# Extract columns from value column
+demo_data = demo_data.withColumn(
+    "year",
+    F.substring(F.col("value"), 1, 4)
+    )
+
+demo_data = demo_data.withColumn(
+    "state",
+    F.substring(F.col("value"), 5, 2)
+    )
+
+demo_data = demo_data.withColumn(
+    "state_fip",
+    F.substring(F.col("value"), 7, 2)
+    )
+
+demo_data = demo_data.withColumn(
+    "county_fip",
+    F.substring(F.col("value"), 9, 3)
+    )
+
+demo_data = demo_data.withColumn(
+    "registry",
+    F.substring(F.col("value"), 12, 2)
+    )
+
+demo_data = demo_data.withColumn(
+    "race",
+    F.substring(F.col("value"), 14,1)
+    )
+
+demo_data = demo_data.withColumn(
+    "origin",
+    F.substring(F.col("value"), 15, 1)
+    )
+
+demo_data = demo_data.withColumn(
+    "sex",
+    F.substring(F.col("value"), 16, 1)
+    )
+
+demo_data = demo_data.withColumn(
+    "age",
+    F.substring(F.col("value"), 17, 2)
+    )
+
+demo_data = demo_data.withColumn(
+    "population",
+    F.substring(F.col("value"), 19, 8)
+    )
+
+# Obtain FIPS code (unique identifier for each county)
+demo_data = demo_data.withColumn(
+    "fips", 
+    F.concat(F.col("state_fip"), F.col("county_fip"))
+    )
+
+# Convert data types
+demo_data = demo_data.withColumn(
+    "population", 
+    demo_data.population.cast(IntegerType())
+    )
+
+demo_data = demo_data.withColumn(
+    "year", 
+    demo_data.year.cast(IntegerType())
+    )
+
+demo_data = demo_data.withColumn(
+    "fips", 
+    demo_data.fips.cast(IntegerType())
+    )
+
+# Drop columns
[…]

### 2024-05-31 code/efsy_cbp_merging.py

@@ -1,27 +1,20 @@
-
-'''
-returns combined_data_three_naics.csv
-
-'''
-
 import os, glob
-import pyspark
-import pandas as pd
-
 from pyspark.sql import SparkSession
 from pyspark.sql import functions as F
 from pyspark.sql.types import IntegerType, StructType, StructField, StringType
+
+'''
+returns combined_data_three_naics.csv, combined_data_four_naics_02to05.csv, 
+combined_data_four_naics_06to10.csv, combined_data_four_naics_11to16.csv
+
+'''
 
 # Build session 
 os.environ["JAVA_HOME"] = 'C:\Program Files\Java\jre-1.8'
 spark = SparkSession.builder.appName('sectors').getOrCreate()
 
 # Get a list of all CSV files in the current folder except 'combined_data_three_naics.csv' and 'combined_data_four_naics.csv'
-csv_files = [file.replace('sector_composition\\', '') for file in glob.glob(r'./data/sector_composition/*.csv')]
-csv_files.pop(csv_files.index('combined_data_three_naics.csv'))
-csv_files.pop(csv_files.index('combined_data_four_naics_02to05.csv'))
-csv_files.pop(csv_files.index('combined_data_four_naics_06to10.csv'))
-csv_files.pop(csv_files.index('combined_data_four_naics_11to16.csv'))
+csv_files = [file.replace('./data/source/sector_composition\\', '') for file in glob.glob(r'./data/source/sector_composition/*.csv')]
 
 # Define an empty schema
 column_names = "fipstate|fipscty|naics|emp|year"
@@ -30,12 +23,10 @@ schema = StructType([StructField(c, StringType()) for c in column_names.split("|
 # Create an empty DataFrame
 combined_data = spark.createDataFrame(data=[], schema=schema)
 
-# https://www.datalumos.org/datalumos/project/117464/version/V1/view?flag=follow&path=/datalumos/117464/fcr:versions/V1/Imputed-CBP-Files&type=folder&pageSize=100&sortOrder=(?title)&sortAsc=true
-
 # Iterate over each CSV file and vertically concatenate the data
 for file in csv_files:
     data = spark.read.csv(
-        f'./sector_composition/{file}', 
+        f'./data/source/sector_composition/{file}', 
         header=True
         )
     data = data.filter(
@@ -70,14 +61,14 @@ for file in csv_files:
         )
 
 # Save the combined data to a new CSV file
[…]

### 2024-05-31 code/labor_market_outcomes.py

@@ -1,0 +1,113 @@
+import pandas as pd
+
+'''
+returns unemployment_rate.csv, unemployment.csv, employment.csv, labor_force.csv, 
+employment_population_ratio.csv, labor_force_participation_rate.csv
+
+'''
+# Read the data 
+df = pd.read_csv(r'./data/source/labor_market_outcomes/ladata64County.txt', sep="\t")
+dfips = pd.read_csv(r'./data/source/fips/county_fips_master.csv',encoding = "ISO-8859-1")
+
+# Drop the 'footnote_codes' column from the DataFrame
+df = df.drop('footnote_codes', axis=1)
+
+# Rename the columns of the DataFrame
+df = df.rename(columns={df.columns.values.tolist()[0]: 'series_id', df.columns.values.tolist()[1]: 'year',
+                        df.columns.values.tolist()[2]: 'period', df.columns.values.tolist()[3]: 'value',})
+
+# Extract the state and county FIPS codes from the 'series_id' column
+df['state_fip'] = df['series_id'].str.slice(5, 7)
+df['county_fip'] = df['series_id'].str.slice(7, 10)
+
+# Create a new column 'fips' by concatenating the state and county FIPS codes
+df["fips"] = df["state_fip"] + df["county_fip"]
+
+# Extract the series code, which represents the type of labor market outcome
+df['series'] = df['series_id'].str.slice(18, 20)
+
+# Extract the month from the 'period' column
+df['month'] = df['period'].str.slice(1, 3)
+
+'''
+(meaning of the codes)
+
+measure_code	measure_text
+
+03	unemployment rate
+
+04	unemployment
+
+05	employment
+
+06	labor force
+
+07	employment-population ratio (empty)
+
+08	labor force participation rate (empty)
+
+09	civilian noninstitutional population
+
+https://users.nber.org/~notom/for_tal/LAUS/readme.txt 
+
+MXX=Monthly, M13=Annual
+
+'''
+
+# Select the desired columns from the 'dfips' DataFrame
+dfips = dfips[['fips','county_name','state_abbr','state_name']]
+
+# Convert the 'fips' column to string and pad with zeros to make it 5 characters long
+dfips['fips'] = dfips['fips'].astype(str).str.rjust(5,'0')
+
+# Filter the 'df' DataFrame to include only rows with series code '03' (unemployment rate) and exclude rows with period 'M13' (annual)
[…]

### 2024-05-31 code/minwage.py

@@ -1,0 +1,90 @@
+import requests
+from bs4 import BeautifulSoup
+import pandas as pd
+import os
+
+'''
+returns minwage_raw.csv
+'''
+
+
+# Get the absolute path of the current script
+abspath = os.path.realpath('minwage.py')
+
+# Get the directory name of the current script
+dname = os.path.dirname(abspath)
+
+# Change the current working directory to the script's directory
+os.chdir(dname)
+
+# URL of the webpage to scrape
+URL = "https://www.dol.gov/agencies/whd/state/minimum-wage/history"
+
+# Send a GET request to the URL and get the page content
+page = requests.get(URL)
+
+# Create a BeautifulSoup object to parse the HTML content
+soup = BeautifulSoup(page.content, "html.parser")
+
+# Find all the tables in the HTML content
+tables = soup.find_all('table')
+
+# Create a dictionary to store the tables
+tables_dict = {}
+for i in range(len(tables)):
+    tables_name = f"table{i}"
+    tables_dict[tables_name] = tables[i]
+
+# Find all the table headers (titles) in the HTML content
+titles = soup.find_all('th')
+
+# Extract the text from the table headers
+text_titles = [i.text for i in titles]
+
+# Create a dictionary to store the data from each table
+mwage_dict = {}
+for i in range(len(tables)):
+    # Find all the rows in the current table
+    col_data = tables_dict[f'table{i}'].find_all('tr')
+
+    # Create a dictionary to store the data for each row
+    single_row_data = {}
+    k = 0
+    for row in col_data:
+        # Find all the cells in the current row
+        row_data = row.find_all('td')
+
+        # Extract the text from each cell and store it in the dictionary
+        single_row_data[f'{k}'] = [data.text.strip() for data in row_data]
+        k += 1
+
+    # Remove the first column (index 0) from the row data
+    single_row_data.pop('0')
+
+    # Create a unique name for the current table's data
+    mwage_name = f"mwage{i}"
+
+    # Create a DataFrame from the row data and store it in the dictionary
+    mwage_dict[mwage_name] = pd.DataFrame.from_dict(single_row_data, orient='index')
+
+# Create an empty DataFrame to store the final combined data
+df = pd.DataFrame()
+
+# Concatenate the DataFrames from each table into the final DataFrame
+for i in range(len(tables)):
+    df = pd.concat([df, mwage_dict[f'mwage{i}']], axis=1)
+
[…]

### 2024-05-31 code/minwage_cleaning.py

@@ -1,27 +1,21 @@
+import pandas as pd
 
 '''
 returns minwage_clean_states_ub.csv and minwage_clean_fed_ub.csv
 
 '''
 
-import pandas as pd
-import os
+# See minwage.py
+df = pd.read_csv(r'./data/source/minimum_wage/minwage_raw.csv', sep=',')
 
-abspath = os.path.realpath('minwage_cleaning.py')
-dname = os.path.dirname(abspath)
-os.chdir(dname)
-
-df = pd.read_csv(r'./data/minimum_wage/minwage_raw.csv',sep=',') # see minwage.py
-
-# drop non-relevant obs
-
-df = df.drop(['1968 (a)','1970 (a)','1972','1976 (a)','1979','1980','1981','1988'],axis=1)
+# Drop non-relevant observations
+df = df.drop(['1968 (a)', '1970 (a)', '1972', '1976 (a)', '1979', '1980', '1981', '1988'], axis=1)
 df = df[df['State or otherjurisdiction'] != 'Guam']
 df = df[df['State or otherjurisdiction'] != 'Puerto Rico']
 df = df[df['State or otherjurisdiction'] != 'U.S. Virgin Islands']
-df = df.replace({'\$':''}, regex=True)
 
-# decisions
+# Remove dollar sign from the values
+df = df.replace({'\$': ''}, regex=True)
 
 '''
 > from: https://www.debt.org/jobs/minimum-wage/#:~:text=What%20Happens%20if%20a%20State's,they%20can%20be%20paid%20less.
@@ -36,8 +30,7 @@ df = df.replace({'\$':''}, regex=True)
 
 '''
 
-# apply effective minimum wage
-
+# Apply effective minimum wage
 for i in range(len(df.iloc[:,0])):
     for j in range(len(df.columns)):
         if df.iloc[i,j] == '...':
@@ -52,8 +45,7 @@ for i in range(len(df.iloc[:,0])):
 
 '''
 
-# select higher minimum wage
-
+# Select higher minimum wage
 df = df.replace({'\[c\]':''}, regex=True).replace({'\(d\)':''}, regex=True).replace({'\(e\)':''}, regex=True)
 
 '''
@@ -62,12 +54,10 @@ df = df.replace({'\[c\]':''}, regex=True).replace({'\(d\)':''}, regex=True).repl
[…]

### 2024-05-31 code/wages_sector.py

@@ -7,34 +7,19 @@ https://www.bls.gov/oes/tables.htm
 
 '''
 
-data_wages_1998 = pd.read_excel(r'./data/sector_wages/nat3d_sic_1998_dl.xls', skiprows=31, header=1)
-data_wages_1999 = pd.read_excel(r'./data/sector_wages/nat3d_sic_1999_dl.xls', skiprows=35, header=1)
-data_wages_2000 = pd.read_excel(r'./data/sector_wages/nat3d_sic_2000_dl.xls', skiprows=33, header=1)
-data_wages_2001 = pd.read_excel(r'./data/sector_wages/nat3d_sic_2001_dl.xls', header=0)
-
+# Define the columns to add
 columns_to_add = ['h_pct10', 'h_pct25', 'h_pct75', 'h_pct90']
 
-for data_wages in [data_wages_1998, data_wages_1999, data_wages_2000]:
-    for column in columns_to_add:
-        data_wages[column] = float('nan')
-
-data_wages_1998 = data_wages_1998[['year', 'sic', 'occ_code', 'tot_emp', 'h_median'] + columns_to_add]
-data_wages_1999 = data_wages_1999[['year', 'sic', 'occ_code', 'tot_emp', 'h_median'] + columns_to_add]
-data_wages_2000 = data_wages_2000[['year', 'sic', 'occ_code', 'tot_emp', 'h_median'] + columns_to_add]
-data_wages_2001 = data_wages_2001[['year', 'sic', 'occ_code', 'tot_emp', 'h_median'] + columns_to_add]
-
-data_wages_98to01 = pd.concat([data_wages_1998, data_wages_1999, data_wages_2000, data_wages_2001], axis=0)
-
-# SIC to NAICS
+# Get the list of wages files, excluding the sic_files
 xls_wages = [file.replace('data/sector_wages\\', '') for file in glob.glob(r'data/sector_wages/*.xls*')]
-
 sic_files = ['nat3d_sic_1998_dl.xls', 'nat3d_sic_1999_dl.xls', 'nat3d_sic_2000_dl.xls', 'nat3d_sic_2001_dl.xls', 'field_descriptions.xls']
-
 xls_wages = [x for x in xls_wages if x not in sic_files]
 
+# Create a dictionary to store the data from each wages file
 data_wages = dict()
 years = []
 
+# Read each wages file and store the data in the dictionary
 for file in xls_wages:
     file_name = file.replace(r".xls*", "")
[…]

### 2024-05-31 data/county_demographics.py

@@ -1,271 +1,0 @@
-
-'''
-returns county_demographics.csv
-
-'''
-
-import os
-import pyspark
-import pandas as pd
-
-from pyspark.sql import SparkSession
-from pyspark.sql import functions as F
-from pyspark.sql.types import IntegerType
-
-### build session 
-
-os.environ["JAVA_HOME"] = 'C:\Program Files\Java\jre-1.8'
-spark = SparkSession.builder.appName('demographics').getOrCreate()
-
-### read the data
-
-demo_data = spark.read.text(r'./data/demographics/us_1990_2020_19ages_adjusted.txt')
-
-### obtain variables from txt file (see https://seer.cancer.gov/popdata/popdic.html for reference)
-
-demo_data = demo_data.withColumn(
-    "year",
-    F.substring(F.col("value"), 1, 4)
-)
-
-demo_data = demo_data.withColumn(
-    "state",
-    F.substring(F.col("value"), 5, 2)
-)
-
-demo_data = demo_data.withColumn(
-    "state_fip",
-    F.substring(F.col("value"), 7, 2)
-)
-
-demo_data = demo_data.withColumn(
-    "county_fip",
-    F.substring(F.col("value"), 9, 3)
-)
-
-demo_data = demo_data.withColumn(
-    "registry",
-    F.substring(F.col("value"), 12, 2)
-)
-
-demo_data = demo_data.withColumn(
-    "race",
-    F.substring(F.col("value"), 14,1)
-)
-
-demo_data = demo_data.withColumn(
-    "origin",
-    F.substring(F.col("value"), 15, 1)
-)
-
-demo_data = demo_data.withColumn(
-    "sex",
-    F.substring(F.col("value"), 16, 1)
-)
-
-demo_data = demo_data.withColumn(
-    "age",
-    F.substring(F.col("value"), 17, 2)
-)
-
-demo_data = demo_data.withColumn(
-    "population",
-    F.substring(F.col("value"), 19, 8)
-)
-
-demo_data = demo_data.withColumn(
-    "fips", 
-    F.concat(F.col("state_fip"), F.col("county_fip"))
-)
-
-demo_data = demo_data.withColumn(
-    "population",
-    demo_data.population.cast(IntegerType())
-)
-
-demo_data = demo_data.withColumn(
-    "year",
-    demo_data.year.cast(IntegerType())
-)
-
-demo_data = demo_data.withColumn(
-    "fips",
[…]

### 2024-05-31 data/county_unemp.py

@@ -1,98 +1,0 @@
-
-'''
-returns unemployment_rate.csv, unemployment.csv, employment.csv, labor_force.csv, 
-employment_population_ratio.csv, labor_force_participation_rate.csv
-
-'''
-
-import pandas as pd
-import os
-
-abspath = os.path.realpath('county_unemp.py')
-dname = os.path.dirname(abspath)
-os.chdir(dname)
-
-df = pd.read_csv(r'./data/labor_market_outcomes/ladata64County.txt',sep="\t")
-df = df.drop('footnote_codes',axis=1)
-df = df.rename(columns={df.columns.values.tolist()[0]: 'series_id', df.columns.values.tolist()[1]: 'year',
-                        df.columns.values.tolist()[2]: 'period', df.columns.values.tolist()[3]: 'value',})
-
-df['state_fip'] = df['series_id'].str.slice(5,7)
-df['county_fip'] = df['series_id'].str.slice(7,10)
-df["fips"] = df["state_fip"] + df["county_fip"]
-df['series'] = df['series_id'].str.slice(18,20)
-df['month'] = df['period'].str.slice(1,3)
-
-'''
-measure_code	measure_text
-
-03	unemployment rate
-
-04	unemployment
-
-05	employment
-
-06	labor force
-
-07	employment-population ratio (empty)
-
-08	labor force participation rate (empty)
-
-09	civilian noninstitutional population
-
-https://users.nber.org/~notom/for_tal/LAUS/readme.txt 
-
-MXX=Monthly, M13=Annual
-
-'''
-
-dfips = pd.read_csv(r'./data/fips/county_fips_master.csv',encoding = "ISO-8859-1")
-dfips = dfips[['fips','county_name','state_abbr','state_name']]
-dfips['fips'] = dfips['fips'].astype(str).str.rjust(5,'0')
-
-
-df_unemp_rate = df.drop(df[(df.series != '03') | (df.period == 'M13')].index)
-df_unemp = df.drop(df[(df.series != '04') | (df.period == 'M13')].index)
-df_emp = df.drop(df[(df.series != '05') | (df.period == 'M13')].index)
-df_lab_force = df.drop(df[(df.series != '06') | (df.period == 'M13')].index)
-
-df_dict = {'unemployment_rate':df_unemp_rate,
-           'unemployment':df_unemp,
[…]

### 2024-05-31 data/data_merge.R

@@ -16,12 +16,6 @@ unemp_rate <- read.csv(
   sep = ","
   )
 
-# unemp <- read.csv(
-#   file.path(getwd(), "labor_market_outcomes/unemployment.csv"),
-#   header = TRUE,
-#   sep = ","
-#   )
-
 emp_rate <- read.csv(
   file.path(getwd(), "labor_market_outcomes/employment_rate.csv"),
   header = TRUE,
@@ -50,7 +44,7 @@ minw_bindgness <- read.csv(
   file.path(getwd(), "sector_wages/county_percentiles.csv"),
   header = TRUE,
   sep = ","
-)
+  )
 
 demographics <- read.csv(
   file.path(getwd(), "demographics/county_demographics.csv"),
@@ -63,6 +57,14 @@ drug_deaths <- read.csv(
   header = TRUE,
   sep = ","
   )
+
+naics <- read.csv(
+  file.path(getwd(), "sector_composition/naics_shares_merged.csv"),
+  header = TRUE,
+  sep = ","
+  )
+
+naics <- naics[,!(names(naics) %in% 'emp')]
 
 # PDMPs
 
@@ -127,6 +129,9 @@ horwitz <- horwitz |>
     first_treatment_mop = (mop_year-1960)*12 + mop_month
     ) 
 
+horwitz_mop <- horwitz[c('state','mop_year','mop_month','first_treatment_mop')]
+horwitz_pmq <- horwitz[c('state','pmq_year','pmq_month','first_treatment_pmq')]
+
 # labor market data
 
 ## Rename, time marker, select and numeric values
@@ -140,10 +145,10 @@ clean_lab_market_data <- function(df){
       state = state_name,
       county = county_name,
       rate = value
-    ) |>
+      ) |>
     mutate(
       time_marker = (year-1960)*12 + month
-    )
+      )
   
   df <- df[,colnames(df) != "state_abbr"]
   
@@ -243,8 +248,7 @@ df <- merge(
     "state",
     "time_marker"
     )
-) 
-
+  ) 
 
 df <- merge(
   df,
@@ -279,12 +283,6 @@ df <- merge(
     "fips"
     ),
   all = T
-  )
-
-df <- merge(
-  df,
-  horwitz,
-  by = "state"
   )
 
 df <- merge(
@@ -350,8 +348,257 @@ df <- rbind(df,df_dc)
 #     indicator_mop = 1*(time_marker >= first_treatment_mop)
[…]

### 2024-05-31 data/minwage.py

@@ -1,50 +1,0 @@
-
-'''
-returns minwage_raw.csv
-
-'''
-
-import requests
-from bs4 import BeautifulSoup
-import pandas as pd
-import os
-
-abspath = os.path.realpath('minwage.py')
-dname = os.path.dirname(abspath)
-os.chdir(dname)
-
-URL = "https://www.dol.gov/agencies/whd/state/minimum-wage/history"
-page = requests.get(URL)
-soup = BeautifulSoup(page.content, "html.parser")
-
-tables = soup.find_all('table')
-tables_dict = {}
-for i in range(len(tables)): 
-    tables_name = f"table{i}"
-    tables_dict[tables_name] = tables[i]
-
-titles = soup.find_all('th')
-text_titles = [i.text for i in titles]
-
-mwage_dict = {}
-for i in range(len(tables)):
-    col_data = tables_dict[f'table{i}'].find_all('tr') 
-    single_row_data = {}
-    k = 0
-    for row in col_data:
-        row_data = row.find_all('td')
-        single_row_data[f'{k}'] = [data.text.strip() for data in row_data]
-        k += 1
-    single_row_data.pop('0')
-    mwage_name = f"mwage{i}"
-    mwage_dict[mwage_name] = pd.DataFrame.from_dict(single_row_data, orient='index')
-
-df = pd.DataFrame()
-for i in range(len(tables)):
-    df = pd.concat([df,mwage_dict[f'mwage{i}']], axis=1)
-df = df.set_axis(text_titles, axis=1)
-states = df.iloc[:,0]
-df = df.drop('State or otherjurisdiction', axis=1)
-df = pd.concat([states,df],axis=1)
-
-df.to_csv(r'./data/minimum_wage/minwage_raw.csv',sep=',',index=False)

### 2024-05-31 data/multiple_causes_death.py

@@ -1,10 +1,0 @@
-import glob
-
-txt_files = glob.glob(r'./data/drugs/*.txt')
-#txt_files.pop(txt_files.index('combined_data_three_naics.csv'))
-
-for file in txt_files:
-    with open(file, 'r') as f:
-        # Do something with the file
-        # For example, print its contents
-        print(f.read())

### 2024-05-31 data/plots.R

@@ -1,6 +1,12 @@
 
 KaitzDensitiesPlot <- function(df){
   
+  df_colnames <- df |> colnames()
+  
+  rel_to_treat_name <- df_colnames[grep("relative_to_treat_", df_colnames)]
+
+  df <- df[df[[rel_to_treat_name]] == 0,]
+  
   density_colors <- c("0.10" = "#0098e9",
                       "0.25" = "#ff5ca8",
                       "0.50" = "#5aa800",
@@ -48,14 +54,9 @@ KaitzDensitiesPlot <- function(df){
   
 }
 
-IndividualEffectsPlot <- function(df,policy,outcome,percentile){
+TitleAndColors <- function(policy, outcome, percentile){
   
-  outcomes <- c("unem_rate", "emp_rate","lab_force_rate")
-  policies <- c("mop", "pmq")
   percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
-  periods <- c("V1","V6","V12","V24")
-  plot_colors <- c("#0098e9","#ff5ca8","#5aa800")
-  plot_yranges <- c(ylim(-1.0,1.5),ylim(-1.5,1.0),ylim(-3.5,1.5))
   
   if (percentile %in% percentiles) {
     # Define the Kaitz-p index percentile
@@ -64,56 +65,106 @@ IndividualEffectsPlot <- function(df,policy,outcome,percentile){
     print("Percentile must be 0.10, 0.25, 0.50, 0.75 or 0.90")
   }
   
-  if ((policy %in% policies) & (outcome %in% outcomes)) {
-    # Select data frame and color
-    plot_data <- df[[paste0('effects_',policy,'_county_',outcome)]]
-    plot_color <- plot_colors[outcomes == outcome]
-    plot_yrange <- plot_yranges[outcomes == outcome] 
-  } else {
-    print("Policy must be either mop or pmq, and outcome must be unem_rate, emp_rate or lab_force_rate")
+  if (policy == "mop"){
+    policy_title <- "Modern System PDMPs"
+  } else if (policy == "pmq") {
+    policy_title <- "Must Query PDMPs"
   }
+  
+  if (outcome == "unem_rate"){
+    outcome_title <- "unemployment rate"
+    plot_color <- "#0098e9"
+    plot_yrange <- ylim(-1.0,3.0)
+  } else if (outcome == "emp_rate"){
+    outcome_title <- "employment rate"
[…]

### 2024-05-31 data/sector_shares.py

@@ -1,0 +1,29 @@
+import pandas as pd
+
+data_naics = dict()
+naics_df = ['four_naics_02to05', 'four_naics_06to10', 'four_naics_11to16']
+for naic in naics_df:
+    naic_subset = pd.read_csv(f'./sector_composition/combined_data_{naic}.csv', header=0)
+    naic_subset['fipstate'] = naic_subset['fipstate'].astype(str).str.rjust(2, '0')
+    naic_subset['fipscty'] = naic_subset['fipscty'].astype(str).str.rjust(3, '0')
+    naic_subset['naics'] = naic_subset['naics'].astype(str).str.ljust(6, '0')
+    data_naics[naic] = naic_subset
+
+for naic, df in data_naics.items():
+    data_naics[naic]['fips'] = data_naics[naic]['fipstate'] + data_naics[naic]['fipscty']
+    data_naics[naic] = data_naics[naic].drop(['fipstate', 'fipscty'], axis=1)
+    data_naics[naic]['naics'] = data_naics[naic]['naics'].str[:2]
+    data_naics[naic] = data_naics[naic].groupby(['naics', 'year', 'fips'])['emp'].sum().reset_index()
+
+unique_naics = data_naics['four_naics_02to05']['naics'].unique()
+for naic, df in data_naics.items():
+    for codes in unique_naics:
+        data_naics[naic]['emp_' + codes + '_ratio'] = (data_naics[naic]['emp'] / data_naics[naic].groupby(['year', 'fips'])['emp'].transform('sum')).round(4)
+
+for naic, df in data_naics.items():
+    data_naics[naic] = data_naics[naic][data_naics[naic]['naics'] == '11']  
+    data_naics[naic] = data_naics[naic].drop(['naics'], axis=1)
+
+
+data_naics_merged = pd.concat([data_naics['four_naics_02to05'], data_naics['four_naics_06to10'], data_naics['four_naics_11to16']])
+data_naics_merged.to_csv(r'./sector_composition/naics_shares_merged.csv',sep=',',index=False)

### 2024-05-31 data/utilities.R

@@ -1,131 +1,190 @@
 
-CountyEffects <- function(df,policy,window){
+IndEffects <- function(data,
+                       yname,
+                       iname,
+                       tname,
+                       kname,
+                       aname,
+                       covariates,
+                       k_min = 0,
+                       k_max = 24,
+                       compute_var_me = FALSE,
+                       only_full_horizon = TRUE) {
   
-  # df => all counties, not only those for which
-  # policy was implemented
+  object <- PrepData(data, yname, iname, tname, kname, aname, covariates,
+                     k_min, k_max, compute_var_me, only_full_horizon)
+  
+  object <- ComputeProjection(object)
+  
+  return(object)
+}
 
-  if (policy %in% c("mop", "pmq")) {
-    # Define unexplained parts of interest
-    labor_outcomes <- c(paste0("unem_rate_tilde_", policy), 
-                        paste0("emp_rate_tilde_", policy), 
-                        paste0("lab_force_rate_tilde_", policy)
-                        )
-    # As well as the subsample of treated counties
-    df_treated <- df[!is.na(df[[paste0("first_treatment_", policy)]]),]
-  } else {
-    print("Policy must be either mop or pmq")
+PrepData <- function(data,
+                     yname,
+                     iname,
+                     tname,
+                     kname,
+                     aname,
+                     covariates,
+                     k_min = 0,
+                     k_max = 24,
+                     compute_var_me = FALSE,
+                     only_full_horizon = TRUE) {
+  
+  t_min <- data[[tname]] |> min()
+  not_yet_treated <- data[data[[kname]] < k_min, ]
+  
+  if (nrow(not_yet_treated) == 0 || fixest:::cpp_isConstant(not_yet_treated[[yname]])) {
+    return(dplyr::tibble())
   }
   
-  # Get the unique identifiers for both the sample and subsample
-  fips_names <- unique(df$fips)
-  fips_names_treated <- unique(df_treated$fips)
+  first_stage <- fixest::feols(
+    stats::as.formula(
+      paste0(yname," ~ ", covariates, " |", iname, " + ", tname)
+    ),
+    data = not_yet_treated,
+    combine.quick = FALSE,
[…]

### 2024-05-31 main.R

@@ -23,18 +23,12 @@ setwd(dirname(file_location)) # set path to location
 source(paste(getwd(), "data/utilities.R", sep = "/"))
 source(paste(getwd(), "data/plots.R", sep = "/"))
 
-# Handy strings
-outcomes <- c("unem_rate", "emp_rate","lab_force_rate")
-policies <- c("mop","pmq")
 percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
-
-# One-side length of effects considered
-window = 24 # two years
 
 # Load the data
 data_location <- paste(
   getwd(),
-  "data/joined_data.csv",
+  "data/joined_data_mop.csv",
   sep = "/"
   )
 
@@ -46,142 +40,337 @@ df <- read.csv(
 
 # Prepare the covariates (as formula)
 names_covar <- names(
-  df[,grep("_ratio$", colnames(df))]
+  df[,c(grep("_ratio$", colnames(df)),
+    grep("scaled_model_values", colnames(df)))]
   )
-
-names_covar <- c(names_covar,"scaled_model_values")
 
 covariates <- paste(
   names_covar,
   collapse = " + "
   )
 
-# Select not yet treated counties
-not_yet_treated_mop <- df[df[["time_marker"]] < df$first_treatment_mop,] # mop
-not_yet_treated_pmq <- df[df[["time_marker"]] < df$first_treatment_pmq,] # pmq
+df[['relative_to_treat_mop']] = df[['time_marker']] - df[['first_treatment_mop']]
+names(df)[names(df) == 'kaitz_median'] <- 'kaitz_pct50'
 
-# First stage regressions, TWFE model:
-# For each policy
-for (policy in policies){
-  # retrieve the not yet treated sample
-  not_yet_treated_df = get(paste0("not_yet_treated_",policy)) 
-  # and for each outcome 
-  for (outcome in outcomes){
-    # regress outcome on fixed effects and covariates,
-    first_stage <- fixest::feols(
-      stats::as.formula(
-        paste0(outcome," ~ ", covariates, " |", "fips", " + ", "time_marker")
-        ),
-      data = not_yet_treated_df,
-      combine.quick = FALSE,
-      warn = FALSE,
-      notes = FALSE
-      )
-    # then create a column with the predicted values (\hat{y})
-    column_explained <- paste0(outcome, "_hat_", policy)
[…]

### 2024-06-10 slides/data.tex
presentation upload
before presentation

--- 2024-02-28 slides/data.tex
@@
     \begin{wideitemize}
-        \item \textcolor{fgre}{USBLS's LAUS Database} \(\to\) county-level unemployment statistics (1990-2023)
-        \item \textcolor{fgre}{US Census Bureau's Population Estimates Program (PEP)} \(\to\) county-level population and demographics (1969-2020)
+        \item \textcolor{fgre}{US Bureau of Labor Statistics (BLS)}  
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{LAUS Database} \(\to\) County-level labor market statistics (1990-2023)
+            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{OEWS Database} \(\to\) National industry-specific occupations and wages statistics (1997-2023) % Occupational Employment and Wage Statistics
+        \end{wideitemize}
+        \item \textcolor{fgre}{Eckert et al. (2020)} \(\to\) County-level industry-specific employment (1975-2016) 
         \item \textcolor{fgre}{USDOL's Changes in Basic Minimum Wages} (1968-2023)
+        \item \textcolor{fgre}{US Census Bureau's Population Estimates Program (PEP)} \(\to\) County-level population and demographics (1969-2020)
+        \item \textcolor{fgre}{DEA's ARCOS} \(\to\) Opioid distribution data (2006-2014)
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{frame}
+
+    \frametitle{Data II} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
         \item \textcolor{fgre}{Horwitz et al. (2020)} \(\to\) Prescription Drug Monitoring Programs (1990-2019)
+        \item \textcolor{fgre}{Centers for Disease Control (CDC)} 
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{National Vital Statistics System} \(\to\) Total drug poisoning mortality by county (2003-2021)
+            \item[\textcolor{fgre}{\textbullet}] \textcolor{fgre}{Multiple Cause of Death} \(\to\) Overdose deaths at the state-year level, for different drugs (1999-2020)
+        \end{wideitemize}
+
     \end{wideitemize}
@@
         \item Unemployment, PDMPs, Minimum Wage rates and demographics for counties of 50 States + DC, 1998-2019
+        \item County-level industry-weighted wage distribution and death mortality for counties of 50 States + DC, 2003-2016
     \end{wideitemize}

### 2024-06-10 slides/estimation.tex
before presentation

@@
     \begin{wideitemize}
-        \item Individual-level event study for different minimum wage quantiles: \textcolor{fgre}{Arkhangelsky, Yanagimoto and Zohar (in progress)}
+        \item Individual-level event study for different minimum wage quantiles: \textcolor{fgre}{Arkhangelsky, Yanagimoto and Zohar (2024)}
     \end{wideitemize}
@@
         \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{c}\): county, \(\textcolor{fblu}{t}\): time (year-month), \(\textcolor{fblu}{k = t - z_c}\): time difference from PDMP implementation year in county \(\textcolor{fblu}{c}\), \(\textcolor{fblu}{z_c}\)
-        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{y_{ct}}\): labor market outcome, \(\textcolor{fblu}{X_{ct}}\): demographic covariates, \(\textcolor{fblu}{\nu_{ct}}\): idiosyncratic shock
+        \item[\textcolor{fblu}{\textbullet}] \(\textcolor{fblu}{y_{ct}}\): labor market outcome; \(\textcolor{fblu}{X_{ct}}\): demographic covariates, sector shares and drug mortality; \(\textcolor{fblu}{\nu_{ct}}\): idiosyncratic shock
     \end{wideitemize}
@@
     State-level changes must be exogenous at the county level for \textcolor{fblu}{identification} and no anticipation up to \(\textcolor{fblu}{h}\) periods
+
+\end{frame}
+
+\begin{frame}
+
+    \label{lab_force_rate_result}
+    \frametitle{Results} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{center} % 
+        %\includegraphics[scale=0.4]{mop10_lab_force_rate.png}
+        \includegraphics[scale=0.4]{pmq10_lab_for_rate.png}
+    \end{center}
+
+    \hyperlink{perc_comparison_3}{\beamerbutton{Percentiles comparison}}
+    \hyperlink{ta_3}{\beamerbutton{Time average}}
+    \hyperlink{presc_3}{\beamerbutton{Prescriptions}}
 
@@
 
-    \begin{center}
-        \includegraphics[scale=0.4]{mop10_unemp_rate.png}
+    \begin{center} %pmq10_unemp_rate.png
+        %\includegraphics[scale=0.4]{mop10_unemp_rate.png}
+        \includegraphics[scale=0.4]{pmq10_unemp_rate.png}
     \end{center}
@@
     \hyperlink{ta_1}{\beamerbutton{Time average}}
+    \hyperlink{presc_1}{\beamerbutton{Prescriptions}}
 
@@
 
-    \begin{center}
-        \includegraphics[scale=0.4]{mop10_emp_rate.png}
+    \begin{center} % pmq10_emp_rate.png
+        %\includegraphics[scale=0.4]{mop10_emp_rate.png}
+        \includegraphics[scale=0.4]{pmq10_emp_rate.png}
     \end{center}
@@
     \hyperlink{ta_2}{\beamerbutton{Time average}}
+    \hyperlink{presc_2}{\beamerbutton{Prescriptions}}
 
@@
 
-    \label{lab_force_rate_result}
+    \label{heroin_result}
     \frametitle{Results} % Title
@@
 
-    \begin{center}
-        \includegraphics[scale=0.4]{mop10_lab_force_rate.png}
+    \begin{center} % pmq10_emp_rate.png
+        %\includegraphics[scale=0.4]{mop10_emp_rate.png}
+        \includegraphics[scale=0.4]{pmq10_heroin.png}
     \end{center}
 
-    \hyperlink{perc_comparison_3}{\beamerbutton{Percentiles comparison}}
-    \hyperlink{ta_3}{\beamerbutton{Time average}}
+    \hyperlink{table_heroin}{\beamerbutton{Coefficients}}
 

### 2024-06-10 slides/introduction.tex
before presentation

@@
         \begin{wideitemize}
-            \item[\textcolor{fblu}{\textbullet}] Negatively, through their effect on \textcolor{fblu}{labor demand for low-skill jobs} 
-            \item[\textcolor{fblu}{\textbullet}] Positively, if they address the reasons that cause \textcolor{fblu}{addiction} in the first place \(\to\) \textcolor{fblu}{Deaths of despair}
+            \item[\textcolor{fblu}{\textbullet}] Negatively, through their effect on \textcolor{fblu}{labor demand} for low-skill jobs %and job-searching behavior for low-skill jobs 
+            \item[\textcolor{fblu}{\textbullet}] Positively, if they address the reasons that cause addiction in the first place \(\to\) \textcolor{fblu}{Deaths of despair}
         \end{wideitemize}
@@
     \begin{wideitemize}
-        \item More stringent regulations on opioid access can make the victims to 
+        \item More stringent regulations on opioid access can make the victims of misuse to 
         \vspace{9pt}
@@
         \end{wideitemize}
-        \item This decision affects the MPL of the worker
-        \item In this scenario, minimum wage rates might
-        \vspace{9pt}
-        %\begin{wideitemize}
-        %    \item[\textcolor{fblu}{\textbullet}] Improve labor outcomes if better living conditions tackle the causes of addiction and relapsing (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding})
-        %    \item[\textcolor{fblu}{\textbullet}] Foster addiction, if quitting doesn't improve labor prospects because \(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{binding} 
-        %\end{wideitemize}
-        \begin{wideitemize}
-            \item[\textcolor{fblu}{\textbullet}] Increase the rates of unemployment of the workers that don't quit (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{binding})
-            \item[\textcolor{fblu}{\textbullet}] Not have an impact on employment rates for addicted workers (\(\textcolor{fblu}{w_{min}}\) \textcolor{fblu}{not binding})
-        \end{wideitemize}
+        \item This decision affects the ability of the individual to participate in the labor market and find a job
+        %\item Withdrawing, through its impact on the MPL of the victim, 
+        \item Minimum wage rates might change the incentives of the victims
+        \item If these are \textcolor{fblu}{binding}, and labor demand is affected, the benefits of quitting will be lower
+        %\item If \(\textcolor{fblu}{w^{min}}\) is too binding, workers might be discouraged to quit opioids
+        %\item Workers that face a more binding minimum wage in their labor markets might be discouraged to withdraw from opioids, as they would face a tougher job market search
     \end{wideitemize}
@@
 
-    \begin{equation*}
-        \begin{aligned}
-            w^{min}\,\,\text{not binding} & \xrightarrow[\text{passed}]{\text{Prescription law}}
-            \begin{cases}
-                \text{\textcolor{fblu}{Quit}} \\
-                \text{LMOs improve} \\
-                \downarrow \text{Prescription rates} \\
-                \downarrow \text{Mortality} \\
-            \end{cases} \\
-            & \\
-            w^{min}\,\,\text{binding} & \xrightarrow[\text{passed}]{\text{Prescription law}}
-            \begin{cases}
-                \text{\textcolor{fblu}{Black market}} \\
-                \text{LMOs worsen} \\
-                \downarrow \text{Prescription rates} \\
-                \uparrow \text{Mortality} \\
-            \end{cases} \\ 
-        \end{aligned}
-    \end{equation*}
+    \begin{wideitemize}
+        \item More stringent regulations on opioid access can make the victims to 
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Quit}, which might imply some disutility, or
+            \item[\textcolor{fblu}{\textbullet}] \textcolor{fblu}{Switch to the black market}
+        \end{wideitemize}
+        \item This decision affects the MPL of the worker
+        \item In this scenario, minimum wage rates might
+        \vspace{9pt}
+        %\begin{wideitemize}
+        %    \item[\textcolor{fblu}{\textbullet}] Increase the rates of unemployment of the workers that don't quit (\(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{binding})
+        %    \item[\textcolor{fblu}{\textbullet}] Not have an impact on employment rates for addicted workers (\(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{not binding})
+        %\end{wideitemize}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] Increase the participation rates of the workers that quit (\(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{not binding})
+            \item[\textcolor{fblu}{\textbullet}] Not have an impact on participation and employment rates for addicted workers (\(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{binding})
+        \end{wideitemize}
+    \end{wideitemize}
 
@@
 
+\begin{frame}
+
+    \frametitle{Incentives} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item If addicted individuals face a binding \(\textcolor{fblu}{w^{min}}\), they
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] Might be discouraged to quit opioids (\textit{Cost of quitting} > \textit{Prospective labor earnings})
+            \item[\textcolor{fblu}{\textbullet}] We shouldn't expect an increase in participation rates
+            \item[\textcolor{fblu}{\textbullet}] Illegal opioid consumption should also be higher
+        \end{wideitemize}
+        \item But if \(\textcolor{fblu}{w^{min}}\) is not binding
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] The incentives to quit opioids are higher
+            \item[\textcolor{fblu}{\textbullet}] An increase in participation rates might follow
+            \item[\textcolor{fblu}{\textbullet}] The impact on unemployment rates is less clear
+        \end{wideitemize}
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{frame}
+
+    \frametitle{I find that} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item Policies aimed at stoping opioid misuse
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] Increased participation rates only where \(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{was not binding}
+            \item[\textcolor{fblu}{\textbullet}] The increase was short-lived, but the impact on unemployment rates was persistent
+            \item[\textcolor{fblu}{\textbullet}] Their impact on prescription rates seem to be the main, but not the only, driver of these results
+        \end{wideitemize}
+        \item At the same time
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] They had no labor market consequences where \(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{was binding}
+            \item[\textcolor{fblu}{\textbullet}] Their impact on heroin and synthetic opioid deaths was increasing in \(\textcolor{fblu}{w^{min}}\) bindingness 
+        \end{wideitemize}
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{comment}
 \begin{frame}
@@
 \end{frame}
+\end{comment}
 
@@
     \begin{wideitemize}
-        \item \textcolor{fgre}{Cengiz et al. (2019)} \(\to\) minimum wage changes reduced employment in low-wage jobs of tradeable sectors
-        \item \textcolor{fgre}{Aliprantis, Fee, and Schweitzer (2023)} \(\to\) large and significant negative impact of prescription rates on labor force participation
-        \item \textcolor{fgre}{Alpert, Powell, and Pacula (2018)} \(\to\) positive relation between OxyContin misuse and heroin deaths 
+        %\item Minimum wage changes reduced employment in low-wage jobs of tradeable sectors: \textcolor{fgre}{Cengiz et al. (2019)} 
+        \item Large and significant negative impact of prescriptions and misuse on labor outcomes: {\footnotesize \textcolor{fgre}{Aliprantis, Fee, and Schweitzer (2023), Beheshti (2022), Greenwood, Guner and Kopecky (2022)}} % , Harris et al.
+(2018)
+        \item Prescription drugs regulations can be effective in reducing opioid misuse: {\footnotesize  \textcolor{fgre}{Buchmueller and Carey (2018), Ukert and Polsky (2023)}}
+        \item Positive relation between legal opioid misuse and heroin deaths: {\footnotesize \textcolor{fgre}{Alpert, Powell, and Pacula (2018), Cicero, Ellis, and Surratt (2012), Cicero and Ellis (2015)} }
     \end{wideitemize}

### 2024-06-10 slides/model.tex
before presentation

@@
 
-    An addicted worker faces
+    An addicted individual faces
     \vspace{9pt}
@@
         \item This cost depends on \textcolor{fblu}{legal opioids consumption} \(l_{it}\) and \textcolor{fblu}{black market opioids consumption} \(h_{it}\) (say, heroin)
-        \item Which opioid is consumed depends on the \textcolor{fblu}{availability of legal opioids} \(z_{it}\) (\(= 1\) if prescription law is in place)
-        \item I assume that illegal opioids are more harmful than legal opioids 
-        \[h_{it} = l_{it} + \mathcal{C}\]
+        \item Which opioid is consumed depends on the \textcolor{fblu}{availability of legal opioids} \(z_{t}\) (\(= 1\) if prescription law is in place)
+        \item I assume that opioid misuse impedes the worker's ability to work
+        %\item I assume that illegal opioids are more harmful than legal opioids 
+        %\[h_{it} = l_{it} + \mathcal{C}\]
     \end{wideitemize}
@@
     \label{decision_making}
-    \frametitle{Policy and decision making} % Title
+    \frametitle{Decision making} % Title
     \framesubtitle{}  % Subtitle
@@
     \begin{wideitemize}
-        \item The \textcolor{fblu}{total cost of opioid consumption} is
+        %\item The \textcolor{fblu}{total cost of opioid consumption} is
+        \item The \textcolor{fblu}{decision rule of opioid consumption} is
         \[
@@
         \]
-        where \(q_{it} = 1\) \textcolor{fblu}{if the worker quits opioids}
+        where \(q_{it} = 1\) \textcolor{fblu}{if she quits opioids}
         \item The worker quits if
@@
         \]
-
+        \vspace{-15pt}
+        \item Quitting and looking for a job are equivalent
     \end{wideitemize}
-    \hyperlink{quitting}{\beamerbutton{Quitting}}
+    %\hyperlink{quitting}{\beamerbutton{Quitting}}
 
@@
 
+\begin{comment}
 \begin{frame}
-
+2
     \frametitle{Labor demand} % Title
@@
 \end{frame}
+\end{comment}
 
+\begin{frame}
+
+    \frametitle{Labor market} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item The market has a \textcolor{fblu}{job finding rate} 
+        \[
+            p_{it} = m_{it}/u_{it} 
+        \]
+        where \(m_{it}\) is the \textcolor{fblu}{number of matches} and \(u_{it}\) is the \textcolor{fblu}{number of unemployed workers}
+
+        \item Matches are an increasing function of vacancies \(v_{it}\)
+        \[
+            m_{it} = M(v_{it}, u_{it})
+        \]
+        where \(M\left(\cdot \right) \) is a matching function
+        \item I assume that more vacancies are open where the minimum wage is less binding ({\footnotesize  \textcolor{fgre}{Flinn, 2010}})
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{frame}
+
+    \frametitle{Quitting decision} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item It follows that
+        \[
+            p_{w_{it} > w^{min}_t} \geq p_{w_{it} = w^{min}_t}
+        \]
+        \vspace{-15pt}
+        \item If \(\textcolor{fblu}{w^{min}_t}\) is not binding
+        \[
+            q_{w_{it} > w^{min}_t} = \mathbbm{1}\{p_{w_{it} > w^{min}_t}\cdot u(w_{it}L_{it}) \geq \mathcal{C}^{quit}\}
+        \]
+        \vspace{-15pt}
+        \item Otherwise
+        \[
+            q_{w_{it} = w^{min}_t} = \mathbbm{1}\{p_{w_{it} = w^{min}_t}\cdot u(w^{min}_t L_{it}) \geq \mathcal{C}^{quit}\} 
+        \]
+
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{frame}
+
+    \frametitle{Impact on labor market outcomes} % Title
+    \framesubtitle{}  % Subtitle
+    \rmfamily % Font
+
+    \begin{wideitemize}
+        \item As long as
+        \[
+            p_{w_{it} > w^{min}_t}\cdot u(w_{it}L_{it}) > \mathcal{C}^{quit} > p_{w_{it} = w^{min}_t}\cdot u(w^{min}_t L_{it})
+        \]
+        we should see (after \(z_{t} = 1\))
+        \vspace{9pt}
+        \begin{wideitemize}
+            \item[\textcolor{fblu}{\textbullet}] An increase in participation rates where the minimum wage is less binding (\(O_{it} = 0\))
+            \item[\textcolor{fblu}{\textbullet}] An increase in illegal opioid consumption where the minimum wage is more binding (\(O_{it} = h_{it}\))
+        \end{wideitemize}
+
+    \end{wideitemize}
+
+\end{frame}
+
+\begin{comment}
 \begin{frame}
@@
     \hyperlink{quitting}{\beamerbutton{Quitting}}
+
 \end{frame}
+\end{comment}

### 2024-06-10 code/data_analysis/plots.R

@@ -1,0 +1,677 @@
+KaitzDensitiesPlot <- function(df) {
+
+  # Define the plot title
+  plot_title <- "Distribution of Kaitz-p indices across counties, 
+  for different percentiles at treatment"
+
+  # Get the column names of the dataframe
+  df_colnames <- df |> colnames()
+
+  # Filter the dataframe to include only observations at treatment
+  rel_to_treat_name <- df_colnames[grep("relative_to_treat_", df_colnames)]
+  df <- df[df[[rel_to_treat_name]] == 0,]
+
+  # Define the colors for the density plots
+  density_colors <- c(
+    "0.10" = "#0098e9",
+    "0.25" = "#ff5ca8",
+    "0.50" = "#5aa800",
+    "0.75" = "#f29318",
+    "0.90" = "#B79F00"
+    )
+
+  # Create the densities plot
+  densities_plot <- ggplot2::ggplot(df) +
+    geom_density(
+      aes(x = kaitz_pct10, ..count.. / sum(..count..), color = "0.10"),
+      linewidth = 1
+      ) +
+    geom_density(
+      aes(x = kaitz_pct25, ..count.. / sum(..count..), color = "0.25"),
+      linewidth = 1
+      ) +
+    geom_density(
+      aes(x = kaitz_pct50, ..count.. / sum(..count..), color = "0.50"),
+      linewidth = 1
+      ) +
+    geom_density(
+      aes(x = kaitz_pct75, ..count.. / sum(..count..), color = "0.75"),
+      linewidth = 1
+      ) +
+    geom_density(
+      aes(x = kaitz_pct90, ..count.. / sum(..count..), color = "0.90"),
+      linewidth = 1
+      ) +
+    scale_color_manual(
+      name = element_blank(),
+      values = density_colors
+      )  +
+    ylab(
+      ""
+      ) +
+    xlab(
+      "Kaitz-p index"
+      ) +
+    labs(
+      title = plot_title
+      )
+
+  # Customize the plot appearance
+  densities_plot <- densities_plot +
+    theme_classic() +
+    theme(
+      text = element_text(size = 16, family = "serif"),
+      plot.title = element_markdown(lineheight = 1.1),
+      legend.position = c(0.15, 0.80)
+      )
+
+  return(densities_plot)
+
+}
+
+TitleAndColors <- function(policy, outcome, percentile) {
+
+  # Define the possible percentiles
[…]

### 2024-06-10 code/data_construction/data_merge.R

@@ -1,0 +1,855 @@
+library(tidyverse)
+
+# Clear the workspace
+rm(list = ls())
+options(max.print = 250)
+
+# Set the seed for reproducibility
+set.seed(123)
+
+# Set the file location
+file_location <- dirname(rstudioapi::getSourceEditorContext()$path)
+setwd(gsub("/code/data_construction", "", file_location))
+
+# Load the data
+
+# Unemployment rate data
+unemp_rate_path <- "data/intermediate/labor_market_outcomes/unemployment_rate.csv"
+
+unemp_rate <- read.csv(
+  file.path(getwd(), unemp_rate_path),
+  header = TRUE,
+  sep = ","
+  )
+
+# Employment rate data
+emp_rate_path <- "data/intermediate/labor_market_outcomes/employment_rate.csv"
+
+emp_rate <- read.csv(
+  file.path(getwd(), emp_rate_path),
+  header = TRUE,
+  sep = ","
+  )
+
+# Labor force data
+lab_force_path <- "data/intermediate/labor_market_outcomes/labor_force.csv"
+
+lab_force <- read.csv(
+  file.path(getwd(), lab_force_path),
+  header = TRUE,
+  sep = ","
+  )
+
+# PDMPs data
+horwitz_path <- "data/source/pdmps/horwitz2020.csv"
+
+horwitz <- read.csv(
+  file.path(getwd(), horwitz_path),
+  header = TRUE,
+  sep = ","
+  )
+
+# NAICS shares of the counties data
+naics_path <- "data/intermediate/sector_shares/naics_shares_merged.csv"
+
+naics <- read.csv(
+  file.path(getwd(), naics_path),
+  header = TRUE,
+  sep = ","
+  )
+
+# Minimum wage data
+minwage_path <- "data/intermediate/minimum_wage/minwage_clean_fed_ub.csv"
+
+minwage <- read.csv(
+  file.path(getwd(), minwage_path),
+  header = TRUE,
+  sep = ","
+  )
+
+# Minimum wage bindingness data
+minw_bindgness_path <- "data/intermediate/sector_wages/county_percentiles.csv"
+
+minw_bindgness <- read.csv(
+  file.path(getwd(), minw_bindgness_path),
+  header = TRUE,
+  sep = ","
+  )
+
+minw_bindgness <- minw_bindgness |>
+  dplyr::select(-fipstate, -fipscty)
+
+# Demographics data
+demo_path <- "data/intermediate/demographics/county_demographics.csv"
+
+demographics <- read.csv(
[…]

### 2024-06-10 code/data_construction/labor_market_outcomes.py

@@ -106,6 +106,9 @@ df_emp_rate['lab_force'] = pd.to_numeric(df_emp_rate['lab_force'], errors='coerc
 # Calculate the employment rate by dividing the 'value' column by the 'lab_force' column and multiplying by 100
 df_emp_rate['value'] = round((df_emp_rate['value'] / df_emp_rate['lab_force']) * 100, 1)
 
+# Drop unnecessary columns from each DataFrame
+df_emp_rate = df_emp_rate.drop(['series_id', 'series', 'period', 'state_fip', 'county_fip', 'lab_force'], axis=1)    
+
 # Merge the 'df_emp_rate' DataFrame with the 'dfips' DataFrame based on the 'fips' column
 df_emp_rate = df_emp_rate.merge(dfips, left_on='fips', right_on='fips')
 

### 2024-06-10 code/data_construction/minwage_bindingness.py

@@ -1,24 +1,29 @@
 import pandas as pd
 
+# Load data_naics dictionary
 data_naics = dict()
 naics_df = ['four_naics_02to05', 'four_naics_06to10', 'four_naics_11to16']
 for naic in naics_df:
-    naic_subset = pd.read_csv(f'./data/sector_composition/combined_data_{naic}.csv', header=0)
+    # Read naic_subset data from CSV files
+    naic_subset = pd.read_csv(f'./data/intermediate/sector_composition/combined_data_{naic}.csv', header=0)
     naic_subset['fipstate'] = naic_subset['fipstate'].astype(str).str.rjust(2, '0')
     naic_subset['fipscty'] = naic_subset['fipscty'].astype(str).str.rjust(3, '0')
     naic_subset['naics'] = naic_subset['naics'].astype(str).str.ljust(6, '0')
     data_naics[naic] = naic_subset
 
-wage_percentiles_02to16 = pd.read_csv(r'./data/sector_wages/industry_weighted_percs_02to16.csv', header=0)
+# Load wage_percentiles_02to16 data
+wage_percentiles_02to16 = pd.read_csv(r'./data/intermediate/sector_wages/industry_weighted_percs_02to16.csv', header=0)
 wage_percentiles_02to16['naics'] = wage_percentiles_02to16['naics'].astype(str)
 
+# Merge data_naics with wage_percentiles_02to16
 for naic, df in data_naics.items():
-    data_naics[naic] = df.merge(wage_percentiles_02to16, on=['naics','year'], how='left')
+    data_naics[naic] = data_naics[naic].merge(wage_percentiles_02to16, on=['naics','year'], how='left')
     data_naics[naic]['fips'] = data_naics[naic]['fipstate'] + data_naics[naic]['fipscty']
 
-
+# Create county_percentiles DataFrame
 county_percentiles = pd.DataFrame(index = range(len(data_naics['four_naics_02to05']['fips'].unique())*(2016-2001)+1), columns = ['year', 'fips','h_pct10', 'h_pct25', 'h_median', 'h_pct75', 'h_pct90'])
 
+# Iterate over data_naics to calculate county_percentiles
 h = -1
 for i, file in enumerate(data_naics):
     for j, year in enumerate(data_naics[file]['year'].unique()):
@@ -40,4 +45,10 @@ for i, file in enumerate(data_naics):
             county_percentiles.iloc[h,5] = (county['h_pct75']*county['industry_emp_shares']).sum()   
[…]

### 2024-06-10 code/data_construction/sector_shares.py

@@ -1,0 +1,81 @@
+import pandas as pd
+import os
+
+from pyspark.sql import SparkSession
+from pyspark.sql import functions as F
+from pyspark.sql.types import IntegerType
+
+# Build session 
+os.environ["JAVA_HOME"] = 'C:\Program Files\Java\jre-1.8'
+spark = SparkSession.builder.appName('sectors_shares_merging').getOrCreate()
+
+# Create an empty dictionary to store the data
+data_naics = dict()
+
+# List of naics dataframes
+naics_df = ['four_naics_02to05', 'four_naics_06to10', 'four_naics_11to16']
+
+# Loop through each naics dataframe
+for naic in naics_df:
+    # Read the files
+    naic_subset = pd.read_csv(f'./data/intermediate/sector_composition/combined_data_{naic}.csv', header=0)
+
+    # Preprocess the columns
+    naic_subset['fipstate'] = naic_subset['fipstate'].astype(str).str.rjust(2, '0')
+    naic_subset['fipscty'] = naic_subset['fipscty'].astype(str).str.rjust(3, '0')
+    naic_subset['naics'] = naic_subset['naics'].astype(str).str.ljust(6, '0')
+
+    # Store the naics dataframe in the dictionary
+    data_naics[naic] = naic_subset
+
+# Loop through each naics dataframe in the dictionary
+for naic, df in data_naics.items():
+    # Create a new column 'fips' by combining 'fipstate' and 'fipscty'
+    data_naics[naic]['fips'] = data_naics[naic]['fipstate'] + data_naics[naic]['fipscty']
+    
+    # Drop unnecessary columns
+    data_naics[naic] = data_naics[naic].drop(['fipstate', 'fipscty'], axis=1)
+    
+    # Extract the first two characters from 'naics' column
+    data_naics[naic]['naics'] = data_naics[naic]['naics'].str[:2]
+    
+    # Group the data by 'naics', 'year', and 'fips' and calculate the sum of 'emp'
+    data_naics[naic] = data_naics[naic].groupby(['naics', 'year', 'fips'])['emp'].sum().reset_index()
+
+# Calculate the ratio of 'emp' column for each unique 'naics' code
+
+unique_naics = data_naics['four_naics_02to05']['naics'].unique()
+for naic, df in data_naics.items():
+    for codes in unique_naics:
[…]

### 2024-06-10 code/data_construction/wages_sector.py

@@ -11,7 +11,7 @@ https://www.bls.gov/oes/tables.htm
 columns_to_add = ['h_pct10', 'h_pct25', 'h_pct75', 'h_pct90']
 
 # Get the list of wages files, excluding the sic_files
-xls_wages = [file.replace('data/sector_wages\\', '') for file in glob.glob(r'data/sector_wages/*.xls*')]
+xls_wages = [file.replace('data/source/sector_wages\\', '') for file in glob.glob(r'data/source/sector_wages/*.xls*')]
 sic_files = ['nat3d_sic_1998_dl.xls', 'nat3d_sic_1999_dl.xls', 'nat3d_sic_2000_dl.xls', 'nat3d_sic_2001_dl.xls', 'field_descriptions.xls']
 xls_wages = [x for x in xls_wages if x not in sic_files]
 

### 2024-06-10 data/data_merge.R

@@ -1,604 +1,0 @@
-
-# CLEAN, COMMENT
-
-rm(list=ls()) 
-set.seed(123)
-
-library(tidyverse)
-
-# Set path
-setwd(dirname(rstudioapi::getSourceEditorContext()$path))
-
-# Load the data
-unemp_rate <- read.csv(
-  file.path(getwd(), "labor_market_outcomes/unemployment_rate.csv"), 
-  header = TRUE,
-  sep = ","
-  )
-
-emp_rate <- read.csv(
-  file.path(getwd(), "labor_market_outcomes/employment_rate.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-lab_force <- read.csv(
-  file.path(getwd(), "labor_market_outcomes/labor_force.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-horwitz <- read.csv(
-  file.path(getwd(), "pdmps/horwitz2020.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-minwage <- read.csv(
-  file.path(getwd(), "minimum_wage/minwage_clean_fed_ub.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-minw_bindgness <- read.csv(
-  file.path(getwd(), "sector_wages/county_percentiles.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-demographics <- read.csv(
-  file.path(getwd(), "demographics/county_demographics.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-drug_deaths <- read.csv(
-  file.path(getwd(), "drugs/NCHS_Drug_Poisoning_Mortality_by_County_United_States.csv"), 
-  header = TRUE,
-  sep = ","
-  )
-
-naics <- read.csv(
-  file.path(getwd(), "sector_composition/naics_shares_merged.csv"),
-  header = TRUE,
-  sep = ","
-  )
-
-naics <- naics[,!(names(naics) %in% 'emp')]
-
-# PDMPs
-
-## Divide months and years in PDMPs dataset
-horwitz <- horwitz |> 
-  separate_wider_delim(
-    Enactment,
-    delim = "-",
-    names = c("enact_month", "enact_year")
-    ) |> 
-  separate_wider_delim(
-    Enactment.funding.contingent,
-    delim = "-",
-    names = c("enact_fc_month", "enact_fc_year"),
-    too_few = "align_start"
-    ) |>
-  separate_wider_delim(
-    Enactment.electronic,
-    delim = "-",
[…]

### 2024-06-10 data/plots.R

@@ -499,6 +499,7 @@ IndividualEffectsPlotPrescriptions <- function(df,percentile){
       color = plot_color
     ) +
     plot_yrange +
+    xlim(-0.8,-0.2) +
     ylab(
       "Individual effects"
     ) + 

### 2024-06-10 data/sector_shares.py

@@ -1,29 +1,0 @@
-import pandas as pd
-
-data_naics = dict()
-naics_df = ['four_naics_02to05', 'four_naics_06to10', 'four_naics_11to16']
-for naic in naics_df:
-    naic_subset = pd.read_csv(f'./sector_composition/combined_data_{naic}.csv', header=0)
-    naic_subset['fipstate'] = naic_subset['fipstate'].astype(str).str.rjust(2, '0')
-    naic_subset['fipscty'] = naic_subset['fipscty'].astype(str).str.rjust(3, '0')
-    naic_subset['naics'] = naic_subset['naics'].astype(str).str.ljust(6, '0')
-    data_naics[naic] = naic_subset
-
-for naic, df in data_naics.items():
-    data_naics[naic]['fips'] = data_naics[naic]['fipstate'] + data_naics[naic]['fipscty']
-    data_naics[naic] = data_naics[naic].drop(['fipstate', 'fipscty'], axis=1)
-    data_naics[naic]['naics'] = data_naics[naic]['naics'].str[:2]
-    data_naics[naic] = data_naics[naic].groupby(['naics', 'year', 'fips'])['emp'].sum().reset_index()
-
-unique_naics = data_naics['four_naics_02to05']['naics'].unique()
-for naic, df in data_naics.items():
-    for codes in unique_naics:
-        data_naics[naic]['emp_' + codes + '_ratio'] = (data_naics[naic]['emp'] / data_naics[naic].groupby(['year', 'fips'])['emp'].transform('sum')).round(4)
-
-for naic, df in data_naics.items():
-    data_naics[naic] = data_naics[naic][data_naics[naic]['naics'] == '11']  
-    data_naics[naic] = data_naics[naic].drop(['naics'], axis=1)
-
-
-data_naics_merged = pd.concat([data_naics['four_naics_02to05'], data_naics['four_naics_06to10'], data_naics['four_naics_11to16']])
-data_naics_merged.to_csv(r'./sector_composition/naics_shares_merged.csv',sep=',',index=False)

### 2024-06-10 main.R

@@ -1,13 +1,3 @@
-
-# This code returns the main estimates.
-# 
-# NOTE: 
-# mop: modern system operational
-# pmq: prescriber must query
-
-rm(list=ls()) 
-set.seed(123)
-
 library(tidyverse)
 library(fixest)
 library(ggtext)
@@ -15,20 +5,745 @@ library(binsreg)
 library(gridExtra)
 library(grid)
 
+rm(list = ls())
+set.seed(123)
+options(max.print = 250)
+
+# This code returns the main estimates.
+#
+# NOTE:
+# mop: modern system operational
+# pmq: prescriber must query
+
 # Set the path
 file_location <- rstudioapi::getSourceEditorContext()$path
 setwd(dirname(file_location)) # set path to location
 
 # Auxiliary functions
-source(paste(getwd(), "data/utilities.R", sep = "/"))
-source(paste(getwd(), "data/plots.R", sep = "/"))
+source(paste(getwd(), "code/data_analysis/utilities.R", sep = "/"))
+source(paste(getwd(), "code/data_analysis/plots.R", sep = "/"))
 
+# Percentiles of the wage distribution
 percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
+
+# Load data's location
+data_location_pmq <- paste(
+  getwd(),
+  "data/processed/joined_data_pmq.csv",
+  sep = "/"
+  )
+
+# Load the data for Must Query PDMPs
+df_pmq <- read.csv(
+  data_location_pmq,
+  header = TRUE,
+  sep = ","
+  )
+
+# Include relative-to-treatement time markers and presc per capita
+
+df_pmq <- df_pmq |>
+  dplyr::mutate(
+    relative_to_treat_pmq = time_marker - first_treatment_pmq,
+    dosage_unit_pc = round(dosage_unit / population, 3)
+  )
+
+# Select the covariate columns
+names_covar_pmq <- names(
+  df_pmq[, c(grep("_ratio$", colnames(df_pmq)),
+        grep("model_values", colnames(df_pmq)))]
+  )
+
+# Prepare the covariates (as formula)
+covariates_pmq <- paste(
+  names_covar_pmq,
+  collapse = " + "
+  )
+
+# Compute the coefficient values
+unem_rate_pmq_effects  <- df_pmq |>
+  IndEffects(yname = "unem_rate",
+             iname = "fips",
[…]

### 2024-06-12 slides/introduction.tex
paper ongoing

@@
     \begin{wideitemize}
-        \item Policies aimed at stoping opioid misuse
+        \item Policies aimed at stopping opioid misuse
         \vspace{9pt}
@@
             \item[\textcolor{fblu}{\textbullet}] They had no labor market consequences where \(\textcolor{fblu}{w^{min}}\) \textcolor{fblu}{was binding}
-            \item[\textcolor{fblu}{\textbullet}] Their impact on heroin and synthetic opioid deaths was increasing in \(\textcolor{fblu}{w^{min}}\) bindingness 
+            \item[\textcolor{fblu}{\textbullet}] Their impact on heroin deaths was increasing in \(\textcolor{fblu}{w^{min}}\) bindingness 
         \end{wideitemize}

### 2024-06-12 slides/paper.tex
paper ongoing

\documentclass[12pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsfonts,bbm}
\usepackage{comment}
\usepackage{hyperref}
\usepackage[backend=bibtex,style=authoryear]{biblatex}

\addbibresource{refs.bib}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={paper},
    pdfpagemode=FullScreen,
}

\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}

\setlength\parindent{0pt}
\addtolength{\parskip}{2mm}

\title{The Opioid Crisis: State Regulations and Labor Market Outcomes \\
[2ex] \large CEMFI}
\author{Guillermo Martínez}
\date{\today}

\begin{document}

\maketitle

\begin{abstract}
    \noindent
    While the effects of opioid misuse on the labor market have been studied in detail, little is known about how labor market regulations, particularly the minimum wage, shape this relationship. Do minimum wage rates affect the relationship between opioid misuse and labor market outcomes? To tackle this question, I use the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) as a source of variation. These policies restrict prescription opioids for those considered to be misusing them, forcing individuals to either quit consumption or switch to black market alternatives. The benefits of quitting opioids depend on the labor market opportunities available to those who withdraw. Through their impact on labor demand, binding minimum wages can reduce the incentives to avoid opioid consumption. I find that labor force participation and unemployment rates increased only where the minimum wage was not binding, while a surge in heroin consumption followed where it was. These results highlight the importance of considering labor market regulations when studying the effects of the opioid crisis on the labor market.
\end{abstract}

% Pablo Muñoz: clarificar a que me refiero con binding

\newpage

\section*{Introduction}

For the past two decades, the United States has been facing the worst public health crisis in its recent history.
The so-called opioid epidemic, characterized by a rising number of deaths due to opioid overdose, has caused more than 680,000 fatalities between 1999 and 2022, mostly among non-college-educated Americans.
This situation has prompted economists to study the crisis's effects on the labor market.
However, little is known about how labor market regulations have shaped this relationship.
Minimum wage rates emerge as a natural candidate for studying this relationship.
These policies primarily affect the same subpopulation of low-educated workers and have increased with varying intensity during the same period.
Do minimum wage rates affect the relationship between opioid misuse and labor market outcomes?

To measure this relationship, I use the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) across states.
These programs require authorized users to check a patient's prescription history before prescribing opioids.
As these measures effectively combat opioid misuse, they force those who are addicted at the moment the law is passed to either quit opioid consumption or switch to illegal alternatives.
The incentives to choose one option over the other will depend on the labor market opportunities available to those who withdraw: quitting opioids, although costly and usually painful, has a benefit not only consistent in better health outcomes, but job related too.
Hence, for those who face a minimum wage strong enough to lower labor demand (binding), the benefits of quitting opioids might be lower than for those who face greater job finding opportunities.

To test this hypothesis I regress the county-level labor force participation and unemployment rates on the rollout of PDMPs using an individual effects event study following \textcite{arkhangelsky2024flexible}.
Critical for identification, I assume that the passing of these state-level policies are exogenous from the point of view of the counties, after controlling for the relevant covariates.
The distribution of coefficients is then compared to how binding the minimum wage rates are in the different counties.
After that, I compare the impact of prescription policies on opioid overdose deaths with the same measure of minimum wage bindingness.

The results are striking in several dimensions.
First, I find that policies aimed at stopping opioid misuse increased participation rates only where the minimum wage was not binding.
Secondly, the increase was short-lived (less than two years), but the impact on unemployment rates was persistent.
Moreover, their impact on prescription rates seem to be the main, but not the only, driver of these results.
Finally, prescription policies had no labor market consequences where where the minimum wage was binding, but their impact on heroin deaths was increasing in minimum wage bindingness.

These results shed light on how labor market regulations in general, and minimum wage rates in particular, give rise to heterogeneities in the effects of the opioid crisis on the labor market.
Moreover, they highlight the importance of blah blah.

\section*{Literature review}

\section*{Model}

An addicted worker \(i\) faces at time \(t\):

\begin{itemize}
    \item A health/performance cost of opioid consumption: \(O_{it}\)
    \item This cost depends on legal opioids consumption \(l_{it}\) and black market opioids consumption \(h_{it}\) (say, heroin)
    \item Which opioid is consumed depends on the availability of legal opioids \(z_{it}\) (\(= 1\) if prescription law is in place)
    \item I assume that opioid misuse impedes the worker's ability to work
\end{itemize}

The decision rule of opioid consumption is hence given by:

\[
    O_{it} = z_{t}\left(1-q_{it}\right)h_{it} + (1-z_{t})l_{it}
\]

where \(q_{it} = 1\) if she quits opioids. The worker quits if

\[
    q_{it} = \mathbbm{1}\{\text{Exp. utility of quitting} \geq \text{Disutility of quitting}\}
\]

Quitting and looking for a job are equivalent.The market has a job finding rate

\[
    p_{it} = m_{it}/u_{it} 
\]

where \(m_{it}\) is the number of matches and \(u_{it}\) is the number of unemployed workers. Matches are a non-decreasing function of vacancies \(v_{it}\)

\[
    m_{it} = M(v_{it}, u_{it})
\]

where \(M\left(\cdot \right) \) is a matching function. I assume that more vacancies are open where the minimum wage is less binding (\cite{flinn2010}). It follows that

\[
    p_{w_{it} > w^{min}_t} \geq p_{w_{it} = w^{min}_t}
\]

If \(w^{min}_t\) is not binding

\[
    q_{w_{it} > w^{min}_t} = \mathbbm{1}\{p_{w_{it} > w^{min}_t}\cdot u(w_{it}L_{it}) \geq \mathcal{C}^{quit}\}
\]

Otherwise

\[
    q_{w_{it} = w^{min}_t} = \mathbbm{1}\{p_{w_{it} = w^{min}_t}\cdot u(w^{min}_t L_{it}) \geq \mathcal{C}^{quit}\} 
\]

As long as

\[
    p_{w_{it} > w^{min}_t}\cdot u(w_{it}L_{it}) > \mathcal{C}^{quit} > p_{w_{it} = w^{min}_t}\cdot u(w^{min}_t L_{it})
\]

we should see (after \(z_{t} = 1\)) an increase in participation rates where the minimum wage is less binding (\(O_{it} = 0\)); an increase in illegal opioid consumption where the minimum wage is more binding (\(O_{it} = h_{it}\)).

\section*{Data}

\section*{Results}

\section*{Conclusions}

\printbibliography

\end{document}

### 2024-06-12 code/data_analysis/utilities.R

@@ -10,13 +10,25 @@ IndEffects <- function(data,
                        k_max = 24,
                        compute_var_me = FALSE,
                        only_full_horizon = TRUE) {
-  
-  object <- PrepData(data, yname, iname, tname, kname, aname, covariates,
-                     k_min, k_max, compute_var_me, only_full_horizon)
-  
+
+  object <- PrepData(
+    data,
+    yname,
+    iname,
+    tname,
+    kname,
+    aname,
+    covariates,
+    k_min,
+    k_max,
+    compute_var_me,
+    only_full_horizon
+  )
+
   object <- ComputeProjection(object)
-  
+
   return(object)
+
 }
 
 PrepData <- function(data,
@@ -30,36 +42,36 @@ PrepData <- function(data,
                      k_max = 24,
                      compute_var_me = FALSE,
                      only_full_horizon = TRUE) {
-  
+
   t_min <- data[[tname]] |> min()
   not_yet_treated <- data[data[[kname]] < k_min, ]
-  
+
   if (nrow(not_yet_treated) == 0 || fixest:::cpp_isConstant(not_yet_treated[[yname]])) {
     return(dplyr::tibble())
   }
-  
+
   first_stage <- fixest::feols(
     stats::as.formula(
-      paste0(yname," ~ ", covariates, " |", iname, " + ", tname)
+      paste0(yname, " ~ ", covariates, " |", iname, " + ", tname)
     ),
     data = not_yet_treated,
     combine.quick = FALSE,
     warn = FALSE,
     notes = FALSE)
-  
+
   data[[paste0(yname, "_hat")]] <- stats::predict(first_stage, newdata = data)
-  
+
   data[[paste0(yname, "_tilde")]] <- data[[yname]] -
     data[[paste0(yname, "_hat")]]
-  
+
   df_indcp <- data[!is.na(data[paste0(yname, "_tilde")]), ]
-  
+
   t_min <- df_indcp[[tname]] |> min()
   t_max <- df_indcp[[tname]] |> max()
-  
+
   a_min <- df_indcp[[aname]] |> min()
   a_max <- df_indcp[[aname]] |> max()
-  
+
   info <- list(yname = yname,
                iname = iname,
                tname = tname,

### 2024-06-12 slides/paper.tex
paper update

@@
 [2ex] \large CEMFI}
-\author{Guillermo Martínez}
+\author{Guillermo Martínez Martínez \\ [2ex] Advisor: Tom Zohar}
 \date{\today}
@@
 To tackle this question, I use the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) as a source of variation.
-These policies restrict prescription opioids for those considered to be misusing them, forcing individuals to either quit consumption or switch to black market alternatives.
+These policies restrict prescription opioids for those considered to be misusing them, forcing individuals to either quit consumption or switch to illegal alternatives.
 The benefits of quitting opioids depend on the labor market opportunities available to those who withdraw.
-Through their impact on labor demand, binding minimum wages can reduce the incentives to avoid opioid consumption.
+Through their negative impact on labor demand, high (binding) minimum wages can reduce the incentives to avoid opioid consumption.
 I find that labor force participation and unemployment rates increased only where the minimum wage was not binding, while a surge in heroin consumption followed where it was.
@@
 These programs require authorized users to check a patient's prescription history before prescribing opioids.
-As these measures effectively combat opioid misuse, they force those who are addicted at the moment the law is passed to either quit opioid consumption or switch to illegal alternatives.
-The incentives to choose one option over the other will depend on the labor market opportunities available to those who withdraw: quitting opioids, although costly and usually painful, has a benefit not only consistent in better health outcomes, but job related too.
-Hence, for those who face a minimum wage strong enough to lower labor demand (binding), the benefits of quitting opioids might be lower than for those who face greater job finding opportunities.
+As these measures effectively combat opioid misuse, they force those who are addicted at the time the law is passed to either quit opioid consumption or switch to illegal alternatives.
+The incentives to choose one option over the other depend on the labor market opportunities available to those who withdraw: quitting opioids, although costly and often painful, offers benefits not only in better health outcomes but also in job prospects.
+Hence, for those facing a minimum wage high enough to lower labor demand (binding), the benefits of quitting opioids might be lower than for those with greater job opportunities.
 
-To test this hypothesis I regress the county-level labor force participation and unemployment rates on the rollout of PDMPs using an individual effects event study following \textcite{arkhangelsky2024flexible}.
-Critical for identification, I assume that the passing of these state-level policies are exogenous from the point of view of the counties, after controlling for the relevant covariates.
-The distribution of coefficients is then compared to how binding the minimum wage rates are in the different counties.
-After that, I compare the impact of prescription policies on opioid overdose deaths with the same measure of minimum wage bindingness.
+If this argument holds, we should observe an increase in job-searching behavior where the minimum wage is not binding and an increase in illegal opioid consumption where it is.
+To test these hypotheses, I regress county-level labor force participation and unemployment rates on the rollout of PDMPs using an individual effects event study following \textcite{arkhangelsky2024flexible}.
+Critical for identification, I assume that the passing of these state-level policies is exogenous from the point of view of the counties, after controlling for the number of overdose deaths, among other variables.
+I then compare the distribution of coefficients to how binding the minimum wage rates are in the different counties.
+In a second stage, I compare the impact of prescription policies on overdose deaths for different opioids using the same measure of minimum wage bindingness.
 
@@
 \section*{Literature review}
+
+This paper is located at the intersection of three strands of literature.
+
+Large and significant negative impact of prescriptions and misuse on labor outcomes: Aliprantis, Fee, and Schweitzer (2023), Beheshti (2022), Greenwood, Guner and Kopecky (2022) % , Harris et al.
+(2018)
+Prescription drugs regulations can be effective in reducing opioid misuse: Buchmueller and Carey (2018), Ukert and Polsky (2023)
+Positive relation between legal opioid misuse and heroin deaths: Alpert, Powell, and Pacula (2018), Cicero, Ellis, and Surratt (2012), Cicero and Ellis (2015)
+
+\textbf{Contribution}: shedding light on how the labor effects of the Opioid Crisis have been affected by minimum wage rates
+
+Interest in how labor market regulations interact with Substance Abuse Disorders
+
+\begin{comment}
+
+#### Negative Impact of Opioid Misuse on Labor Outcomes
+
+Extensive research has documented the large and significant negative impact of opioid prescriptions and misuse on labor market outcomes.
+Aliprantis, Fee, and Schweitzer (2023) highlight how opioid addiction leads to substantial declines in employment rates and labor force participation.
+Similarly, Beheshti (2022) finds that opioid misuse is linked to reduced work hours and earnings, particularly among low-educated workers.
+Greenwood, Guner, and Kopecky (2022) show that regions with high opioid prescription rates experience notable reductions in economic productivity and workforce stability.
+Harris et al.
+(2018) further corroborate these findings, demonstrating the broad economic costs associated with opioid addiction, including increased unemployment and decreased job-seeking behavior.
+
+#### Effectiveness of Prescription Drug Regulations
+
+Research also indicates that prescription drug regulations can effectively reduce opioid misuse.
+Buchmueller and Carey (2018) provide evidence that policies limiting opioid prescriptions lead to substantial declines in opioid abuse and related health issues.
+Ukert and Polsky (2023) build on this work by showing that state-level interventions, such as Prescription Drug Monitoring Programs (PDMPs), significantly decrease the availability of prescription opioids and reduce rates of opioid misuse.
+These studies collectively suggest that regulatory measures can be an essential tool in combating the opioid crisis.
+
+#### Relationship Between Legal Opioid Misuse and Heroin Deaths
+
+The transition from legal opioid misuse to heroin use has been well-documented, with several studies illustrating a positive relationship between the two.
+Alpert, Powell, and Pacula (2018) find that restrictions on prescription opioids often lead individuals to seek out heroin as a cheaper and more accessible alternative, resulting in increased heroin overdose deaths.
+Cicero, Ellis, and Surratt (2012) also identify this substitution effect, noting a significant rise in heroin use following efforts to curtail prescription opioid abuse.
+Further research by Cicero and Ellis (2015) supports these findings, emphasizing the need for comprehensive approaches that address both prescription and illicit opioid markets to effectively reduce overall opioid-related mortality.
+
+Feel free to adjust the details or add more context based on the specific focus and needs of your paper!
+\end{comment}
 
@@
 
+US Bureau of Labor Statistics (BLS):
+LAUS Database: County-level labor market statistics (1990-2023) % Local Area Unemployment Statistics
+OEWS Database: National industry-specific occupations and wages statistics (1997-2023) % Occupational Employment and Wage Statistics
+
+Eckert et al. (2020): County-level industry-specific employment (1975-2016)
+
+USDOL's Changes in Basic Minimum Wages (1968-2023)
+
+US Census Bureau's Population Estimates Program (PEP) : County-level population and demographics (1969-2020)
+
+DEA's ARCOS : Opioid distribution data (2006-2014)
+
+Horwitz et al. (2020) : Prescription Drug Monitoring Programs (1990-2019)
+
+Centers for Disease Control (CDC):
+National Vital Statistics System: Total drug poisoning mortality by county (2003-2021)
+Multiple Cause of Death: Overdose deaths at the state-year level, for different drugs (1999-2020)
+
+Current working sample:
+LMOs, PDMPs, Minimum Wage rates and demographics for counties of 50 States + DC, 1998-2019
+County-level industry-weighted wage distribution and death mortality for counties of 50 States + DC, 2003-2016
+
+Prescription Drug Monitoring Programs (PDMPs): electronic database that tracks controlled substance prescriptions
+Modern System PDMPs: the electronic database becomes accessible to any authorized user (eg, physician, pharmacist, or member of law enforcement)
+Must Query PDMPs: the law mandates the prescriber to check the database before prescribing a listed opioid
+
 \section*{Results}

### 2024-06-16 slides/paper.tex
paper

--- 2024-06-12 slides/paper.tex
@@
 \usepackage{graphicx}
-\usepackage{amsmath,amssymb,amsfonts,bbm}
-\usepackage{comment}
+\usepackage{amsmath,amssymb,amsfonts,bbm} % For math
+\usepackage{comment} % Comment blocks
 \usepackage{hyperref}
+\usepackage{setspace}
+\usepackage{booktabs} % For tables
+\usepackage{longtable}
 \usepackage[backend=bibtex,style=authoryear]{biblatex}
+
+\renewcommand{\baselinestretch}{1.25}
 
@@
     urlcolor=cyan,
+    citecolor=cyan,
     pdftitle={paper},
@@
 [2ex] \large CEMFI}
-\author{Guillermo Martínez}
+\author{Guillermo Martínez \\ [2ex] Advisor: Tom Zohar}
 \date{\today}
@@
 For the past two decades, the United States has been facing the worst public health crisis in its recent history.
-The so-called opioid epidemic, characterized by a rising number of deaths due to opioid overdose, has caused more than 680,000 fatalities between 1999 and 2022, mostly among non-college-educated Americans.
-This situation has prompted economists to study the crisis's effects on the labor market.
+The so-called opioid epidemic, characterized by a rising number of deaths due to opioid overdose, has caused more than 680,000 fatalities between 1999 and 2022 (\cite{CDCMCD9920}, \cite{CDCMCD1822}), mostly among non-college-educated Americans.
+The sheer magnitude of this crisis has left a significant footprint on the labor market, prompting economists to study the relationship between opioid misuse and labor market outcomes.
 However, little is known about how labor market regulations have shaped this relationship.
-Minimum wage rates emerge as a natural candidate for studying this relationship.
-These policies primarily affect the same subpopulation of low-educated workers and have increased with varying intensity during the same period.
-Do minimum wage rates affect the relationship between opioid misuse and labor market outcomes?
+Minimum wage rates emerge as a natural candidate for studying this dynamic since they affect the same subpopulation of low-educated workers and have increased with varying intensity during the same period.
 
+%This situation has prompted economists to study the crisis's effects on the labor market.
+%Do minimum wage rates affect the relationship between opioid misuse and labor market outcomes?
+
+\begin{comment}
+    MORITZ:
+    What do you mean by recent? % common in the literature
+    Why so-called? 
+    Is it important that you mention non-college educated people NOW?
+    I think you need a better connection from opioid deaths to the study of labor market regulations
+    I am not sure you have to talk about labor market regulation in general terms at all, you can immediately go into minimum wage rates that is very interesting in its own right % I still need regulations to give a better intuition
+    I think I know the connection between low education levels and minimum wage rates (based on bindingness) but the connection here could be clearer
+    I have questions about the importance of access to opioids in driving the epidemic
+\end{comment}
+
+This paper argues that minimum wage rates affect the relationship between opioid misuse and labor market outcomes.
 To measure this relationship, I use the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) across states.
 These programs require authorized users to check a patient's prescription history before prescribing opioids.
-As these measures effectively combat opioid misuse, they force those who are addicted at the moment the law is passed to either quit opioid consumption or switch to illegal alternatives.
-The incentives to choose one option over the other will depend on the labor market opportunities available to those who withdraw: quitting opioids, although costly and usually painful, has a benefit not only consistent in better health outcomes, but job related too.
-Hence, for those who face a minimum wage strong enough to lower labor demand (binding), the benefits of quitting opioids might be lower than for those who face greater job finding opportunities.
+By effectively combating opioid misuse, these measures force those addicted to either quit opioid consumption or switch to illegal alternatives.
+The incentives to choose one option over the other depend on the labor market opportunities available to those who withdraw. 
+Quitting opioids, although costly and often painful, offers benefits not only in better health outcomes but also in job prospects. 
+However, for those facing a minimum wage high enough to lower labor demand (binding), the benefits of quitting opioids might be lower than for those with greater job opportunities.
+My estimations support this intuition, showing increases in labor force participation and unemployment rates as a result of prescription policies only where the minimum wage was not binding, while a surge in heroin consumption occurred where it was.
 
-To test this hypothesis I regress the county-level labor force participation and unemployment rates on the rollout of PDMPs using an individual effects event study following \textcite{arkhangelsky2024flexible}.
-Critical for identification, I assume that the passing of these state-level policies are exogenous from the point of view of the counties, after controlling for the relevant covariates.
-The distribution of coefficients is then compared to how binding the minimum wage rates are in the different counties.
-After that, I compare the impact of prescription policies on opioid overdose deaths with the same measure of minimum wage bindingness.
+%at the time the law is passed
+
+% Condensate the whole background part into a single (two) paragraph
+These findings contribute to the growing body of literature studying the economic effects of the opioid crisis.
+Among this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation.
+This perspective has led economists to focus on supply-side interventions to explore the causal relationship between opioid consumption and labor market outcomes.
+While well-intentioned, many of these policies have had unintended consequences, such as the rise in heroin consumption following the reformulation of OxyContin.
+Prescription Drug Monitoring Programs (PDMPs), however, have been less scrutinized, with recent research indicating that, in their more stringent form, they can effectively reduce prescription opioid misuse.
+
+Yet, how heterogenous these effects are across labor markets remains an open question.
+Different incentives can lead victims of opioid misuse to take different decisions regarding both their health and their economic prospects. 
+Labor market regulations, by altering the labor outcomes of workers, influence this decision-making process by affecting the incentives to quit opioids -- a family of drugs that can severely impair the ability to work.
+To understand how these regulations have shaped the ongoing crisis, I focus on minimum wage rates, which have increased in all states at different speeds during the same period as the opioid crisis, thereby affecting the labor market opportunities of low-educated workers.
+Critical to our argument is not the rates themselves, but how binding they are across different labor markets: the more binding the minimum wage, the less incentives workers have to quit opioids, as they face a tougher job market search.
+
+I use several data sources to obtain a distribution of coefficients of the impact of opioid misuse on labor market outcomes across different minimum wage bindingness levels.
+The US Bureau of Labor Statistics (BLS) LAUS database provides county-level labor market statistics, while the US Census Bureau's Population Estimates Program (PEP) offers county-level population and demographic data.
+From \textcite{Horwitz2021} I obtain data on Prescription Drug Monitoring Programs (PDMPs). 
+The BLS's Occupational Employment and Wage Statistics (OEWS) database provides national industry-specific occupations and wages statistics, \textcite{Eckert2020} offers county-level industry-specific employment data, and the US Department of Labor's Changes in Basic Minimum Wages provides information on minimum wage rates; all of these data are relevant for constructing a measure of minimum wage bindingness.
+The Centers for Disease Control (CDC) National Vital Statistics System offers data on total drug poisoning mortality by county, while the CDC's Multiple Cause of Death database provides overdose deaths at the state-year level for different drugs.
+Finally, the Drug Enforcement Administration's Automation of Reports and Consolidated Orders System (ARCOS) offers opioid distribution data.
+The joint panel covers the United states for the period 2003-2016.
 
 The results are striking in several dimensions.
+After constructing a measure of minimum wage bindingness known as the Kaitz-\(\rho\) index, I find that the impact of prescription policies on labor market outcomes is indeed heterogeneous across different levels of minimum wage bindingness.
 First, I find that policies aimed at stopping opioid misuse increased participation rates only where the minimum wage was not binding.
@@
 
+\textbf{Contribution}: shedding light on how the labor effects of the Opioid Crisis have been affected by minimum wage rates
+
+Interest in how labor market regulations interact with Substance Abuse Disorders
+
 These results shed light on how labor market regulations in general, and minimum wage rates in particular, give rise to heterogeneities in the effects of the opioid crisis on the labor market.
@@
 
-\section*{Literature review}
+OUTLINE of the paper (the paper is structured as follows...)
 
-\section*{Model}
+\section*{Background: Opioid Crisis and regulatory changes}
 
-An addicted worker \(i\) faces at time \(t\):
+In this section, I provide a summary of how the opioid crisis started and what effects it had on the labor market in general. 
+Then, I discuss some of the policies implemented to curb the opioid misuse, and how some of them resulted ineffective or simply backfired. 
+Finally, I describe how labor market incentives, and in particular minimum wage rates, can explain some of these policy failures and the heterogeneity of the labor market outcomes.
 
-\begin{itemize}
-    \item A health/performance cost of opioid consumption: \(O_{it}\)
-    \item This cost depends on legal opioids consumption \(l_{it}\) and black market opioids consumption \(h_{it}\) (say, heroin)
-    \item Which opioid is consumed depends on the availability of legal opioids \(z_{it}\) (\(= 1\) if prescription law is in place)
-    \item I assume that opioid misuse impedes the worker's ability to work
-\end{itemize}
+\subsection*{The origins of the opioid epidemic and its impact on the labor market}
+
+% CHECKED
+It is difficult to overstate the importance of the opioid crisis in the United States. 
+In 2022 alone, 110,000 people died from drug abuse, three-quarters of them due to opioids.
+For context, the Vietnam War -- one of the most traumatic events in recent American history -- killed 58.000 US soldiers in total.
+Given the severity of the crisis, it is not surprising that a large body of literature has emerged to study its origins and consequences.
+
+% CHECKED
+Although demand-side explanations of the crisis exist, with \textcite{Deaton2017} and \textcite{Deaton2015} being some of the most influential, a supply-side view of the epidemic's origins is now widely accepted (\cite{Eichmeyer2022}, \cite{Ruhm2018}).
+This shift is partly due to mixed evidence on the causal impact of declining economic conditions on opioid consumption (\cite{Currie2018}).
+Moreover, the efforts by pharmaceutical companies and healthcare providers to increase opioid prescriptions since the late 1990s have been well documented (\cite{Alpert2019}, \cite{Kolodny2015}, \cite{VanZee2008}).
+Purdue Pharma's aggressive marketing campaign of OxyContin, a powerful opioid painkiller, is often cited as the starting point of the crisis (\cite{Chow2019}).
+
+% CHECKED
+Economists quickly began studying the effects of the opioid crisis on the labor market.
+In a seminal paper, \textcite{Krueger2017} found that labor force participation rates were falling on greater numbers in areas with higher opioid prescription rates, an analysis later refined by \textcite{Aliprantis2023}.
+This relationship has also been emphasized from a macroeconomic perspective (\cite{Greenwood2022a}, \cite{Greenwood2022b}).
+Many researchers have used exogenous variations in prescription rates to further establish the causal relationship between opioid consumption and worsened labor market outcomes (\cite{Beheshti2023}, \cite{Powell2022}, \cite{Harris2020}).
+These papers typically rely on federal- or state-level regulatory changes implemented in response to the ongoing public health crisis.
+However, none have delved into how labor market regulations can create heterogeneity in these effects.
+
+\subsection*{Policy efforts and unintended consequences}
+
+% CHECKED
+Policy efforts to curb the opioid crisis have been numerous and varied, although not all of them have been equally effective.
+A much-discussed example is the reformulation of OxyContin in 2010, which made the dosage units more difficult to crush or dissolve, thus reducing the drug's abuse potential through non-oral routes.
+This change led to a surge in heroin consumption as users switched to a more readily accessible alternative for misuse (\cite{Alpert2018}, \cite{Evans2019}, \cite{Cicero2012}, \cite{Butler2013}).
+Additionally, the switch to heroin seemed to result in a greater prevalence of blood-borne diseases in the most affected areas (\cite{Beheshti2019}).
+Lastly, there is anectodic evidence that those who did not choose a black market alternative still found ways to make the drug injectable (\cite{Cicero2015}).
+
+% CHECKED
+Abuse-deterrent formulations were not the only efforts that backfired. 
+For instance, \textcite{Doleac2022} find that increased access to Naloxone (a drug used to reverse opioid overdoses) led to a rise in opioid-related emergency room visits and opioid-related crime. 
+The impact of these policies on opioid-related mortality seems to be restricted to early-adopter states (\cite{Rees2019}).
+Medical marijuana laws (\cite{Powell2018}), Medicare Part D (\cite{Powell2020}), substance abuse treatment availability (\cite{Swensen2015}), and physicians' education (\cite{Schnell2018}) are among the many policies and factors that economists have studied in relation to the opioid epidemic, contributing to a vast body of literature.
+
+% CHECKED
+Fewer concerns have been raised about the unintented consequences of Prescription Drug Monitoring Programs (PDMPs).
+According to the \textcite{CDC2021}, a PDMP is ``\textit{a database that collect information about dispensed prescription drugs.}''.
+The collected data can only be accessed by authorized users (e.g., physicians, pharmacists, or law enforcement members) and is used to identify and prevent drug misuse.
+Although early studies questioned the effectiveness of these programs in reducing opioid misuse (\cite{Paulozzi2011}, \cite{Maughan2015}, \cite{Meara2016}) and the effectiveness of supply-side interventions in drug-related policies is mixed (\cite{Pollack2014}, \cite{Dobkin2019}, \cite{Dobkin2014}), more recent work has found that in states where prescribers are mandated to check the database before prescribing a listed opioid, the number of opioid misuse cases related to prescription opioids decreased (\cite{Buchmueller2018}, \cite{Sacks2021}, \cite{Grecu2019}).
+These more stringent PDMPs, known as Must Query PDMPs, therefore offer a unique opportunity to study the relationship between opioid misuse and labor market outcomes.
+Their staggered rollout across states, combined with the variation at the county level in how binding minimum wage rates are, allows for a quasi-experimental design that can shed light on how labor market regulations shape the effects of the opioid crisis.
+
+\subsection*{Bringing minimum wage rates in}
+
+% CHECKED
+During the same period as the Opioid Crisis, minimum wage rates have been increasing in all states, albeit at different speeds.
+While some states such as California and New York have raised their minimum wage to over \$15 per hour, others have kept it at the federal minimum of \$7.25.
+This spatial and temporal variability makes minimum wage rates a natural candidate to study the mediation of labor market regulations on the relationship between opioid misuse and labor market outcomes.
+However, its relevance hinges on whether these minimum wage rates affect labor demand.
+
+% CHECKED
+Although the significance of the impact of minimum wages on labor demand has been questioned since \textcite{Card1994} published their seminal paper, recent research shows that minimum wage increases can have negative effects on employment in certain contexts.
+For instance, \textcite{Cengiz2019} found that minimum wage increases in the United States led to a decrease in employment in tradeable sectors.
+This finding is consistent with \textcite{Harasztosi2019}, who observed that the impact on employment is larger in industries where passing costs to consumers is more difficult, using data from Hungary.
+\textcite{Clemens2019} showed that minimum wage increases in the United States led to a decrease in employment in states bound by the federal minimum wage increase during the Great Recession.
+Furthermore, \textcite{Clemens2021} showed that large increases in minimum wages have reduced employment rates among individuals with low levels of education, in accordance with \textcite{Neumark2004},although the effect is negligible for small increases.
+In a recent paper, \textcite{Engbom2022} pointed out that the effect of these increases on employment can be muted by reallocations of workers across firms.
+Crucially, they argue that lower-productivity firms cut vacancy creation and hiring, which in our context could lead to lower job finding rates where the minimum wage is more binding.
+
+% CHECKED
+If relatively high minimum wages reduce labor demand for low-skilled workers, or reduce it in areas where lower-productivity firms are more prevalent, the prospective labor earnings of non-college-educated workers should be lower in these areas.
+This could, in turn, reduce the incentives for opioid addicts to withdraw as prescription drug regulations are passed. 
+The difficult and often painful process of withdrawal might be less appealing compared to opting for illegal alternatives if job opportunities are scarce.
+This intuition, which I formalize in the next section, suggests that the effects of prescription drug regulations on labor market outcomes should be more pronounced in areas where the minimum wage is not binding.
+
+\section*{A model of job searching and opioid consumption}
+
+The following model formalizes the intuition outlined above. 
+An addicted worker \(i\) faces a health/performance cost of opioid consumption \(O_{it}\) at time \(t\).
+This cost depends on whether she consumes prescription/legal opioids \(l_{it}\) or illegal opioids \(h_{it}\) (e.g., heroin).
+The type of opioid consumed depends on the availability of legal opioids \(z_{it}\), which equals 1 if an effective prescription law is in place.
+I assume that opioid misuse impedes the individual's ability to work.
 
@@
 
-\[
+\begin{equation}
     O_{it} = z_{t}\left(1-q_{it}\right)h_{it} + (1-z_{t})l_{it}
-\]
+\end{equation}
 
-where \(q_{it} = 1\) if she quits opioids. The worker quits if
+where \(q_{it} = 1\) if she quits opioids. 
+Quitting and looking for a job are equivalent in this setting.
 
-\[
-    q_{it} = \mathbbm{1}\{\text{Exp. utility of quitting} \geq \text{Disutility of quitting}\}
-\]
+From (1), we see that the determinants of the decision to quit are key to understand what happens when a prescription law is passed.
+This will, in turn, depend on the labor market opportunities the individual faces.
+The addicted individual faces a labor market with a job finding rate
 
-Quitting and looking for a job are equivalent.The market has a job finding rate
+\begin{equation}
+    p_{c(i)t} = \dfrac{m_{c(i)t}}{u_{c(i)t}}
+\end{equation}
 
-\[
-    p_{it} = m_{it}/u_{it} 
-\]
+where \(m_{c(i)t}\) is the number of matches and \(u_{c(i)t}\) is the number of unemployed workers in the county the agent lives, \(c(i)\). 
+I assume that matches are a non-decreasing function of vacancies \(v_{c(i)t}\)
 
-where \(m_{it}\) is the number of matches and \(u_{it}\) is the number of unemployed workers. Matches are a non-decreasing function of vacancies \(v_{it}\)
+\begin{equation}
+    m_{c(i)t} = M(v_{c(i)t}, u_{c(i)t})
+\end{equation}
 
-\[
-    m_{it} = M(v_{it}, u_{it})
-\]
+where \(M\left(\cdot \right) \) is a matching function. 
+Following \textcite{Flinn2010}, I assume that more vacancies are open where the minimum wage is less binding. 
+Then, it is easy to see that
 
-where \(M\left(\cdot \right) \) is a matching function. I assume that more vacancies are open where the minimum wage is less binding (\cite{flinn2010}). It follows that
+\begin{equation}
+    p_{\{w_{it} > w^{min}_t\}} \geq p_{\{w_{it} = w^{min}_t\}}
+\end{equation}
 
-\[
-    p_{w_{it} > w^{min}_t} \geq p_{w_{it} = w^{min}_t}
-\]
+that is, the job finding rate the individual faces is higher if the minimum wage is not binding.
 
-If \(w^{min}_t\) is not binding
+The decision to quit opioids is hence affected by the minimum wage in the following way:
+If the individual quits, she will look for a job which returns an utility \(u(w_{it}L_{it})\), where \(w_{it}\) is the wage she receives and \(L_{it}\) is the number of hours she works.
+The probability of finding a job is given by the job finding rate.
+If the product of the two is greater than the disutility of quitting, \(\mathcal{C}^{quit}\), the individual will quit.
+Formally
 
-\[
-    q_{w_{it} > w^{min}_t} = \mathbbm{1}\{p_{w_{it} > w^{min}_t}\cdot u(w_{it}L_{it}) \geq \mathcal{C}^{quit}\}
-\]
+\begin{equation}
+    q_{it}  = \mathbbm{1}\{p_{{c(i)t}} \cdot u(w_{it}L_{it}) \geq \mathcal{C}^{quit}\}
+\end{equation}
 
-Otherwise
+Unless a corner solution exists (i.e, the disutility of quitting is greater or lower than the maximum utility the individual can get from working), there must be some wage level such that
 
-\[
-    q_{w_{it} = w^{min}_t} = \mathbbm{1}\{p_{w_{it} = w^{min}_t}\cdot u(w^{min}_t L_{it}) \geq \mathcal{C}^{quit}\} 
-\]
+\begin{equation}
+    p_{\{w_{it} > w^{min}_t\}}\cdot u(w_{it}L_{it}) > \mathcal{C}^{quit} > p_{\{w_{it} = w^{min}_t\}}\cdot u(w^{min}_t L_{it})
+\end{equation}
 
-As long as
+that is, for some wage level, the job finding rate is such that the utility from working is greater than the disutility of quitting.
+Notice that this is irrespective of the differences in levels of wages, since the job finding rate is also playing a role.
 
-\[
-    p_{w_{it} > w^{min}_t}\cdot u(w_{it}L_{it}) > \mathcal{C}^{quit} > p_{w_{it} = w^{min}_t}\cdot u(w^{min}_t L_{it})
-\]
-
-we should see (after \(z_{t} = 1\)) an increase in participation rates where the minimum wage is less binding (\(O_{it} = 0\)); an increase in illegal opioid consumption where the minimum wage is more binding (\(O_{it} = h_{it}\)).
+If condition (6) holds, after a prescription law is passed (\(z_{t} = 1\)) we should see an increase in participation rates where the minimum wage is less binding (\(O_{it} = 0\)) but also an increase in illegal opioid consumption where the minimum wage is more binding (\(O_{it} = h_{it}\)).
+We test these hypotheses in the next sections.
 
@@
 
-\section*{Results}
+To uncover the heterogeneity in the effects of the opioid crisis on the labor market created by minimum wage rates, I use several data sources.
+The result is a panel of county-level data for the United States for the period 2003-2016 that includes labor, demographic and health variables.
+Where monthly data is not available, I use the data available for the year.
+The same holds for the minimum wage rates, which are set at the state level but are inputed to the counties.
+
+\textit{Labor market data and demographics}: I use the US Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics (LAUS) database to obtain county-level labor market statistics. 
+In particular, I obtain monthly data on labor force participation, employment, and unemployment rates at the county level.
+The dataset covers the period 1990-2023. US Census Bureau's Population Estimates Program (PEP) I am able to retrieve demograpic shares at the county-year level, from 1969 to 2020.
+The data gives me the population of each area for each gender, race and origin (hispainc or non-hispanic).
+Moreover, I can obtain the number of people in each of the 19 age groups considered, which is used to construct the labor force participation rate.
+
+\textit{Minimum wage bindingness}: I use the US Department of Labor's Changes in Basic Minimum Wages to obtain information on minimum wage rates at the state-level, from 1968 to 2023.
+Estimates of the wage distribution at the county level are requiered to construct a measure of minimum wage bindingness, which is outlined in the next section.
+For this purpose, I use the US Bureau of Labor Statistics (BLS) Occupational Employment and Wage Statistics (OEWS) database to obtain national industry-specific occupations and wages statistics.
+This dataset reports the number of workers in each occupation and the 10th, 25th, median, 75th and 90th wage percentiles for each industry, from 1997 to 2023.
+Moreover, I use the data from \textcite{Eckert2020} to obtain county-level industry-specific employment data, from 1975 to 2016.
+These last two datasets have a yearly frequency. 4-digit NAICS codes are used to match the industries across datasets.
+
+\textit{Prescription Drug Monitoring Programs (PDMPs)}: I use data from \textcite{Horwitz2021} to obtain information on Prescription Drug Monitoring Programs (PDMPs) from 1990 to 2019.
+The data is at the state-year level and reports the type of PDMP implemented in each state.
+Table 1 shows the timing of the implementation of Modern System (the last two columns) and Must Query (the first two) PDMPs across states.
+A Modern System PDMP is one where the electronic database becomes accessible to any authorized user, while a Must Query PDMP is one where the law mandates the prescriber to check the database before prescribing a listed opioid.
+This last type of PDMP is the one that is expected to have the most impact on opioid misuse, and the one I employ in the estimations.
+
+\textit{Prescriptions and overdose deaths}: I use data from the Centers for Disease Control (CDC) National Center for Health Statistics (NCHS) to obtain information on total drug poisoning mortality by county, from 2003 to 2021.
+To avoid suppression of data, \textcite{CLMort} employs Hierarchical Bayesian models with spatial and temporal random effects to estimate the number of deaths where the data is not available.
+I further use the Multiple Cause of Death database (\cite{CDCMCD9920}), which reports overdose deaths at the state-year level for different drugs, from 1999 to 2020, to obtain information on overdose deaths related to heroin (T.40.1), methadone (T.40.2), other opioids (T.40.3) and other synthetic narcotics (T.40.4).
+The codes reffer to the Multiple Causes of Death ICD-10 classification. 
+Overdose deaths are identified using the Underlying Cause of Death ICD-10 codes X40-X44, X60-X64, X85, and Y10-Y14.
+
+Finally, I use the Drug Enforcement Administration's Automation of Reports and Consolidated Orders System (ARCOS) to obtain opioid distribution data, from 2006 to 2014.
+This dataset records the distribution of all Schedule II substances by active ingredient at the 3-digit ZIP code level.
+The data is aggregated at the county-month level.
+The shorter time span does not affect the main specifications, since I use prescription rates only to assure that the PDMPs are effective.
+% Schedule II -> oxycodone and hydrocodone
+
+\section*{Findings: how minimum wages shaped the economic effects of the opioid epidemic}
+
+This section presents the main results of the paper. 
+I start describing the construction of the Kaitz-\(\rho\) index, which is used to measure the bindingness of the minimum wage.
+Then, I present the main especification of the paper, which is employed to estimate the effects of Prescription Drug Monitoring Programs on labor force participation and unemployment rates.
+The coefficients of the regression are compared across the spatial distribution of minimum wage bindingness measures.
+Finally, I repeat the same exercise using overdose deaths for different opioids as the dependent variable.
+
+\subsection*{The Kaitz-\(\rho\) index}
+
+The argument outlined above hinges on the assumption that the minimum wage can affect the labor market opportunities of low-educated workers.
+In other words, at least in some areas it must be the case that the minimum wage is an effective, or binding, constraint on the labor market.
+A common approach to measure this bindingness is through the Kaitz-\(\rho\) index, which can be defined as
+
+\begin{equation}
+    \text{\textit{kaitz}}_{ct}(\rho) \equiv \log{w^{\text{\textit{min}}}_{s(c)t}} - \log{w^{\rho}_{ct}}
+\end{equation}
+
+where \(w^{\text{\textit{min}}}_{s(c)t}\) is the minimum wage in the state where the county \(c\) is located at time \(t\), and \(w^{\rho}_{ct}\) is the \(\rho\) percentile of the wage distribution.
+This measure generates an spatial distribution of the bindingness of the minimum wage, which can be used to compare the effects of the opioid crisis across different labor markets.
+Figure 1 shows the distribution of the Kaitz-\(\rho\) index across counties in the United States for the year a Must Query PDMP is implemented in their states, for each of the different percentiles of the wage distribution included in the panel.
+For the main results, I use the 10th percentile, but the findings are similar across the distribution -- the counties for which the minimum wage is more binding with respect to lower percentiles tend to be the same as those for which it is more binding with respect to higher percentiles.
+
+The indices are obtained using the data on the wage distribution at the occupation-industry level, and the shares of employment of each industry in each county. 
+In particular, let \(w^{\rho}_{oit}\) be the percentile \(\rho\) of the wages at the occupation \(o\) in the industry \(i\) at time \(t\). 
+Then
+
+\begin{equation}
+    w^{\rho}_{it} = \sum_{o\in i} \dfrac{e_{oi}}{E_{it}}\,w^{\rho}_{oit}
+\end{equation}
+
+where \(e_{oi}\) is the number of workers in the occupation \(o\) and \(E_{it}\) is the total number of workers in the industry \(i\) at time \(t\).
+That is, by weighting the wage percentiles by the share of employment of each occupation in the industry, I obtain the wage percentile for the industry.
+After that, I compute
+
+\begin{equation}
+    w^{\rho}_{ct} = \sum_{i\in c} \dfrac{e_{ci}}{E_c}\,w^{\rho}_{it}
+\end{equation}
+
+where \(e_{ci}\) is the number of workers in the industry \(i\) in the county \(c\) and \(E_c\) is the total number of workers in the county.
+\(w^{\rho}_{ct}\) is the wage percentile for the county.
+Finally, the minimum wage rates at the state level are already included in the panel.
+
+\subsection*{Estimation}
+
+The main specification computes individual-level effects following \textcite{Arkhangelsky2024} and has the form
+
+\begin{equation}
+    y_{ct} = \alpha_c + \lambda_t + \sum_{k\geq h}\tau_{ck} + \beta X_{ct} + \nu_{ct}
+\end{equation}
+
+Here, \(y_{ct}\) is a labor market outcome, either the labor force participation rate or the unemployment rate, \(\alpha_c\) are county fixed effects, and \(\lambda_t\) are time fixed (year-month) effects.
+Relative time is represented with \(k = t - z_c\), the time difference of the calendar time from the implementation of a PDMP in the state, represented by \(z_s\).
+\(\tau_{ck}\) are the coefficients of interest and \(\nu_{ct}\) is an idiosyncratic shock.
+
+The covariates \(X_{ct}\) include demographic shares (race, origin, sex and age), sector shares of employment (at the 2-digits NAICS codes), and total drug poisoning mortality, which might be a relevant variable in policy-making.
+Crucial for identification, I assume that the state-level rollout of the PDMPs is exogenous at the county level after conditioning on the covariates.
+Moreover, this strategy requires that there is no anticipation of the policy up to \(h\) periods. 
+I assume no anticipation whatsoever, so that the coefficients are identified from the very start of the policy.
+In this context, this means that no addicted individual decided to look for a job or switch to heroin because of the expected implementation of the PDMP.
+
+After running (10), a distribution of individual (county-level) coefficients is obtained.
+These coefficients are then compared across the spatial distribution of the Kaitz-\(\rho\) index at the moment of treatment.
+The comparison is made by running a binscatter following \textcite{Cattaneo2024}, which allows for an optimal binning of the data.
+
+\subsection*{Labor force participation rates}
+
+\subsection*{Unemployment rates}
+
+\subsection*{Overdose deaths}
 
@@
 
+\newpage
+
 \printbibliography
 
+\newpage
+
+\begin{longtable}[c]{lcccc}
+    \caption{Prescriber Must Query and Modern Operational PDMPs Dates by State, from Horwitz et al. (2021)} \\
+    \toprule
+    \textbf{State} & \textbf{PMQ Year} & \textbf{PMQ Month} & \textbf{MOP Year} & \textbf{MOP Month} \\
+    \midrule
+    \endfirsthead
+
+    \caption[]{(continued)} \\
+    \toprule
+    \textbf{State} & \textbf{PMQ Year} & \textbf{PMQ Month} & \textbf{MOP Year} & \textbf{MOP Month} \\
+    \midrule
+    \endhead
+
+    \midrule \multicolumn{5}{r}{Continued on next page} \\
+    \endfoot
+
+    \bottomrule
+    \endlastfoot
+
+    Alabama & 2017 & 3 & 2006 & 4 \\
+    Alaska & 2018 & 6 & 2012 & 1 \\
+    Arizona & 2017 & 10 & 2008 & 12 \\
+    Arkansas & 2017 & 8 & 2013 & 5 \\
+    California & 2018 & 10 & 2009 & 9 \\
+    Colorado & - & - & 2008 & 2 \\
+    Connecticut & 2015 & 10 & 2008 & 7 \\
+    Delaware & - & - & 2012 & 8 \\
+    District of Columbia & - & - & 2016 & 10 \\
+    Florida & 2018 & 7 & 2011 & 10 \\
+    Georgia & 2018 & 7 & 2013 & 5 \\
+    Hawaii & 2018 & 7 & 2012 & 2 \\
+    Idaho & - & - & 2008 & 4 \\
+    Illinois & 2018 & 1 & 2009 & 12 \\
+    Indiana & - & - & 2007 & 7 \\
+    Iowa & 2018 & 7 & 2009 & 3 \\
+    Kansas & - & - & 2011 & 4 \\
+    Kentucky & 2012 & 7 & 1999 & 7 \\
+    Louisiana & 2014 & 8 & 2009 & 1 \\
+    Maine & 2017 & 1 & 2005 & 1 \\
+    Maryland & 2018 & 7 & 2013 & 12 \\
+    Massachusetts & 2014 & 12 & 2011 & 1 \\
+    Michigan & 2018 & 6 & 2003 & 1 \\
+    Minnesota & - & - & 2010 & 4 \\
+    Mississippi & - & - & 2008 & 7 \\
+    Missouri & - & - & - & - \\
+    Montana & - & - & 2012 & 10 \\
+    Nebraska & - & - & 2017 & 1 \\
+    Nevada & 2015 & 10 & 2011 & 2 \\
+    New Hampshire & 2016 & 5 & 2014 & 10 \\
+    New Jersey & 2015 & 11 & 2012 & 1 \\
+    New Mexico & 2013 & 2 & 2005 & 8 \\
+    New York & 2013 & 8 & 2013 & 6 \\
+    North Carolina & 2017 & 6 & 2007 & 7 \\
+    North Dakota & 2018 & 1 & 2008 & 10 \\
+    Ohio & 2015 & 4 & 2006 & 10 \\
+    Oklahoma & 2015 & 11 & 2006 & 7 \\
+    Oregon & - & - & 2011 & 9 \\
+    Pennsylvania & 2015 & 6 & 2016 & 8 \\
+    Rhode Island & 2016 & 6 & 2012 & 9 \\
+    South Carolina & 2017 & 5 & 2008 & 2 \\
+    South Dakota & - & - & 2012 & 3 \\
+    Tennessee & 2013 & 4 & 2010 & 1 \\
+    Texas & 2019 & 9 & 2012 & 8 \\
+    Utah & 2018 & 5 & 2006 & 1 \\
+    Vermont & 2013 & 11 & 2009 & 1 \\
+    Virginia & 2015 & 7 & 2006 & 6 \\
+    Washington & - & - & 2012 & 1 \\
+    West Virginia & 2012 & 6 & 2013 & 5 \\
+    Wisconsin & 2017 & 4 & 2013 & 6 \\
+    Wyoming & - & - & 2013 & 7 \\
+
+\end{longtable}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{kaitz_percentiles.png}
+    \caption{Only treated states are included in the sample.
+The figure shows the distribution of the Kaitz-\(\rho\) index across counties in the United States for the year a Must Query PDMP is implemented in their states.
+As lower percentiles from the wage distribution are considered, the minimum wage is more binding.}
+    \label{fig:kaitz_perc}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate.png}
+    \caption{}
+    \label{fig:lab_for_rate}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_ta.png}
+    \caption{}
+    \label{fig:lab_for_rate_ta}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_prescriptions.png}
+    \caption{}
+    \label{fig:prescriptions}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate.png}
+    \caption{}
+    \label{fig:unemp_rate}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_ta.png}
+    \caption{}
+    \label{fig:unemp_rate_ta}
+\end{figure}
+
+\newpage
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_heroin.png}
+    \caption{}
+    \label{fig:drugs}
+\end{figure}
+
+\newpage
+
+\section*{Appendix}
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_p.png}
+    \caption{}
+    \label{fig:lab_for_rate_p}
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_comp6.png}
+    \caption{}
+    \label{fig:lab_for_rate_comp6}
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_comp24.png}
+    \caption{}
+    \label{fig:lab_for_rate_comp24}
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_p.png}
+    \caption{}
+    \label{fig:unemp_rate_p}
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_comp6.png}
+    \caption{}
+    \label{fig:unemp_rate_comp6}
+\end{figure}
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_comp24.png}
+    \caption{}
+    \label{fig:unemp_rate_comp24}
+\end{figure}
+
 \end{document}

### 2024-06-16 code/data_analysis/results.R

@@ -1,0 +1,555 @@
+
+# Percentiles of the wage distribution
+percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
+
+# Load data's location
+data_location_pmq <- paste(
+  getwd(),
+  "data/processed/joined_data_pmq.csv",
+  sep = "/"
+  )
+
+# Load the data for Must Query PDMPs
+df_pmq <- read.csv(
+  data_location_pmq,
+  header = TRUE,
+  sep = ","
+  )
+
+# Include relative-to-treatement time markers and presc per capita
+df_pmq <- df_pmq |>
+  dplyr::mutate(
+    relative_to_treat_pmq = time_marker - first_treatment_pmq,
+    dosage_unit_pc = round(dosage_unit / population, 3)
+  )
+
+# Select the covariate columns
+names_covar_pmq <- names(
+  df_pmq[, c(grep("_ratio$", colnames(df_pmq)),
+        grep("model_values", colnames(df_pmq)))]
+  )
+
+# Prepare the covariates (as formula)
+covariates_pmq <- paste(
+  names_covar_pmq,
+  collapse = " + "
+  )
+
+# Compute the coefficient values
+unem_rate_pmq_effects  <- df_pmq |>
+  IndEffects(yname = "unem_rate",
+             iname = "fips",
+             tname = "time_marker",
+             kname = "relative_to_treat_pmq",
+             aname = "first_treatment_pmq",
+             covariates = covariates_pmq,
+             only_full_horizon = FALSE)
+
+emp_rate_pmq_effects  <- df_pmq |>
+  IndEffects(yname = "emp_rate",
+             iname = "fips",
+             tname = "time_marker",
+             kname = "relative_to_treat_pmq",
+             aname = "first_treatment_pmq",
+             covariates = covariates_pmq,
+             only_full_horizon = FALSE)
+
+lab_force_rate_pmq_effects  <- df_pmq |>
+  IndEffects(yname = "lab_force_rate",
+             iname = "fips",
+             tname = "time_marker",
+             kname = "relative_to_treat_pmq",
+             aname = "first_treatment_pmq",
+             covariates = covariates_pmq,
+             only_full_horizon = FALSE)
+
[…]

### 2024-06-17 slides/paper.tex
replication

@@
 
-% Pablo Muñoz: clarificar a que me refiero con binding
-
 \newpage
@@
 The so-called opioid epidemic, characterized by a rising number of deaths due to opioid overdose, has caused more than 680,000 fatalities between 1999 and 2022 (\cite{CDCMCD9920}, \cite{CDCMCD1822}), mostly among non-college-educated Americans.
-The sheer magnitude of this crisis has left a significant footprint on the labor market, prompting economists to study the relationship between opioid misuse and labor market outcomes.
+The sheer magnitude of this crisis has significantly impacted the labor market, prompting economists to study the relationship between opioid misuse and labor market outcomes.
 However, little is known about how labor market regulations have shaped this relationship.
@@
 
+\begin{comment}
     This paper argues that minimum wage rates affect the relationship between opioid misuse and labor market outcomes.
@@
     My estimations support this intuition, showing increases in labor force participation and unemployment rates as a result of prescription policies only where the minimum wage was not binding, while a surge in heroin consumption occurred where it was.
+\end{comment}
+
+This paper argues that minimum wage rates influence the relationship between opioid misuse and labor market outcomes.
+I measure this relationship using the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) across states, which require authorized users to check a patient's prescription history before prescribing opioids.
+These programs force addicts to either quit opioids or switch to illegal alternatives, with the choice depending on available labor market opportunities.
+Quitting opioids, though difficult, improves health and job prospects.
+However, in areas with high (binding) minimum wages, reduced labor demand might lower the benefits of quitting.
+My estimations show increases in labor force participation and unemployment rates only where the minimum wage was not binding, while a surge in heroin consumption occurred where it was.
 
@@
 
-% Condensate the whole background part into a single (two) paragraph
+% Condensate the whole background part into a single (two) paragraph maybe not mention lit yet
+
 These findings contribute to the growing body of literature studying the economic effects of the opioid crisis.
-Among this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation.
+Within this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation.
 This perspective has led economists to focus on supply-side interventions to explore the causal relationship between opioid consumption and labor market outcomes.
 While well-intentioned, many of these policies have had unintended consequences, such as the rise in heroin consumption following the reformulation of OxyContin.
-Prescription Drug Monitoring Programs (PDMPs), however, have been less scrutinized, with recent research indicating that, in their more stringent form, they can effectively reduce prescription opioid misuse.
+However, Prescription Drug Monitoring Programs (PDMPs), have been less scrutinized.
+Recent research indicates that, in their more stringent form, these programs can effectively reduce prescription opioid misuse.
+
+\begin{comment}
+    These findings contribute to the growing body of literature studying the economic effects of the opioid crisis and of the policies implemented to curb it.
+    Within this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation.
+    This perspective has led economists to focus on supply-side interventions to explore the causal relationship between opioid consumption and labor market outcomes.
+    Prescription Drug Monitoring Programs (PDMPs) 
+    Recent research indicates that, in their more stringent form, these programs can effectively reduce prescription opioid misuse.
+\end{comment}
 
@@
 To understand how these regulations have shaped the ongoing crisis, I focus on minimum wage rates, which have increased in all states at different speeds during the same period as the opioid crisis, thereby affecting the labor market opportunities of low-educated workers.
-Critical to our argument is not the rates themselves, but how binding they are across different labor markets: the more binding the minimum wage, the less incentives workers have to quit opioids, as they face a tougher job market search.
+Critical to our argument are not the rates themselves, but how binding they are across different labor markets: the more binding the minimum wage, the less incentives workers have to quit opioids, as they face a tougher job market search.
 
@@
 The BLS's Occupational Employment and Wage Statistics (OEWS) database provides national industry-specific occupations and wages statistics, \textcite{Eckert2020} offers county-level industry-specific employment data, and the US Department of Labor's Changes in Basic Minimum Wages provides information on minimum wage rates; all of these data are relevant for constructing a measure of minimum wage bindingness.
+
 The Centers for Disease Control (CDC) National Vital Statistics System offers data on total drug poisoning mortality by county, while the CDC's Multiple Cause of Death database provides overdose deaths at the state-year level for different drugs.
@@
 
-\textbf{Contribution}: shedding light on how the labor effects of the Opioid Crisis have been affected by minimum wage rates
+Therefore, this paper sheds light on how labor market regulations, and in particular minimum wage rates, give rise to heterogeneities in the effects of the opioid crisis on the labor market, contributing to a better understanding of the ongoing crisis and the policies implemented to address it.
+More broadly, it contributes to the understanding of how labor market regulations interact with drug disorders.
 
-Interest in how labor market regulations interact with Substance Abuse Disorders
-
-These results shed light on how labor market regulations in general, and minimum wage rates in particular, give rise to heterogeneities in the effects of the opioid crisis on the labor market.
-Moreover, they highlight the importance of blah blah.
-
-OUTLINE of the paper (the paper is structured as follows...)
+The paper is structured as follows.
+First, it provides a summary of the opioid crisis and the effects it has had on the labor market as uncovered by the literature.
+Then, it discusses some of the policies implemented to curb the opioid misuse and its effectiveness.
+Later, I describe how different minimum wage rates can explain some of these policy consequences and the heterogeneity of the labor market outcomes. 
+A model is introduced to formalize the intuition behind this argument.
+The employed data is discussed before presenting the empirical strategy and obtained results.
 
@@
 
-In this section, I provide a summary of how the opioid crisis started and what effects it had on the labor market in general. 
-Then, I discuss some of the policies implemented to curb the opioid misuse, and how some of them resulted ineffective or simply backfired. 
-Finally, I describe how labor market incentives, and in particular minimum wage rates, can explain some of these policy failures and the heterogeneity of the labor market outcomes.
+% CHECKED
+In this section, I summarize the origins of the opioid crisis and its general effects on the labor market.
+I then discuss various policies implemented to curb opioid misuse, and how some of them resulted ineffective or counterproductive. 
+Finally, I explain how labor market incentives, in particular minimum wage rates, can account for some of these policy failures and the heterogeneity in labor market outcomes.
 
@@
 % CHECKED
-Although the significance of the impact of minimum wages on labor demand has been questioned since \textcite{Card1994} published their seminal paper, recent research shows that minimum wage increases can have negative effects on employment in certain contexts.
-For instance, \textcite{Cengiz2019} found that minimum wage increases in the United States led to a decrease in employment in tradeable sectors.
-This finding is consistent with \textcite{Harasztosi2019}, who observed that the impact on employment is larger in industries where passing costs to consumers is more difficult, using data from Hungary.
-\textcite{Clemens2019} showed that minimum wage increases in the United States led to a decrease in employment in states bound by the federal minimum wage increase during the Great Recession.
-Furthermore, \textcite{Clemens2021} showed that large increases in minimum wages have reduced employment rates among individuals with low levels of education, in accordance with \textcite{Neumark2004},although the effect is negligible for small increases.
-In a recent paper, \textcite{Engbom2022} pointed out that the effect of these increases on employment can be muted by reallocations of workers across firms.
-Crucially, they argue that lower-productivity firms cut vacancy creation and hiring, which in our context could lead to lower job finding rates where the minimum wage is more binding.
+Although the impact of minimum wages on labor demand has been questioned since \textcite{Card1994} published their seminal paper, recent research indicates that minimum wage increases can negatively affect employment in certain contexts.
+For instance, we know that minimum wage increases can lead to a decrease in employment in tradeable sectors (\cite{Cengiz2019}), and in industries where passing costs to consumers is more difficult (\cite{Harasztosi2019}).
+This impact seems to be more pronounced for low-skilled workers (\cite{Neumark2004}), although the effect is negligible for small increases (\cite{Clemens2021}).
+The macroeconomic situation can also influence the effect of minimum wage increases on employment if the former is particularly binding (\cite{Clemens2019}). 
+However, \textcite{Engbom2022} noted that these effects can be muted by reallocations of workers across firms.
+Crucially, they argue that lower-productivity firms reduce vacancy creation and hiring, which in our context could lead to lower job finding rates where the minimum wage is more binding.
 
@@
 
-From (1), we see that the determinants of the decision to quit are key to understand what happens when a prescription law is passed.
-This will, in turn, depend on the labor market opportunities the individual faces.
+From (1), we see that the determinants of the decision to quit are crucial to understanding the effects of a prescription law.
+This decision depends on the labor market opportunities available to the individual.
 The addicted individual faces a labor market with a job finding rate
@@
 
-The decision to quit opioids is hence affected by the minimum wage in the following way:
-If the individual quits, she will look for a job which returns an utility \(u(w_{it}L_{it})\), where \(w_{it}\) is the wage she receives and \(L_{it}\) is the number of hours she works.
+The decision to quit opioids is hence influenced by the minimum wage as follows:
+If the individual quits, she will seek a job that provides utility \(u(w_{it}L_{it})\), where \(w_{it}\) is the wage she receives and \(L_{it}\) is the number of hours she works.
 The probability of finding a job is given by the job finding rate.
-If the product of the two is greater than the disutility of quitting, \(\mathcal{C}^{quit}\), the individual will quit.
+If the product of the utility and the job finding rate is greater than the disutility of quitting, \(\mathcal{C}^{quit}\), the individual will quit.
 Formally
@@
 
-Unless a corner solution exists (i.e, the disutility of quitting is greater or lower than the maximum utility the individual can get from working), there must be some wage level such that
+Unless a corner solution exists (i.e., the disutility of quitting is greater or lower than the maximum utility the individual can get from working), there must be some wage level such that
 
@@
 
-that is, for some wage level, the job finding rate is such that the utility from working is greater than the disutility of quitting.
+that is, for some wage level, the job finding rate ensures that the utility from working exceeds the disutility of quitting.
 Notice that this is irrespective of the differences in levels of wages, since the job finding rate is also playing a role.
 
-If condition (6) holds, after a prescription law is passed (\(z_{t} = 1\)) we should see an increase in participation rates where the minimum wage is less binding (\(O_{it} = 0\)) but also an increase in illegal opioid consumption where the minimum wage is more binding (\(O_{it} = h_{it}\)).
+If condition (6) holds, after a prescription law is enacted (\(z_{t} = 1\)) we should observe an increase in participation rates where the minimum wage is less binding (\(O_{it} = 0\)) but also an increase in illegal opioid consumption where the minimum wage is more binding (\(O_{it} = h_{it}\)).
 We test these hypotheses in the next sections.
@@
 
+% CHECKED
 To uncover the heterogeneity in the effects of the opioid crisis on the labor market created by minimum wage rates, I use several data sources.
-The result is a panel of county-level data for the United States for the period 2003-2016 that includes labor, demographic and health variables.
-Where monthly data is not available, I use the data available for the year.
-The same holds for the minimum wage rates, which are set at the state level but are inputed to the counties.
+The result is a panel of county-level data for the United States for the period 2003-2016 encompassing labor market, demographic and health variables.
+Monthly data is utilized wherever available; otherwise, I use the data available for the year.
+Similarly, state-level minimum wage rates are allocated to the counties.
 
-\textit{Labor market data and demographics}: I use the US Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics (LAUS) database to obtain county-level labor market statistics. 
-In particular, I obtain monthly data on labor force participation, employment, and unemployment rates at the county level.
-The dataset covers the period 1990-2023. US Census Bureau's Population Estimates Program (PEP) I am able to retrieve demograpic shares at the county-year level, from 1969 to 2020.
-The data gives me the population of each area for each gender, race and origin (hispainc or non-hispanic).
+% CHECKED
+\textit{Labor market data and demographics}: 
+I use the US Bureau of Labor Statistics (BLS) Local Area Unemployment Statistics (LAUS) database to access county-level labor market statistics. 
+Specifically, I gather monthly data on labor force participation, employment, and unemployment rates at the month-county level.
+This dataset covers the period 1990-2023. 
+Additionally, through the US Census Bureau's Population Estimates Program (PEP), I obtain demographic information at the county-year level spanning from 1969 to 2020. 
+This dataset provides the population counts of each county by gender, race and origin (Hispainc or Non-Hispanic), on a yearly basis.
 Moreover, I can obtain the number of people in each of the 19 age groups considered, which is used to construct the labor force participation rate.
 
-\textit{Minimum wage bindingness}: I use the US Department of Labor's Changes in Basic Minimum Wages to obtain information on minimum wage rates at the state-level, from 1968 to 2023.
-Estimates of the wage distribution at the county level are requiered to construct a measure of minimum wage bindingness, which is outlined in the next section.
-For this purpose, I use the US Bureau of Labor Statistics (BLS) Occupational Employment and Wage Statistics (OEWS) database to obtain national industry-specific occupations and wages statistics.
+% CHECKED
+\textit{Minimum wage bindingness}: 
+I utilize the US Department of Labor's Changes in Basic Minimum Wages dataset to gather information on minimum wage rates at the state level from 1968 to 2023.
+Estimates of the wage distribution at the county level are requiered to construct a measure of minimum wage bindingness, as outlined in the next section.
+For this purpose, I rely on the US Bureau of Labor Statistics (BLS) Occupational Employment and Wage Statistics (OEWS) database to obtain national industry-specific occupations and wages statistics.
 This dataset reports the number of workers in each occupation and the 10th, 25th, median, 75th and 90th wage percentiles for each industry, from 1997 to 2023.
-Moreover, I use the data from \textcite{Eckert2020} to obtain county-level industry-specific employment data, from 1975 to 2016.
-These last two datasets have a yearly frequency. 4-digit NAICS codes are used to match the industries across datasets.
+Additionally, I use the data from \textcite{Eckert2020} to obtain county-level industry-specific employment data spanning from 1975 to 2016.
+These last two datasets have an annual frequency. 4-digit NAICS codes are used to match the industries across datasets.
 
-\textit{Prescription Drug Monitoring Programs (PDMPs)}: I use data from \textcite{Horwitz2021} to obtain information on Prescription Drug Monitoring Programs (PDMPs) from 1990 to 2019.
-The data is at the state-year level and reports the type of PDMP implemented in each state.
-Table 1 shows the timing of the implementation of Modern System (the last two columns) and Must Query (the first two) PDMPs across states.
-A Modern System PDMP is one where the electronic database becomes accessible to any authorized user, while a Must Query PDMP is one where the law mandates the prescriber to check the database before prescribing a listed opioid.
-This last type of PDMP is the one that is expected to have the most impact on opioid misuse, and the one I employ in the estimations.
+% CHECKED
+\textit{Prescription Drug Monitoring Programs (PDMPs)}: 
+I use data from \textcite{Horwitz2021} to obtain information on Prescription Drug Monitoring Programs (PDMPs) from 1990 to 2019.
+This dataset offers state-year level detailing the implementation of PDMP types across states, as illustrated in Table 1. 
+PDMPs are categorized into Modern System and Must Query types. 
+A Modern System PDMP grants electronic database access to authorized users, whereas a Must Query PDMP mandates prescribers to consult the database before prescribing listed opioids. 
+This last type of PDMP is the one most likely to have an impact on opioid misuse, and it forms the basis of my estimations.
 
-\textit{Prescriptions and overdose deaths}: I use data from the Centers for Disease Control (CDC) National Center for Health Statistics (NCHS) to obtain information on total drug poisoning mortality by county, from 2003 to 2021.
-To avoid suppression of data, \textcite{CLMort} employs Hierarchical Bayesian models with spatial and temporal random effects to estimate the number of deaths where the data is not available.
+% CHECKED
+\textit{Prescriptions and overdose deaths}: 
+I use data from the Centers for Disease Control (CDC) National Center for Health Statistics (NCHS) to obtain information on total drug poisoning mortality by county, from 2003 to 2021.
+To overcome the suppression of data following anonimity concerns, \textcite{CLMort} employs Hierarchical Bayesian models with spatial and temporal random effects to estimate the number of deaths where the data is not available.
 I further use the Multiple Cause of Death database (\cite{CDCMCD9920}), which reports overdose deaths at the state-year level for different drugs, from 1999 to 2020, to obtain information on overdose deaths related to heroin (T.40.1), methadone (T.40.2), other opioids (T.40.3) and other synthetic narcotics (T.40.4).
@@
 
-Finally, I use the Drug Enforcement Administration's Automation of Reports and Consolidated Orders System (ARCOS) to obtain opioid distribution data, from 2006 to 2014.
+% CHECKED
+Finally, I use the Drug Enforcement Administration's Automation of Reports and Consolidated Orders System (ARCOS) to obtain opioid distribution data spanning from 2006 to 2014.
 This dataset records the distribution of all Schedule II substances by active ingredient at the 3-digit ZIP code level.
-The data is aggregated at the county-month level.
-The shorter time span does not affect the main specifications, since I use prescription rates only to assure that the PDMPs are effective.
-% Schedule II -> oxycodone and hydrocodone
+The data is aggregated on semi-synthetic opioid analgesics (i.e. oxycodone, hydrocodone) prescribed at the county-month level.
+The shorter time span does not affect the main specifications, since I use prescription rates only to assure that the PDMPs are effective in a complementary exercise.
 
@@
 
+Using the labor force participation rate as the dependent variable, the results of the main specification are displayed in Figure 2.
+The figure shows the distribution of the coefficients of the impact of the passing of the Must Query PDMPs on labor force participation rates across different levels of minimum wage bindingness, for different periods relative to treatment.
+In the upper right panel we can see the immediate impact of the policy.
+For those counties where the minimum wage is less binding, the effect is positive (1.37 p.p.) and significant (0.513, 2.22) at conventional levels.
+The effects become smaller and more uncertain as the minimum wage becomes more binding.
+For those counties in which the minimum wage has a higher impact on labor demand, the point estimate becomes negative (-0.693) but not significant (-2.54, 1.16).
+
+The effect seems to be persistent in the short run (upper right and lower right panels), but fades away after two years (lower right)\footnote{The confidence interval of the most binnding counties is not displayed in this last panel for visualization purposes.
+The values are (-3.10, 1.90)}.
+Figure 3 shows the time averages of the coefficients for two years after the law is passed.
+The results remain positive and significant for those counties in which the minimum wage is less binding (1.06 p.p., (0.584, 1.55)), while they are not significant for those where the minimum wage was more binding.
+
+The results seem to be driven mostly by the impact of the policy on prescription rates.
+Although the overlap between the passing of the policies (Table 1) and the data for prescription rates (2006-2014) is scarce, we can still obtain som estimates on the impact of Must Query PDMPs on prescription rates per capita.
+Figure 4 runs the model (10) with the prescription rates per capita as the dependent variable.
+The two months after the treatment is passed are displayed.
+We see that the policy is able to decrease the amount of prescriptions, with the biggest drops occuring where the minimum wage is the least and the most binding (the effects in these areas are not statistically different).
+Although the target of these policies where not to decrease prescriptions \textit{per se}, but only tackling misuse cases, the results suggest that the policies are effective in reducing the amount of opioids prescribed on the aggregate.
+
+Moreover, this seems to be the main channel through which the policies affect labor market outcomes: Figure 8 in the appendix shows that the relation between labor market outcomes and minimum wage bindingness is not significant when controlling for prescription rates.
+Figures 9 and 10 in the Appendix show the the point estimates of the binscatter for different Kaitz indices, for six months and two years after the policy is passed.
+It is easy to see that the results do not change much across different measures, this being especially true the closer in time the treatment was implemented.
+
+\begin{comment}
+After six months:
+1.79 (0.955, 2.63)
+0.333 (-0.665, 1.33)
+
+After two years:
+-0.118 (-1.66, 1.42)
+-0.599 (-3.10, 1.90)
+
+Time averages:
+1.06  (0.584, 1.55)
+0.182 (-0.335, 0.698)
+\end{comment}
+
 \subsection*{Unemployment rates}
+
+The analysis is repeated using the unemployment rate as the dependent variable.
+The results are displayed in the same fashion in Figure 5.
+The distribution of the coefficients reveal that almost all the positive effect it was found in the labor force participation rates translated into an increase in the unemployment rate.
+For those counties where the minimum wage is less binding, the effect is positive (1.02 p.p.) and significant (1.46, 0.575) at conventional levels. 
+The point estimate is only 0.35 p.p. below the one it was found for the labor force participation rate.
+Although a positive effect is also found for those counties where the minimum wage is more binding (0.640 p.p.), the effect is not significant (-0.0252, 1.31).
+
+These results shouldn't be surprising, at least in the short run: if people with severe opioid addictions start looking for a job after the implementation of a PDMP, they might not be able to find one immediately even in those counties where the minimum wage is less binding.
+Strikingly though, the impact on unemployment persists after two years.
+The lower right panel shows that, even after the increase in labor force participation rates fades away, the increase in unemployment rates remains high (2.77 p.p.) and significant (1.87, 3.68) where the minimum wage is less binding.
+Whether the causes of this persistence are related to opioid consumption (through, say, relapsing) or to characterisctics of the workers or the labor market (say, if these workers face more frictional unemployment) remains an open question.
+
+\begin{comment}
+    After six months:
+    1.01   1.48   0.543
+    -0.0981 0.199 -0.396
+
+    After two years:
+    2.77   3.68   1.87
+    -0.0720 0.265 -0.409 
+\end{comment}
+
+The time average of the coefficients, displayed in Figure 6, confirm our previous result.
+The effect of the policy on unemployment rates is positive and significant for those counties where the minimum wage is less binding (1.25 p.p., (0.97, 1.52)), while it is not significant for those where the minimum wage is more binding (0.025 p.p., (-0.11, 0.16)).
+Again, the results seem to be driven by the impact of the policy on prescription rates, especially in the medium run: Figure 11 in the Appendix shows that the impact stops being significantly different from zero after one year once we control for prescriptions per capita.
+Figures 12 and 13 in the Appendix show the binscatter for different Kaitz indices, for six months and two years after the policy is passed.
+The results do not change substantially across different measures.
 
@@
 
-\section*{Conclusions}
+Finally, and in accordance with the model outlined above, I show some evidence that the rollout of Must Query PDMPs has had a positive impact on the consumption of illegal opioids in those counties where the minimum wage was more binding.
+For this purpose, I run the same estimation strategy as before, but using the number of overdose deaths per capita for different opioids as the dependent variable. 
+The data is at the state-year level, and the number of total poisoning deaths is not used as a control variable.
+Figure 7 displays the scatter plot of the time average of the coefficients on the Kaitz-0.10 index of the states, for the two years after the policy is passed.
+The opioids included are heroin, methadone, other opioids and other synthetic narcotics, following the ICD-10 classification.
+
+The coefficients are regressed on the Kaitz index and a constant. 
+Tables 2-5 show the values of the regression coefficients for each of the different opioids.
+Although admittedly weaker than the evidence presented in previous sections due to data constaints, it is interesting that the only marginally significant relation is found for heroin.
+This suggest that the positive relation between black market switching and low labor demand is present, but we cannot assert this relationship with the same degree of certainty.
+
+It seems implausible that part of the effect is absorbed by the consumption of other opioids.
+Methadone is often used as a treatment for opioid adiction and can be legally used under a doctor's supervision.
+Other opioids include drugs such as morphine and codeine, but also semi-synthetic opioid analgesics (e.g.
+drugs such as oxycodone, hydrocodone, hydromorphone, and oxymorphone), some of them targeted by the PDMPs.
+Finally, although the category of other synthetic narcotics includes fentanyl and its analogues, its use didn't used to be as widespread as it is today during the period considered.
+
+\section*{Conclusion}
+
+This paper contributes to the understanding of how labor market regulations, particularly minimum wage rates, shape the relationship between opioid misuse and labor market outcomes.
+By leveraging the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) across states, this study provides a nuanced view of how labor market opportunities influence the incentives for opioid users to quit or switch to illegal alternatives.
+
+The findings reveal that labor force participation and unemployment rates increased as a result of prescription policies only in areas where the minimum wage was not binding.
+In contrast, a surge in heroin consumption occurred where the minimum wage was binding, highlighting the complex interplay between labor market conditions and opioid misuse.
+These results underscore the importance of considering labor market regulations when evaluating the economic impacts of the opioid crisis.
+
+Furthermore, this study adds to the growing body of literature that examines the economic effects of the opioid epidemic.
+While previous research has primarily focused on the direct consequences of opioid misuse on labor market outcomes, this paper emphasizes the role of labor market regulations in mediating these effects.
+The heterogeneity in the impact of prescription policies across different levels of minimum wage bindingness suggests that policies aimed at curbing opioid misuse must account for the broader economic context to be effective.
+
+In summary, the analysis demonstrates that higher minimum wages, by reducing labor demand for low-skilled workers, can diminish the incentives for opioid users to quit, thereby exacerbating the shift towards illegal alternatives like heroin.
+These insights highlight the need for a comprehensive approach to addressing the opioid crisis, one that integrates labor market policies with public health interventions to effectively mitigate the epidemic's adverse effects on both health and economic outcomes.
+
+Future research should continue to explore the interplay between labor market regulations and health crises, considering the potential unintended consequences of well-intentioned policies.
+By doing so, policymakers can design more effective interventions that address the root causes of public health crises while also supporting economic stability and growth.
+
+[MOVING FORWARD]
 
@@
 
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lcccc}
+    \toprule
+     & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr($>|t|$)} \\
+    \midrule
+    Intercept  & 5.600 & 2.135 & 2.623 & 0.0211 * \\
+    Kaitz-0.10 & 9.613 & 5.021 & 1.914 & 0.0778 . \\
+    \midrule
+    \multicolumn{5}{l}{\textit{Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1}} \\
+    \midrule
+    \multicolumn{5}{l}{\textbf{Residual standard error}: 1.95 on 13 degrees of freedom} \\
+    \multicolumn{5}{l}{\textbf{Multiple R-squared}: 0.2199, \textbf{Adjusted R-squared}: 0.1599} \\
+    \multicolumn{5}{l}{\textbf{F-statistic}: 3.665 on 1 and 13 DF, \textbf{p-value}: 0.07783} \\
+    \bottomrule
+    \end{tabular}
+    \caption{Regression Results: Heroin Overdose Deaths}
+    \label{tab:regression_results_heroin}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lcccc}
+    \toprule
+     & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr($>|t|$)} \\
+    \midrule
+    Intercept  & -0.4800 & 0.8021 & -0.598 & 0.560 \\
+    Kaitz-0.10 & -0.6460 & 1.8900 & -0.342 & 0.738 \\
+    \midrule
+    \multicolumn{5}{l}{\textit{Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1}} \\
+    \midrule
+    \multicolumn{5}{l}{\textbf{Residual standard error}: 0.7576 on 13 degrees of freedom} \\
+    \multicolumn{5}{l}{\textbf{Multiple R-squared}: 0.008906, \textbf{Adjusted R-squared}: -0.06733} \\
+    \multicolumn{5}{l}{\textbf{F-statistic}: 0.1168 on 1 and 13 DF, \textbf{p-value}: 0.738} \\
+    \bottomrule
+    \end{tabular}
+    \caption{Regression Results: Methadone Overdose Deaths}
+    \label{tab:regression_results_methadone}
+\end{table}
+
+\newpage
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lcccc}
+    \toprule
+     & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr($>|t|$)} \\
+    \midrule
+    Intercept  & -3.486 & 2.530 & -1.378 & 0.192 \\
+    Kaitz-0.10 & -7.537 & 5.949 & -1.267 & 0.227 \\
+    \midrule
+    \multicolumn{5}{l}{\textit{Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1}} \\
+    \midrule
+    \multicolumn{5}{l}{\textbf{Residual standard error}: 2.31 on 13 degrees of freedom} \\
+    \multicolumn{5}{l}{\textbf{Multiple R-squared}: 0.1099, \textbf{Adjusted R-squared}: 0.04143} \\
+    \multicolumn{5}{l}{\textbf{F-statistic}: 1.605 on 1 and 13 DF, \textbf{p-value}: 0.2274} \\
+    \bottomrule
+    \end{tabular}
+    \caption{Regression Results: Other opioids Overdose Deaths}
+    \label{tab:regression_results_otheropioids}
+\end{table}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{lcccc}
+    \toprule
+     & \textbf{Estimate} & \textbf{Std. Error} & \textbf{t value} & \textbf{Pr($>|t|$)} \\
+    \midrule
+    Intercept  & 4.944 & 5.995 & 0.825 & 0.424 \\
+    Kaitz-0.10 & 8.056 & 14.097 & 0.571 & 0.577 \\
+    \midrule
+    \multicolumn{5}{l}{\textit{Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1}} \\
+    \midrule
+    \multicolumn{5}{l}{\textbf{Residual standard error}: 5.475 on 13 degrees of freedom} \\
+    \multicolumn{5}{l}{\textbf{Multiple R-squared}: 0.0245, \textbf{Adjusted R-squared}: -0.05053} \\
+    \multicolumn{5}{l}{\textbf{F-statistic}: 0.3266 on 1 and 13 DF, \textbf{p-value}: 0.5774} \\
+    \bottomrule
+    \end{tabular}
+    \caption{Regression Results: synthetic opioids Overdose Deaths}
+    \label{tab:regression_results_synthetic}
+\end{table}
+
+\newpage
+
 \section*{Appendix}

### 2024-06-17 code/data_analysis/plots.R

@@ -1,15 +1,14 @@
 KaitzDensitiesPlot <- function(df) {
 
   # Define the plot title
-  plot_title <- "Distribution of Kaitz-p indices across counties, 
-  for different percentiles at treatment"
+  plot_title <- "Distribution of Kaitz-p indices across counties, for different \n percentiles at treatment (Must Query PDMP)"
 
   # Get the column names of the dataframe
   df_colnames <- df |> colnames()
 
   # Filter the dataframe to include only observations at treatment
   rel_to_treat_name <- df_colnames[grep("relative_to_treat_", df_colnames)]
-  df <- df[df[[rel_to_treat_name]] == 0,]
+  df <- df[df[[rel_to_treat_name]] == 0, ]
 
   # Define the colors for the density plots
   density_colors <- c(
@@ -61,8 +60,8 @@ KaitzDensitiesPlot <- function(df) {
     theme_classic() +
     theme(
       text = element_text(size = 16, family = "serif"),
-      plot.title = element_markdown(lineheight = 1.1),
-      legend.position = c(0.15, 0.80)
+      plot.title = element_text(hjust = 0.5, lineheight = 1.1),
+      legend.position = c(0.10, 0.80)
       )
 
   return(densities_plot)
@@ -104,9 +103,9 @@ TitleAndColors <- function(policy, outcome, percentile) {
     plot_color <- "#0098e9"
     plot_yrange <- ylim(-0.01, 0.01)
   } else if (outcome == "dosage_unit_pc") {
-    outcome_title <- "prescriptions_pc"
+    outcome_title <- "prescriptions per capita"
     plot_color <- "#0098e9"
-    plot_yrange <- ylim(-3.0, 3.0)
+    plot_yrange <- ylim(-1.5, 0.5)
   } else if (outcome == "deaths_pc") {
     outcome_title <- "opioid deaths_pc"
     plot_color <- "#0098e9"
@@ -159,6 +158,7 @@ IndividualEffectsPlot <- function(df,
 
   # Create an empty list to store the plots
   plot_collection <- list()
+  bins_collection <- list()
 
   for (period in periods){
 
@@ -166,8 +166,13 @@ IndividualEffectsPlot <- function(df,
     plot_data <- df[df[[paste0("relative_to_treat_", policy)]] == period, ]
 
     # Perform binscatter regression
[…]

### 2024-06-17 code/data_analysis/results.R

@@ -1,555 +1,602 @@
+regressions <- function(location) {
 
-# Percentiles of the wage distribution
-percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
+  # Percentiles of the wage distribution
+  percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
 
-# Load data's location
-data_location_pmq <- paste(
-  getwd(),
-  "data/processed/joined_data_pmq.csv",
-  sep = "/"
-  )
+  # Load data's location
+  data_location_pmq <- paste(
+    location,
+    "data/processed/joined_data_pmq.csv",
+    sep = "/"
+    )
 
-# Load the data for Must Query PDMPs
-df_pmq <- read.csv(
-  data_location_pmq,
-  header = TRUE,
-  sep = ","
-  )
+  # Load the data for Must Query PDMPs
+  df_pmq <- read.csv(
+    data_location_pmq,
+    header = TRUE,
+    sep = ","
+    )
 
-# Include relative-to-treatement time markers and presc per capita
-df_pmq <- df_pmq |>
-  dplyr::mutate(
-    relative_to_treat_pmq = time_marker - first_treatment_pmq,
-    dosage_unit_pc = round(dosage_unit / population, 3)
-  )
+  # Include relative-to-treatement time markers and presc per capita
+  df_pmq <- df_pmq |>
+    dplyr::mutate(
+      relative_to_treat_pmq = time_marker - first_treatment_pmq,
+      dosage_unit_pc = round(dosage_unit / population, 3)
+    )
 
-# Select the covariate columns
-names_covar_pmq <- names(
-  df_pmq[, c(grep("_ratio$", colnames(df_pmq)),
-        grep("model_values", colnames(df_pmq)))]
-  )
+  # Select the covariate columns
+  names_covar_pmq <- names(
+    df_pmq[, c(grep("_ratio$", colnames(df_pmq)),
+          grep("model_values", colnames(df_pmq)))]
+    )
 
-# Prepare the covariates (as formula)
-covariates_pmq <- paste(
-  names_covar_pmq,
-  collapse = " + "
-  )
+  # Prepare the covariates (as formula)
+  covariates_pmq <- paste(
+    names_covar_pmq,
+    collapse = " + "
+    )
 
-# Compute the coefficient values
-unem_rate_pmq_effects  <- df_pmq |>
-  IndEffects(yname = "unem_rate",
[…]

### 2024-06-17 code/data_construction/data_merge.R

@@ -1,855 +1,844 @@
-library(tidyverse)
+datasets <- function(location) {
+  # Load the data
 
-# Clear the workspace
-rm(list = ls())
-options(max.print = 250)
+  # Unemployment rate data
+  unemp_rate_path <- "data/intermediate/labor_market_outcomes/unemployment_rate.csv"
 
-# Set the seed for reproducibility
-set.seed(123)
+  unemp_rate <- read.csv(
+    file.path(getwd(), unemp_rate_path),
+    header = TRUE,
+    sep = ","
+    )
 
-# Set the file location
-file_location <- dirname(rstudioapi::getSourceEditorContext()$path)
-setwd(gsub("/code/data_construction", "", file_location))
+  # Employment rate data
+  emp_rate_path <- "data/intermediate/labor_market_outcomes/employment_rate.csv"
 
-# Load the data
+  emp_rate <- read.csv(
+    file.path(getwd(), emp_rate_path),
+    header = TRUE,
+    sep = ","
+    )
 
-# Unemployment rate data
-unemp_rate_path <- "data/intermediate/labor_market_outcomes/unemployment_rate.csv"
+  # Labor force data
+  lab_force_path <- "data/intermediate/labor_market_outcomes/labor_force.csv"
 
-unemp_rate <- read.csv(
-  file.path(getwd(), unemp_rate_path),
-  header = TRUE,
-  sep = ","
-  )
+  lab_force <- read.csv(
+    file.path(getwd(), lab_force_path),
+    header = TRUE,
+    sep = ","
+    )
 
-# Employment rate data
-emp_rate_path <- "data/intermediate/labor_market_outcomes/employment_rate.csv"
+  # PDMPs data
+  horwitz_path <- "data/source/pdmps/horwitz2020.csv"
 
-emp_rate <- read.csv(
-  file.path(getwd(), emp_rate_path),
-  header = TRUE,
-  sep = ","
-  )
+  horwitz <- read.csv(
+    file.path(getwd(), horwitz_path),
+    header = TRUE,
+    sep = ","
+    )
 
-# Labor force data
-lab_force_path <- "data/intermediate/labor_market_outcomes/labor_force.csv"
+  # NAICS shares of the counties data
+  naics_path <- "data/intermediate/sector_shares/naics_shares_merged.csv"
 
-lab_force <- read.csv(
-  file.path(getwd(), lab_force_path),
-  header = TRUE,
-  sep = ","
-  )
+  naics <- read.csv(
[…]

### 2024-06-17 code/data_construction/labor_market_outcomes.py

@@ -1,5 +1,3 @@
-import pandas as pd
-
 '''
 returns unemployment_rate.csv, unemployment.csv, employment.csv, labor_force.csv, 
 employment_population_ratio.csv, labor_force_participation_rate.csv

### 2024-06-17 code/data_construction/minwage_bindingness.py

@@ -1,5 +1,3 @@
-import pandas as pd
-
 # Load data_naics dictionary
 data_naics = dict()
 naics_df = ['four_naics_02to05', 'four_naics_06to10', 'four_naics_11to16']

### 2024-06-17 code/data_construction/minwage_cleaning.py

@@ -1,5 +1,3 @@
-import pandas as pd
-
 '''
 returns minwage_clean_states_ub.csv and minwage_clean_fed_ub.csv
 

### 2024-06-17 code/data_construction/sector_shares.py

@@ -55,9 +55,11 @@ for naic, df in data_naics.items():
     data_naics[naic].columns = data_naics[naic].columns.droplevel().rename(None)
     data_naics[naic] = data_naics[naic].reset_index()
     data_naics[naic].columns = 'emp_' + data_naics[naic].columns + '_ratio'
+    data_naics[naic] = data_naics[naic].rename({'emp_year_ratio' : 'year', 'emp_fips_ratio' : 'fips'}, axis=1)
+    data_naics[naic]['state_code'] = data_naics[naic]['fips'].str[:2]
+    data_naics[naic]['county_code'] = data_naics[naic]['fips'].str[2:]
 
-#data_naics['four_naics_02to05']
-# RENAME WHEN WIFI AVAILABLE
+data_naics['four_naics_02to05']
 
 # Save the dataframes as CSV files
 data_naics['four_naics_02to05'].to_csv(r'./data/intermediate/sector_shares/naics_shares_02to05.csv', sep=',', index=False)
@@ -71,9 +73,6 @@ naics_shares_11to16 = spark.createDataFrame(data_naics['four_naics_11to16'])
 
 # Union the Spark DataFrames
 naics_shares = naics_shares_02to05.union(naics_shares_06to10).union(naics_shares_11to16)
-
-# Drop the 'emp' column
-#naics_shares = naics_shares.drop('emp')
 
 # Save the merged DataFrame as a CSV file
 naics_shares.toPandas().to_csv(r'./data/intermediate/sector_shares/naics_shares_merged.csv', sep=',', index=False)

### 2024-06-17 code/data_construction/wages_sector.py

@@ -1,7 +1,3 @@
-import pandas as pd
-import glob
-import re
-
 '''
 https://www.bls.gov/oes/tables.htm
 

### 2024-06-17 data/plots.R

@@ -1,649 +1,0 @@
-
-KaitzDensitiesPlot <- function(df){
-  
-  df_colnames <- df |> colnames()
-  
-  rel_to_treat_name <- df_colnames[grep("relative_to_treat_", df_colnames)]
-
-  df <- df[df[[rel_to_treat_name]] == 0,]
-  
-  density_colors <- c("0.10" = "#0098e9",
-                      "0.25" = "#ff5ca8",
-                      "0.50" = "#5aa800",
-                      "0.75" = "#f29318",
-                      "0.90" = "#B79F00")
-  
-  densities_plot <- ggplot(df) +
-    geom_density(
-      aes(x = kaitz_pct10, ..count../sum(..count..), color = "0.10"), linewidth = 1
-      ) +
-    geom_density(
-      aes(x = kaitz_pct25, ..count../sum(..count..), color = "0.25"), linewidth = 1
-      ) + 
-    geom_density(
-      aes(x = kaitz_pct50, ..count../sum(..count..), color = "0.50"), linewidth = 1
-      ) + 
-    geom_density(
-      aes(x = kaitz_pct75, ..count../sum(..count..), color = "0.75"), linewidth = 1
-      ) + 
-    geom_density(
-      aes(x = kaitz_pct90, ..count../sum(..count..), color = "0.90"), linewidth = 1
-      ) +
-    scale_color_manual(
-      name = element_blank(), values = density_colors
-      )  +
-    ylab(
-      ""
-      ) + 
-    xlab(
-      "Kaitz-p index"
-      ) + 
-    labs(
-      title = "Distribution of Kaitz-p indices across counties, for different percentiles<br>at treatment"
-      )
-  
-  densities_plot <- densities_plot + 
-    theme_classic() + 
-    theme(
-      text = element_text(size = 16, family="serif"), 
-      plot.title = element_markdown(lineheight = 1.1),
-      legend.position = c(0.15, 0.80)
-      ) 
-  
-  return(densities_plot)
-  
-}
-
-TitleAndColors <- function(policy, outcome, percentile){
-  
-  percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
-  
-  if (percentile %in% percentiles) {
-    # Define the Kaitz-p index percentile
-    percentile_name <- paste0("kaitz_pct",gsub("0.","",percentile))
-  } else {
-    print("Percentile must be 0.10, 0.25, 0.50, 0.75 or 0.90")
-  }
-  
-  if (policy == "mop"){
[…]

### 2024-06-17 main.R

@@ -1,3 +1,10 @@
+install.packages("tidyverse")
+install.packages("fixest")
+install.packages("ggtext")
+install.packages("binsreg")
+install.packages("gridExtra")
+install.packages("grid")
+
 library(tidyverse)
 library(fixest)
 library(ggtext)
@@ -22,1013 +29,10 @@ setwd(dirname(file_location)) # set path to location
 # Auxiliary functions
 source(paste(getwd(), "code/data_analysis/utilities.R", sep = "/"))
 source(paste(getwd(), "code/data_analysis/plots.R", sep = "/"))
+source(paste(getwd(), "code/data_construction/data_merge.R", sep = "/"))
+source(paste(getwd(), "code/data_analysis/results.R", sep = "/"))
 
-# Percentiles of the wage distribution
-percentiles <- c("0.10", "0.25", "0.50", "0.75", "0.90")
+# Obtain the results
+datasets(getwd())
+regressions(getwd())
 
-# Load data's location
-data_location_pmq <- paste(
-  getwd(),
-  "data/processed/joined_data_pmq.csv",
-  sep = "/"
-  )
-
-# Load the data for Must Query PDMPs
-df_pmq <- read.csv(
-  data_location_pmq,
-  header = TRUE,
-  sep = ","
-  )
-
-# Include relative-to-treatement time markers and presc per capita
-
-df_pmq <- df_pmq |>
-  dplyr::mutate(
-    relative_to_treat_pmq = time_marker - first_treatment_pmq,
-    dosage_unit_pc = round(dosage_unit / population, 3)
-  )
-
-# Select the covariate columns
-names_covar_pmq <- names(
-  df_pmq[, c(grep("_ratio$", colnames(df_pmq)),
-        grep("model_values", colnames(df_pmq)))]
-  )
-
-# Prepare the covariates (as formula)
-covariates_pmq <- paste(
-  names_covar_pmq,
-  collapse = " + "
-  )
-
-# Compute the coefficient values
-unem_rate_pmq_effects  <- df_pmq |>
-  IndEffects(yname = "unem_rate",
-             iname = "fips",
-             tname = "time_marker",
-             kname = "relative_to_treat_pmq",
-             aname = "first_treatment_pmq",
-             covariates = covariates_pmq,
-             only_full_horizon = FALSE)
[…]

### 2024-06-17 main.py

@@ -1,0 +1,35 @@
+import subprocess
+import sys
+import os
+import subprocess
+import glob
+import re
+
+# Function to install a package using pip
+def install(package):
+    subprocess.check_call([sys.executable, "-m", "pip", "install", package])
+
+# Install the 'pandas' package
+install('pandas')
+import pandas as pd
+
+# Get the current working directory
+folder_path = os.getcwd()
+
+# Define file paths for different Python scripts
+file_path_lmo = os.path.join(folder_path, 'code\\data_construction\\', 'labor_market_outcomes.py')
+file_path_mwc = os.path.join(folder_path, 'code\\data_construction\\', 'minwage_cleaning.py')
+file_path_wsec = os.path.join(folder_path, 'code\\data_construction\\', 'wages_sector.py')
+file_path_mwb = os.path.join(folder_path, 'code\\data_construction\\', 'minwage_bindingness.py')
+
+# Run the 'labor_market_outcomes.py' script using the 'python' command
+subprocess.run(['python', file_path_lmo])
+
+# Run the 'minwage_cleaning.py' script using the 'python' command
+subprocess.run(['python', file_path_mwc])
+
+# Run the 'wages_sector.py' script using the 'python' command
+subprocess.run(['python', file_path_wsec])
+
+# Run the 'wages_sector.py' script using the 'python' command (this line seems to be a duplicate)
+subprocess.run(['python', file_path_wsec])

### 2024-06-18 slides/paper.tex
readme

@@
 
+% CHECKED
 For the past two decades, the United States has been facing the worst public health crisis in its recent history.
@@
 
+% CHECKED
 This paper argues that minimum wage rates influence the relationship between opioid misuse and labor market outcomes.
@@
 
-%at the time the law is passed
+% CHECKED
+These findings contribute to the growing body of literature studying the economic effects of the opioid crisis and the policies implemented to curb it.
+Within this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation (\cite{Ruhm2018}).
+This perspective has led economists to focus on supply-side interventions to explore the causal relationship between opioid consumption and labor market outcomes.
+While well-intentioned, many of these policies have had unintended consequences, such as the rise in heroin consumption following the reformulation of OxyContin (\cite{Alpert2018}).
+However, Prescription Drug Monitoring Programs (PDMPs) have been less scrutinized.
+Recent research indicates that, in their more stringent form, these programs can effectively reduce prescription opioid misuse (\cite{Buchmueller2018}).
 
-% Condensate the whole background part into a single (two) paragraph maybe not mention lit yet
-
-These findings contribute to the growing body of literature studying the economic effects of the opioid crisis.
-Within this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation.
-This perspective has led economists to focus on supply-side interventions to explore the causal relationship between opioid consumption and labor market outcomes.
-While well-intentioned, many of these policies have had unintended consequences, such as the rise in heroin consumption following the reformulation of OxyContin.
-However, Prescription Drug Monitoring Programs (PDMPs), have been less scrutinized.
-Recent research indicates that, in their more stringent form, these programs can effectively reduce prescription opioid misuse.
-
-\begin{comment}
-    These findings contribute to the growing body of literature studying the economic effects of the opioid crisis and of the policies implemented to curb it.
-    Within this literature, the origins of the epidemic have been widely discussed, with a supply-side view of its origins now largely accepted over a demand-side explanation.
-    This perspective has led economists to focus on supply-side interventions to explore the causal relationship between opioid consumption and labor market outcomes.
-    Prescription Drug Monitoring Programs (PDMPs) 
-    Recent research indicates that, in their more stringent form, these programs can effectively reduce prescription opioid misuse.
-\end{comment}
-
+% CHECKED
 Yet, how heterogenous these effects are across labor markets remains an open question.
@@
 
-I use several data sources to obtain a distribution of coefficients of the impact of opioid misuse on labor market outcomes across different minimum wage bindingness levels.
+% CHECKED
+I use several data sources to obtain a distribution of coefficients for the impact of opioid misuse on labor market outcomes across different levels of minimum wage bindingness.
 The US Bureau of Labor Statistics (BLS) LAUS database provides county-level labor market statistics, while the US Census Bureau's Population Estimates Program (PEP) offers county-level population and demographic data.
-From \textcite{Horwitz2021} I obtain data on Prescription Drug Monitoring Programs (PDMPs). 
-The BLS's Occupational Employment and Wage Statistics (OEWS) database provides national industry-specific occupations and wages statistics, \textcite{Eckert2020} offers county-level industry-specific employment data, and the US Department of Labor's Changes in Basic Minimum Wages provides information on minimum wage rates; all of these data are relevant for constructing a measure of minimum wage bindingness.
+Data on Prescription Drug Monitoring Programs (PDMPs) is obtained from \textcite{Horwitz2021}. 
+The BLS's Occupational Employment and Wage Statistics (OEWS) database provides national industry-specific occupations and wage statistics. 
+County-level industry-specific employment data comes from \textcite{Eckert2020}, and information on minimum wage rates is sourced from the US Department of Labor's Changes in Basic Minimum Wages.
+These last three datasets are crucial for constructing a measure of minimum wage bindingness.
 
+% CHECKED
 The Centers for Disease Control (CDC) National Vital Statistics System offers data on total drug poisoning mortality by county, while the CDC's Multiple Cause of Death database provides overdose deaths at the state-year level for different drugs.
@@
 
+% CHECKED
 The results are striking in several dimensions.
 After constructing a measure of minimum wage bindingness known as the Kaitz-\(\rho\) index, I find that the impact of prescription policies on labor market outcomes is indeed heterogeneous across different levels of minimum wage bindingness.
-First, I find that policies aimed at stopping opioid misuse increased participation rates only where the minimum wage was not binding.
+First, policies aimed at stopping opioid misuse increased labor force participation rates only where the minimum wage was not binding. 
 Secondly, the increase was short-lived (less than two years), but the impact on unemployment rates was persistent.
-Moreover, their impact on prescription rates seem to be the main, but not the only, driver of these results.
-Finally, prescription policies had no labor market consequences where where the minimum wage was binding, but their impact on heroin deaths was increasing in minimum wage bindingness.
+Moreover, the impact on prescription rates appears to be the main, but not the only, driver of these results.
+Finally, prescription policies had no labor market consequences where the minimum wage was binding, but their impact on heroin deaths increased with minimum wage bindingness.
 
-Therefore, this paper sheds light on how labor market regulations, and in particular minimum wage rates, give rise to heterogeneities in the effects of the opioid crisis on the labor market, contributing to a better understanding of the ongoing crisis and the policies implemented to address it.
-More broadly, it contributes to the understanding of how labor market regulations interact with drug disorders.
+% CHECKED
+Therefore, this paper sheds light on how labor market regulations, particularly minimum wage rates, create heterogeneities in the effects of the opioid crisis on the labor market, contributing to a better understanding of the ongoing crisis and the policies implemented to address it.
+More broadly, it contributes to understanding how labor market regulations interact with drug disorders.
 
-The paper is structured as follows.
-First, it provides a summary of the opioid crisis and the effects it has had on the labor market as uncovered by the literature.
-Then, it discusses some of the policies implemented to curb the opioid misuse and its effectiveness.
-Later, I describe how different minimum wage rates can explain some of these policy consequences and the heterogeneity of the labor market outcomes. 
+% CHECKED
+The paper is structured as follows:
+First, it provides a summary of the opioid crisis and its effects on the labor market as uncovered by the literature.
+Then, it discusses some of the policies implemented to curb opioid misuse and their effectiveness.
+Later, it describes how different minimum wage rates can explain some of these policy consequences and the heterogeneity of labor market outcomes. 
 A model is introduced to formalize the intuition behind this argument.
-The employed data is discussed before presenting the empirical strategy and obtained results.
+The employed data is then discussed, followed by the empirical strategy and obtained results.
 
@@
 
+% CHECKED (the whole section)
 The following model formalizes the intuition outlined above. 
@@
 
+% CHECKED
 This section presents the main results of the paper. 
-I start describing the construction of the Kaitz-\(\rho\) index, which is used to measure the bindingness of the minimum wage.
-Then, I present the main especification of the paper, which is employed to estimate the effects of Prescription Drug Monitoring Programs on labor force participation and unemployment rates.
-The coefficients of the regression are compared across the spatial distribution of minimum wage bindingness measures.
-Finally, I repeat the same exercise using overdose deaths for different opioids as the dependent variable.
+First, I describe the construction of the Kaitz-\(\rho\) index, which measures the bindingness of the minimum wage.
+Next, I introduce the primary specification used to estimate the effects of Prescription Drug Monitoring Programs (PDMPs) on labor force participation and unemployment rates. 
+The regression coefficients are then compared across the spatial distribution of minimum wage bindingness measures.
+Finally, I repeat the analysis using overdose deaths for different opioids as the dependent variable.
 
@@
 
+% CHECKED
 The argument outlined above hinges on the assumption that the minimum wage can affect the labor market opportunities of low-educated workers.
-In other words, at least in some areas it must be the case that the minimum wage is an effective, or binding, constraint on the labor market.
-A common approach to measure this bindingness is through the Kaitz-\(\rho\) index, which can be defined as
+In other words, at least in some areas, the minimum wage must act as an effective, or binding, constraint on the labor market.
+A common approach to measure this bindingness is through the Kaitz-\(\rho\) index, which can be defined as:
 
@@
 
+% CHECKED
 where \(w^{\text{\textit{min}}}_{s(c)t}\) is the minimum wage in the state where the county \(c\) is located at time \(t\), and \(w^{\rho}_{ct}\) is the \(\rho\) percentile of the wage distribution.
@@
 Figure 1 shows the distribution of the Kaitz-\(\rho\) index across counties in the United States for the year a Must Query PDMP is implemented in their states, for each of the different percentiles of the wage distribution included in the panel.
-For the main results, I use the 10th percentile, but the findings are similar across the distribution -- the counties for which the minimum wage is more binding with respect to lower percentiles tend to be the same as those for which it is more binding with respect to higher percentiles.
+For the main results, I use the 10th percentile, but the findings are similar across the distribution -- the counties where the minimum wage is more binding at lower percentiles tend to be the same as those where it is more binding at higher percentiles.
 
-The indices are obtained using the data on the wage distribution at the occupation-industry level, and the shares of employment of each industry in each county. 
-In particular, let \(w^{\rho}_{oit}\) be the percentile \(\rho\) of the wages at the occupation \(o\) in the industry \(i\) at time \(t\). 
-Then
+% CHECKED
+The indices are obtained using data on the wage distribution at the occupation-industry level and the employment shares of each industry in each county. 
+Specifically, let \(w^{\rho}_{oit}\) be the \(\rho\) percentile of the wages in occupation \(o\) within industry \(i\) at time \(t\). 
+Then,
 
@@
 
-where \(e_{oi}\) is the number of workers in the occupation \(o\) and \(E_{it}\) is the total number of workers in the industry \(i\) at time \(t\).
-That is, by weighting the wage percentiles by the share of employment of each occupation in the industry, I obtain the wage percentile for the industry.
-After that, I compute
+% CHECKED
+where \(e_{oi}\) is the number of workers in occupation \(o\) and \(E_{it}\) is the total number of workers in industry \(i\) at time \(t\).
+That is, by weighting the wage percentiles by the share of employment in each occupation within the industry, I obtain the wage percentile for the industry.
+Next, I compute
 
@@
 
-where \(e_{ci}\) is the number of workers in the industry \(i\) in the county \(c\) and \(E_c\) is the total number of workers in the county.
-\(w^{\rho}_{ct}\) is the wage percentile for the county.
-Finally, the minimum wage rates at the state level are already included in the panel.
+% CHECKED
+where \(e_{ci}\) is the number of workers in industry \(i\) in county \(c\) and \(E_c\) is the total number of workers in the county.
+\(w^{\rho}_{ct}\) represents the wage percentile for the county.
+The state-level minimum wage rates are already included in the panel.
 
@@
 
-The main specification computes individual-level effects following \textcite{Arkhangelsky2024} and has the form
+% CHECKED
+The main specification computes individual-level effects following \textcite{Arkhangelsky2024} and is given by
 
@@
 
-Here, \(y_{ct}\) is a labor market outcome, either the labor force participation rate or the unemployment rate, \(\alpha_c\) are county fixed effects, and \(\lambda_t\) are time fixed (year-month) effects.
-Relative time is represented with \(k = t - z_c\), the time difference of the calendar time from the implementation of a PDMP in the state, represented by \(z_s\).
+% CHECKED
+where \(y_{ct}\) is a labor market outcome, either the labor force participation rate or the unemployment rate, \(\alpha_c\) are county fixed effects, and \(\lambda_t\) are time fixed effects (year-month).
+Relative time is represented by \(k = t - z_c\), the time difference from the implementation of a PDMP in the state (\(z_s\)).
 \(\tau_{ck}\) are the coefficients of interest and \(\nu_{ct}\) is an idiosyncratic shock.
 
-The covariates \(X_{ct}\) include demographic shares (race, origin, sex and age), sector shares of employment (at the 2-digits NAICS codes), and total drug poisoning mortality, which might be a relevant variable in policy-making.
+% CHECKED
+The covariates \(X_{ct}\) include demographic shares (race, origin, sex and age), sector shares of employment (at the 2-digits NAICS code level), and total drug poisoning mortality, which might be a relevant variable in policy-making.
 Crucial for identification, I assume that the state-level rollout of the PDMPs is exogenous at the county level after conditioning on the covariates.
-Moreover, this strategy requires that there is no anticipation of the policy up to \(h\) periods. 
+Additionally, this strategy requires that there is no anticipation of the policy up to \(h\) periods. 
 I assume no anticipation whatsoever, so that the coefficients are identified from the very start of the policy.
-In this context, this means that no addicted individual decided to look for a job or switch to heroin because of the expected implementation of the PDMP.
+In other words, no addicted individual decided to look for a job or switch to heroin due to the expected implementation of the PDMP.
 
+% CHECKED
 After running (10), a distribution of individual (county-level) coefficients is obtained.
@@
 
+% CHECKED
 Using the labor force participation rate as the dependent variable, the results of the main specification are displayed in Figure 2.
-The figure shows the distribution of the coefficients of the impact of the passing of the Must Query PDMPs on labor force participation rates across different levels of minimum wage bindingness, for different periods relative to treatment.
-In the upper right panel we can see the immediate impact of the policy.
-For those counties where the minimum wage is less binding, the effect is positive (1.37 p.p.) and significant (0.513, 2.22) at conventional levels.
+The figure shows the distribution of the coefficients of the impact of the Must Query PDMPs on labor force participation rates across different levels of minimum wage bindingness for various periods relative to treatment.
+
+% CHECKED
+In the upper right panel, we see the immediate impact of the policy.
+For counties where the minimum wage is less binding, the effect is positive (1.37 p.p.) and significant (0.513, 2.22) at conventional levels.
 The effects become smaller and more uncertain as the minimum wage becomes more binding.
-For those counties in which the minimum wage has a higher impact on labor demand, the point estimate becomes negative (-0.693) but not significant (-2.54, 1.16).
+In counties where the minimum wage has a higher impact on labor demand, the point estimate becomes negative (-0.693) but not significant (-2.54, 1.16).
 
-The effect seems to be persistent in the short run (upper right and lower right panels), but fades away after two years (lower right)\footnote{The confidence interval of the most binnding counties is not displayed in this last panel for visualization purposes.
+% CHECKED
+The effect appears to be persistent in the short run (upper right and lower right panels) but fades away after two years (lower right)\footnote{The confidence interval for the most binding counties is not displayed in this last panel for visualization purposes.
 The values are (-3.10, 1.90)}.
 Figure 3 shows the time averages of the coefficients for two years after the law is passed.
-The results remain positive and significant for those counties in which the minimum wage is less binding (1.06 p.p., (0.584, 1.55)), while they are not significant for those where the minimum wage was more binding.
+The results remain positive and significant for the counties where the minimum wage is less binding (1.06 p.p., (0.584, 1.55)), while they are not significant for those where the minimum wage was more binding.
 
-The results seem to be driven mostly by the impact of the policy on prescription rates.
-Although the overlap between the passing of the policies (Table 1) and the data for prescription rates (2006-2014) is scarce, we can still obtain som estimates on the impact of Must Query PDMPs on prescription rates per capita.
-Figure 4 runs the model (10) with the prescription rates per capita as the dependent variable.
-The two months after the treatment is passed are displayed.
-We see that the policy is able to decrease the amount of prescriptions, with the biggest drops occuring where the minimum wage is the least and the most binding (the effects in these areas are not statistically different).
-Although the target of these policies where not to decrease prescriptions \textit{per se}, but only tackling misuse cases, the results suggest that the policies are effective in reducing the amount of opioids prescribed on the aggregate.
+% CHECKED
+The results seem to be driven primarily by the impact of the policy on prescription rates.
+Although the overlap between the passing of the policies (Table 1) and the data for prescription rates (2006-2014) is limited, we can still obtain some estimates of the impact of Must Query PDMPs on prescription rates per capita.
+Figure 4 runs the model (10) with prescription rates per capita as the dependent variable, showing the two months after the treatment is passed.
+The policy is effective in decreasing the amount of prescriptions, with the biggest drops occurring where the minimum wage is the least and the most binding (the effects in these areas are not statistically different).
+Although the target of these policies was not to decrease prescriptions \textit{per se}, but to tackle misuse cases, the results suggest that the policies are effective in reducing the overall amount of opioids prescribed.
 
-Moreover, this seems to be the main channel through which the policies affect labor market outcomes: Figure 8 in the appendix shows that the relation between labor market outcomes and minimum wage bindingness is not significant when controlling for prescription rates.
-Figures 9 and 10 in the Appendix show the the point estimates of the binscatter for different Kaitz indices, for six months and two years after the policy is passed.
-It is easy to see that the results do not change much across different measures, this being especially true the closer in time the treatment was implemented.
+% CHECKED
+Moreover, this seems to be the main channel through which the policies affect labor market outcomes.
+Figure 8 in the appendix shows that the relation between labor market outcomes and minimum wage bindingness is not significant when controlling for prescription rates.
+Figures 9 and 10 in the appendix show the point estimates of the binscatter for different Kaitz indices for six months and two years after the policy is passed.
+The results do not change much across different measures, especially the closer in time the treatment was implemented.
 
@@
 
+% CHECKED
 The analysis is repeated using the unemployment rate as the dependent variable.
-The results are displayed in the same fashion in Figure 5.
-The distribution of the coefficients reveal that almost all the positive effect it was found in the labor force participation rates translated into an increase in the unemployment rate.
-For those counties where the minimum wage is less binding, the effect is positive (1.02 p.p.) and significant (1.46, 0.575) at conventional levels. 
-The point estimate is only 0.35 p.p. below the one it was found for the labor force participation rate.
-Although a positive effect is also found for those counties where the minimum wage is more binding (0.640 p.p.), the effect is not significant (-0.0252, 1.31).
+The results are displayed in the same manner in Figure 5.
+The distribution of the coefficients reveals that almost all the positive effects found in the labor force participation rates translated into an increase in the unemployment rate.
+For counties where the minimum wage is less binding, the effect is positive (1.02 p.p.) and significant (1.46, 0.575) at conventional levels. 
+The point estimate is only 0.35 p.p. lower than the one found for the labor force participation rate.
+Although a positive effect is also found for counties where the minimum wage is more binding (0.640 p.p.), the effect is not significant (-0.0252, 1.31).
 
-These results shouldn't be surprising, at least in the short run: if people with severe opioid addictions start looking for a job after the implementation of a PDMP, they might not be able to find one immediately even in those counties where the minimum wage is less binding.
+% CHECKED
+These results shouldn't be surprising, at least in the short run: if people with severe opioid addictions start looking for a job after the implementation of a PDMP, they might not be able to find one immediately even in counties where the minimum wage is less binding.
 Strikingly though, the impact on unemployment persists after two years.
 The lower right panel shows that, even after the increase in labor force participation rates fades away, the increase in unemployment rates remains high (2.77 p.p.) and significant (1.87, 3.68) where the minimum wage is less binding.
-Whether the causes of this persistence are related to opioid consumption (through, say, relapsing) or to characterisctics of the workers or the labor market (say, if these workers face more frictional unemployment) remains an open question.
+Whether the causes of this persistence are related to opioid consumption (through, say, relapsing) or to characteristics of the workers or the labor market (e.g., if these workers face more frictional unemployment) remains an open question.
+
+% CHECKED
+The time average of the coefficients, displayed in Figure 6, confirm our previous result.
+The effect of the policy on unemployment rates is positive and significant for counties where the minimum wage is less binding (1.25 p.p., (0.97, 1.52)), while it is not significant for those where the minimum wage is more binding (0.025 p.p., (-0.11, 0.16)).
+Again, the results seem to be driven by the impact of the policy on prescription rates, especially in the medium run: Figure 11 in the appendix shows that the impact stops being significantly different from zero after one year once we control for prescriptions per capita.
+Figures 12 and 13 in the appendix show the binscatter for different Kaitz indices for six months and two years after the policy is passed.
+The results do not change substantially across different measures.
 
@@
 
-The time average of the coefficients, displayed in Figure 6, confirm our previous result.
-The effect of the policy on unemployment rates is positive and significant for those counties where the minimum wage is less binding (1.25 p.p., (0.97, 1.52)), while it is not significant for those where the minimum wage is more binding (0.025 p.p., (-0.11, 0.16)).
-Again, the results seem to be driven by the impact of the policy on prescription rates, especially in the medium run: Figure 11 in the Appendix shows that the impact stops being significantly different from zero after one year once we control for prescriptions per capita.
-Figures 12 and 13 in the Appendix show the binscatter for different Kaitz indices, for six months and two years after the policy is passed.
-The results do not change substantially across different measures.
-
 \subsection*{Overdose deaths}
 
-Finally, and in accordance with the model outlined above, I show some evidence that the rollout of Must Query PDMPs has had a positive impact on the consumption of illegal opioids in those counties where the minimum wage was more binding.
-For this purpose, I run the same estimation strategy as before, but using the number of overdose deaths per capita for different opioids as the dependent variable. 
+% CHECKED
+Finally, and in accordance with the model outlined above, I present evidence that the rollout of Must Query PDMPs has had a positive impact on the consumption of illegal opioids in counties where the minimum wage was more binding.
+For this purpose, I apply the same estimation strategy as before, but use the number of overdose deaths per capita for different opioids as the dependent variable. 
 The data is at the state-year level, and the number of total poisoning deaths is not used as a control variable.
-Figure 7 displays the scatter plot of the time average of the coefficients on the Kaitz-0.10 index of the states, for the two years after the policy is passed.
-The opioids included are heroin, methadone, other opioids and other synthetic narcotics, following the ICD-10 classification.
+Figure 7 displays the scatter plot of the time average of the coefficients on the Kaitz-0.10 index for the states, for the two years after the policy is passed.
+The opioids included are heroin, methadone, other opioids, and other synthetic narcotics, following the ICD-10 classification.
 
+% CHECKED
 The coefficients are regressed on the Kaitz index and a constant. 
 Tables 2-5 show the values of the regression coefficients for each of the different opioids.
-Although admittedly weaker than the evidence presented in previous sections due to data constaints, it is interesting that the only marginally significant relation is found for heroin.
-This suggest that the positive relation between black market switching and low labor demand is present, but we cannot assert this relationship with the same degree of certainty.
+Although the evidence presented here is admittedly weaker than in previous sections due to data constraints, it is interesting that the only marginally significant relation is found for heroin.
+This suggests that the positive relation between black market switching and low labor demand is present, but we cannot assert this relationship with the same degree of certainty.
 
-It seems implausible that part of the effect is absorbed by the consumption of other opioids.
-Methadone is often used as a treatment for opioid adiction and can be legally used under a doctor's supervision.
-Other opioids include drugs such as morphine and codeine, but also semi-synthetic opioid analgesics (e.g.
-drugs such as oxycodone, hydrocodone, hydromorphone, and oxymorphone), some of them targeted by the PDMPs.
-Finally, although the category of other synthetic narcotics includes fentanyl and its analogues, its use didn't used to be as widespread as it is today during the period considered.
+% CHECKED
+It seems unlikely that part of the effect is absorbed by the consumption of other opioids.
+Methadone is often used as a treatment for opioid addiction and can be legally used under a doctor's supervision.
+Other opioids include drugs such as morphine and codeine, but also semi-synthetic opioid analgesics (e.g., oxycodone, hydrocodone, hydromorphone, and oxymorphone), some of which are targeted by the PDMPs.
+Finally, although the category of other synthetic narcotics includes fentanyl and its analogues, its use was not as widespread during the period considered as it is today.
 
@@
 
+% CHECKED
 This paper contributes to the understanding of how labor market regulations, particularly minimum wage rates, shape the relationship between opioid misuse and labor market outcomes.
-By leveraging the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) across states, this study provides a nuanced view of how labor market opportunities influence the incentives for opioid users to quit or switch to illegal alternatives.
+By leveraging the staggered rollout of Must Query Prescription Drug Monitoring Programs (PDMPs) across states, this paper provides evidence of how labor market opportunities influence the incentives for opioid users to quit or switch to illegal alternatives.
 
+% CHECKED
 The findings reveal that labor force participation and unemployment rates increased as a result of prescription policies only in areas where the minimum wage was not binding.
-In contrast, a surge in heroin consumption occurred where the minimum wage was binding, highlighting the complex interplay between labor market conditions and opioid misuse.
-These results underscore the importance of considering labor market regulations when evaluating the economic impacts of the opioid crisis.
+In contrast, a surge in heroin consumption occurred where the minimum wage was binding.
+These results underscore the importance of considering labor market regulations when evaluating the economic impacts of the opioid crisis, as they can generate heterogeneities often overlooked in the literature.
+Hence, this study adds to the growing body of research that examines the economic effects of the opioid epidemic, emphasizing the need for a comprehensive approach that considers the incentives faced by addicted individuals in different labor market contexts.
 
-Furthermore, this study adds to the growing body of literature that examines the economic effects of the opioid epidemic.
-While previous research has primarily focused on the direct consequences of opioid misuse on labor market outcomes, this paper emphasizes the role of labor market regulations in mediating these effects.
-The heterogeneity in the impact of prescription policies across different levels of minimum wage bindingness suggests that policies aimed at curbing opioid misuse must account for the broader economic context to be effective.
-
-In summary, the analysis demonstrates that higher minimum wages, by reducing labor demand for low-skilled workers, can diminish the incentives for opioid users to quit, thereby exacerbating the shift towards illegal alternatives like heroin.
-These insights highlight the need for a comprehensive approach to addressing the opioid crisis, one that integrates labor market policies with public health interventions to effectively mitigate the epidemic's adverse effects on both health and economic outcomes.
-
-Future research should continue to explore the interplay between labor market regulations and health crises, considering the potential unintended consequences of well-intentioned policies.
-By doing so, policymakers can design more effective interventions that address the root causes of public health crises while also supporting economic stability and growth.
-
-[MOVING FORWARD]
+% CHECKED
+Future research should address some limitations of this study.
+First, more detailed data on illegal opioid consumption and effective minimum wages (which can vary in metropolitan areas) could provide more precise estimates of the effects of labor market regulations on the opioid crisis's impact on labor outcomes.
+More importantly, the mechanisms through which these regulations affect the incentives of addicted individuals should be further explored.
+The relevance of minimum wage rates hinges on their impact on labor demand, so further exploration of this mechanism, including the impact of these policies on vacancies, would be valuable.
+Finally, a more dynamic model of workers' incentives might account for the short-run effects on labor rates participation rates and the persistent increase in unemployment rates observed.
 
@@
     \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate.png}
-    \caption{}
+    \caption{Results for (10) using labor force participation rates as the dependent variable.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:lab_for_rate}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_ta.png}
-    \caption{}
+    \caption{Time average of the coefficients of Figure 2, for two years after the policy is passed.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:lab_for_rate_ta}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_prescriptions.png}
-    \caption{}
+    \caption{Distribution of the unit-level effects using prescriptions per capita as the dependent variable.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:prescriptions}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate.png}
-    \caption{}
+    \caption{Results for (10) using unemployment rates as the dependent variable.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:unemp_rate}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_ta.png}
-    \caption{}
+    \caption{Time average of the coefficients of Figure 5, for two years after the policy is passed.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:unemp_rate_ta}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_heroin.png}
-    \caption{}
+    \caption{Results for (10) using different opioids drug overdose deaths as the dependent variables, at the state-year level. The coefficients of the linear regression are shown in Tables 2-5.}
     \label{fig:drugs}
@@
     \end{tabular}
-    \caption{Regression Results: Other opioids Overdose Deaths}
+    \caption{Regression Results: Other Opioids Overdose Deaths}
     \label{tab:regression_results_otheropioids}
@@
     \end{tabular}
-    \caption{Regression Results: synthetic opioids Overdose Deaths}
+    \caption{Regression Results: Synthetic Opioids Overdose Deaths}
     \label{tab:regression_results_synthetic}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_p.png}
-    \caption{}
+    \caption{Results for (10) using labor force participation rates as the dependent variable and adding prescriptions per capita as controls.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:lab_for_rate_p}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_comp6.png}
-    \caption{}
+    \caption{Comparison of the point estimates for the labor force participation rate across percentiles of the wage distribution, six months after treatment.
+The bins are computed optimally following Cattaneo et al.
+(2024).}
     \label{fig:lab_for_rate_comp6}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_lab_for_rate_comp24.png}
-    \caption{}
+    \caption{Comparison of the point estimates for the labor force participation rate across percentiles of the wage distribution, two years after treatment.
+The bins are computed optimally following Cattaneo et al.
+(2024).}
     \label{fig:lab_for_rate_comp24}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_p.png}
-    \caption{}
+    \caption{Results for (10) using unemployment rates as the dependent variable and adding prescriptions per capita as controls.
+The bins are computed optimally following Cattaneo et al.
+(2024).
+A confidence interval is displayed around the point estimates, at 95\% level.}
     \label{fig:unemp_rate_p}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_comp6.png}
-    \caption{}
+    \caption{Comparison of the point estimates for the unemployment rate across percentiles of the wage distribution, six months after treatment.
+The bins are computed optimally following Cattaneo et al.
+(2024).}
     \label{fig:unemp_rate_comp6}
@@
     \includegraphics[width=0.8\textwidth]{pmq10_unemp_rate_comp24.png}
-    \caption{}
+    \caption{Comparison of the point estimates for the unemployment rate across percentiles of the wage distribution, two years after treatment.
+The bins are computed optimally following Cattaneo et al.
+(2024).}
     \label{fig:unemp_rate_comp24}

### 2024-06-18 code/data_analysis/results.R

@@ -560,7 +560,6 @@ regressions <- function(location) {
 
   # Prescriptions
 
-
   data_location_prescriptions_pmq <- paste(
     location,
     "data/processed/prescription_monthly_pmq.csv",
